---
title: "ggstates, ggfips"
author: "Evangeline Reynolds"
date: "11/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem: positioning with SF feels spooky and data prep can feel effortful

W/ ggplot2, we are able to say 'where' with x and y.  These are positional aesthetics.  With SF the 'where' is implied with the geometry column. This feels non-declarative and spooky to me. 

Also, it often requires pre-processing to work with data you'd like to represent.

When we ask the question 'where' in a geographical context, a human being will respond with the name of a region; not coordinates and not a set of coordinates that define a region.  Is it possible to create an interface where you define 'where' with geographical names or other ids?  For example: `aes(country = my_countries_variable)` could be used with a geom/stat like geom_polygon_countries().


See step-by-step:

https://evamaerey.github.io/ggplot2_grammar_guide/geom_sf.html#1


```{r, eval = F}
library(tidyverse)
 
nato_names <- c("Albania", "Belgium", "Bulgaria", "Canada", "Croatia", "Czech Republic", "Denmark", "Estonia", "France", "Germany", "Greece", "Hungary",  
                "Iceland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Montenegro", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Turkey", "United Kingdom", "United States")  
library(gapminder)  
gapminder %>%  
  filter(year == 2002) %>%  
  rename(name = country) ->  
gapminder_2002_prepped  

rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%  
  select(name, pop_est, gdp_md_est,  
         continent, geometry) %>%  
  filter(name %in% nato_names) %>%  
  left_join(gapminder_2002_prepped, by = "name") ->  
  nato_countries  

ggplot(data = nato_countries) +  
  # This part feels spooky to me 
  ggplot2::geom_sf(data = nato_countries) +  
    aes(fill = lifeExp) +  
    scale_fill_viridis_c(option = "magma", direction = -1) +  
  ggplot2::coord_sf(xlim = c(-175, 47.5),  
           ylim = c(23, 85),  
           expand = FALSE) +  
  labs(title = "Life Expectancy, 2002",  
         subtitle = "NATO Member States",  
         fill = "Share of GDP (%)") +  
  theme_bw() +  
  theme(plot.title =  
            element_text(hjust = 0.5)) +  
  theme(plot.subtitle =
            element_text(hjust = 0.5))
```





## A different API?

Here is what I think might be an easy to use API.

```
ggworld(data = gapminder_2002, projection = 'a good one') + 
  aes(country = country, # positioning aesthetic
      fill = lifeExp) + 
  geom_polygon_country()
```

Note: country names are really hard. Other codes could be used.  One could try to use the wonderful countrycodes as well when 

# Proof of concept? (without SF, and with easy statenames case)

Because of my limited knowledge of SF, let's start w/ the x and y inputs approach to mapping with ggplot2 and see if we can cook something up. 

## Step 00: See how someone gets it done

I was here:
https://remiller1450.github.io/s230s19/Intro_maps.html

## Step 0: get it done w/ ggplot2

Look at mapping w/ ggplot2, no sf, we use geom_polygon and the x, y and group aesthetics).  It is not very efficient (lots of redundant info in the ), but we won't worry about that for this demonstration.  A little rewrite to get inspired.

```{r}
library(ggplot2) 
library(dplyr)
library(maps)
MainStates <- map_data("state")
StatePopulation <- read.csv("https://raw.githubusercontent.com/ds4stats/r-tutorials/master/intro-maps/data/StatePopulation.csv", as.is = TRUE)

MergedStates <- inner_join(MainStates, StatePopulation, by = "region")

StatePopulation %>% # becomes ggus()
  right_join(MainStates) %>% # becomes ggus()
  ggplot() + # becomes ggus()
  geom_polygon() + # becomes geom_polygon state
  aes(x = long, # becomes state = region
      y = lat, # becomes state = region
      group = group, # becomes state = region
      fill = population/1000000) + 
  coord_map(projection = "orthographic", 
            orientation = c(41, -74, 0))
```


# Step 1: compute function and test

```{r}
library(maps)
MainStates <- map_data("state")
# AllCounty <- map_data("county")
# MainCities <- filter(us.cities, long>=-130)

MainStates %>% 
  rename(state = region) %>% 
  rename(which_polygon = group) ->
continuous_states

continuous_states %>% 
  count(state, which_polygon) %>% 
  sample_n(10)

StatePopulation <- read.csv("https://raw.githubusercontent.com/ds4stats/r-tutorials/master/intro-maps/data/StatePopulation.csv", as.is = TRUE)

compute_state <- function(data, scales){
 
  data %>% 
    inner_join(continuous_states, multiple = "all") %>% 
    mutate(x = long, y = lat) %>% 
    arrange(order) %>%
    # mutate(group = paste(group, ".", which_polygon)) %>%
    # group_by(which_polygon) %>%
    data.frame()

}

StatePopulation %>% 
  rename(state = region) %>% 
  mutate(group = row_number()) %>% 
  compute_state() ->
computed

computed %>% sample_n(30)

computed %>% 
  ggplot() + 
  aes(x = x, y = y) + 
  geom_point() + 
  geom_polygon(color = "white") +
  aes(group = which_polygon)

my_setup_data <- function(data, params){
                                    if(data$group[1] == -1){
                                      nrows <- nrow(data)
                                      data$group <- seq_len(nrows)
                                    }
                                    data
}
```


# Step 2: pass to ggproto

```{r}
StatPolygonstate <- ggplot2::ggproto(`_class` = "StatPolygonstate",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("state"),
                                  # setup_data = my_setup_data,
                                  compute_panel = compute_state,
                                  default_aes = aes(group = after_stat(which_polygon))
                                  )
```


# Step 3: write geom_* function

```{r}
geom_polygon_state <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatPolygonstate,  # proto object from step 2
    geom = ggplot2::GeomPolygon,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

# step 4: test geom_* function, celebrate!

```{r, eval = T}
StatePopulation %>% #count(region)
  mutate(region = region %>% stringr::str_trim() %>% as.factor()) %>% 
  ggplot(data = .) + 
  aes(state = region, # state indicates position instead of x and y 
      fill = population) +
  geom_polygon_state() + 
  coord_map(projection = "orthographic", 
            orientation = c(41, -74, 0)) + 
  geom_polygon_state(data = . %>% 
                       filter(region == "new york"),
                     color = "red" ) + 
  geom_polygon_state(data = . %>% 
                       filter(region == "colorado"),
                     color = "green" ) + 
  scale_fill_viridis_c(option = "magma", end = .9) + 
  ggstamp::stamp_point(y = 30.2, x = -97.7, 
                       color = "goldenrod1",
                       size = 2)
```


# Step 5: What's not working and taking a break.

1. Solved!  "Not sure what's going on w/ merge so that we only get the 4 states," now much better!  there was a grouping variable in the polygons data set named 'group' - this really messed with things because ggplot2 uses that var internally.  We renamed!
2. state = var, var must be factor (not character).  I've seen this before and I'd like that not to be the case, but fine for now. 
3. There are subregion.  You can see the tell tail signs. Seems like setup_data should help w/ group management
4. Take a break! 



---

# 0 County - ggfips?

```{r}
AllCounty <- map_data("county")

ggplot() + geom_polygon(data=AllCounty, aes(x=long, y=lat, group=group),
                color="darkblue", fill="lightblue", size = .1 ) +
  
          geom_polygon( data=MainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue",  size = .15, alpha = .3)
```

# Circle pack seems similar...


https://evamaerey.github.io/mytidytuesday/2022-05-28-circle-pack/circle_pack_stipple.html


# Example simple feature to data frame...

```{r}
library(sf)
library(sp)
library(tidyverse)

## using a MULTIPOLYGON data set supplied with library(sf)
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))


nc %>% 
  ggplot() + 
  geom_sf(aes(fill = AREA))

# pull out all the data frames that define polygons
nc_polygons <- nc$geometry %>% purrr::flatten() %>% purrr::flatten()
# There's sometimes multiple polygons associated with a row
# 
nsubgroup <- nc$geometry %>% purrr::map_dbl(.f = length)

tibble(nsubgroup) %>% 
  mutate(group = 1:n()) %>% 
  uncount(weights = nsubgroup) %>% 
  group_by(group) %>% 
  mutate(subgroup = 1:n()) %>% 
  mutate(region = paste0(group,'.', subgroup)) %>% 
  ungroup() %>% 
  mutate(polygon = 1:n()) ->
groups_and_polygons_nc


nc_df <- nc
nc_df$geometry <- NULL
str(nc_df)
nc_df$group <- 1:nrow(nc_df)


polygons_out <- data.frame()

for (i in 1:length(nc_polygons)){
  
  temp <- nc_polygons[[i]] %>% 
    data.frame() %>% 
    mutate(polygon = i)
  
  polygons_out = bind_rows(polygons_out, temp)
  }

polygons_out %>% 
  left_join(groups_and_polygons_nc) %>% 
  left_join(nc_df) %>% 
  ggplot() + 
  aes(x = X1, y = X2 , group = polygon, fill = AREA) +
  geom_polygon() + 
  coord_map(projection = "orthographic", 
            orientation = c(31, -74, -3)) 
```
# Write functions for ggnorthcarolina

```{r}
# imagine you only have nc_df
nc_df %>% head()

geometries_frame_to_id_group <- function(geometries_frame){

    nsubgroup <- geometries_frame$geometry %>% purrr::map_dbl(.f = length)

    tibble(nsubgroup) %>% 
      mutate(mgroup = 1:n()) %>% # row number 
      uncount(weights = nsubgroup) %>% 
      group_by(mgroup) %>% 
      mutate(subgroup = 1:n()) %>% 
      mutate(region = paste0(mgroup,'.', subgroup)) %>% 
      ungroup() %>% 
      mutate(polygon = 1:n()) %>% 
      ungroup()

}

geometries_frame_to_id_group(nc)

geometries_frame_to_df <- function(geometries_frame){

      polygons_flat <- data.frame()
    
      polygons <- geometries_frame$geometry %>% purrr::flatten() %>% purrr::flatten()
      
    for (i in 1:length(nc_polygons)){
      
      temp <- nc_polygons[[i]] %>% 
        data.frame() %>% 
        mutate(polygon = i)
      
      polygons_flat = bind_rows(polygons_flat, temp)
  

}

  polygons_flat %>% 
    rename(x = X1, 
           y = X2)
}
  
geometries_frame_to_df(nc)

geometries_frame_to_ggplot_reference <- function(geometries_frame = nc){
  
  keep_frame <- geometries_frame
  
  keep_frame$geometry = NULL
  
  keep_frame %>% 
    data.frame() %>% 
    janitor::clean_names() %>% 
    mutate(mgroup = row_number()) %>% 
    left_join(geometries_frame_to_id_group(nc)) %>% 
    left_join(geometries_frame_to_df(nc))
  
}
  

geometries_frame_to_id_group(nc) %>% 
  count(mgroup)
  
nc_ggplot2_reference <- geometries_frame_to_ggplot_reference(nc)
  

compute_county_nc <- function(data, scales){
 
  data %>% 
    inner_join(nc_ggplot2_reference, multiple = "all") 

}

nc_df %>% 
  rename(fips = FIPS) %>% 
  compute_county_nc()

```


# Step 2: pass to ggproto

```{r}
StatPolygonnccounty <- ggplot2::ggproto(`_class` = "StatPolygonnccounty",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("fips"),
                                  # setup_data = my_setup_data,
                                  compute_panel = compute_county_nc,
                                  default_aes = aes(group = after_stat(polygon))
                                  )
```


# Step 3: write geom_* function

```{r}
geom_polygon_nccounty <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatPolygonnccounty,  # proto object from step 2
    geom = ggplot2::GeomPolygon,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

# step 4: test geom_* function, celebrate!

```{r, eval = T}
nc_df %>% #count(region)
  mutate(FIPS = FIPS %>% as.factor()) %>% 
  ggplot(data = .) + 
  aes(fips = FIPS, # state indicates position instead of x and y 
      fill = PERIMETER) +
  geom_polygon_nccounty() + 
  coord_map(projection = "orthographic", 
            orientation = c(41, -74, 0))
```
