---
title: "Exploded code"
subtitle: "Using flipbookr and xaringan"
author: "Me"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = F, warning = FALSE, comment = "", cache = F)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping 
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```


---

# existing ggplot2

plotting observations and linear model is fairly easy with existing ggplot2 code.

---


`r flipbookr::chunk_reveal("start", left_assign = "detect")`


```{r start, message=F, warning=F, include = FALSE}
library(tidyverse)
anscombe %>% 
  ggplot() + 
  aes(x = x3, y = y3) + 
  geom_point() +
  geom_smooth(method = lm,
              se = F)
```

---

# plotting predictions

plotting the predicted values of the observations is a bit harder, but we can do it with pre calculation. 

---


`r flipbookr::chunk_reveal("lm", left_assign = "detect")`


```{r lm, include = FALSE}
# calc fitted and add points at fitted
lm(formula = anscombe$y3 ~ anscombe$x3) ->
  my_model

data.frame(x3 = anscombe$x3, 
           predicted_y3 = my_model$fitted.values) ->
  fitted_data

anscombe %>% 
  ggplot() + 
  aes(x = x3, y = y3) + 
  geom_point() + 
  geom_smooth(method = lm, 
              se = F) + 
  geom_point(data = fitted_data,
             aes(x = x3, 
                 y = predicted_y3),
             color = "blue") +
  geom_text(data = fitted_data,
             aes(x = x3, 
                 y = predicted_y3,
                 label = round(predicted_y3, 
                               digits = 1)
                 ),
             color = "midnightblue") 
```

---

# Or extend ggplot2


---


`r flipbookr::chunk_reveal("extend", left_assign = "detect")`


```{r extend, include = FALSE}
StatOlsfittedpoint <- ggplot2::ggproto(`_class` = "StatOlsfittedpoint",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = function(data, scales) {
                                    
                                    model <- lm(formula = data$y ~ data$x)
                                  
                                    data.frame(x = data$x,
                                               y = model$fitted.values)
                                  }

)


geom_point_lm_fitted <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatOlsfittedpoint, 
    geom = ggplot2::GeomPoint, 
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


---


`r flipbookr::chunk_reveal("use", left_assign = "detect")`


```{r use, include = FALSE}
library(ggplot2)
ggplot(cars) + 
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method = lm) + 
  geom_point_lm_fitted(color = "plum4")
```


---


`r flipbookr::chunk_reveal("extend2", left_assign = "detect")`


```{r extend2, include = FALSE}
StatOlsfitted <- ggplot2::ggproto(`_class` = "StatOlsfitted",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = function(data, scales) {
                                    
                                    model <- lm(formula = data$y ~ data$x)
                                  
                                    data.frame(x = data$x,
                                               y = model$fitted.values,
                                               label = 
                                                 round(x = model$fitted.values,
                                                       digits = 2))
                                  }

)


geom_text_lm_fitted <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatOlsfitted, 
    geom = ggplot2::GeomText, 
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


---


`r flipbookr::chunk_reveal("use2", left_assign = "detect")`


```{r use2, include = FALSE}
library(ggplot2)
ggplot(cars) + 
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method = lm) + 
  geom_text_lm_fitted(color = "plum4")
```

