---
title: "layer one liners"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Campitelli:  stat_rasa (now stat_group() and extending to stat_panel()) 

https://eliocamp.github.io/codigo-r/en/2018/05/how-to-make-a-generic-stat-in-ggplot2/

### Backend StatRasa, stat_group() 

Define group-wise computation.

```{r}
library(tidyverse)

compute_rasa <- function(data, scales, compute_fun, fun.args) {
     # Change default arguments of the function to the 
     # values in fun.args
     args <- formals(compute_fun)
     for (i in seq_along(fun.args)) {
        if (names(fun.args[i]) %in% names(fun.args)) {
           args[[names(fun.args[i])]] <- fun.args[[i]]
        } 
     }
     formals(compute_fun) <- args
     
     # Apply function to data
     compute_fun(data)
}


StatRasagroup <- ggplot2::ggproto(
  `_class` = "StatRasagroup", 
  `_inherit` = ggplot2::Stat,
  compute_group = compute_rasa
  )

# stat function used in ggplot - but reordered from conventional!
stat_group <- function(compute_fun = NULL,
                       geom = "point", 
                       mapping = NULL, 
                       data = NULL,
                       position = "identity",
                       required_aes = NULL, 
                       default_aes = NULL, 
                       dropped_aes = NULL,
                      ...,
                      show.legend = NA,
                      inherit.aes = TRUE) {
   # Check arguments 
   if (!is.function(compute_fun)) stop("compute_group_fun must be a function")
   
   # Pass dotted arguments to a list
   fun.args <- match.call(expand.dots = FALSE)$`...`
   
   if(!is.null(required_aes)){StatRasagroup$required_aes <- required_aes}
   if(!is.null(default_aes)){StatRasagroup$default_aes <- default_aes}
   if(!is.null(dropped_aes)){StatRasagroup$dropped_aes <- dropped_aes}
   
   ggplot2::layer(
      data = data,
      mapping = mapping,
      stat = StatRasagroup,
      geom = geom,
      position = position,
      show.legend = show.legend,
      inherit.aes = inherit.aes,
      check.aes = FALSE,
      check.param = FALSE,
      params = list(
         compute_fun = compute_fun, 
         fun.args = fun.args,
         na.rm = FALSE,
         ...
      )
   )
}
```



---

# extending: Define panel-wise computation

```{r}
library(tidyverse)


StatRasapanel <- 
  ggplot2::ggproto("StatRasapanel", 
                   ggplot2::Stat,
                   compute_panel = compute_rasa)


# stat function used in ggplot - we reorder from conventional
stat_panel <- function(compute_fun = NULL, 
                       geom = "point", 
                       mapping = NULL, data = NULL,
                      position = "identity",
                      required_aes = NULL,
                      default_aes = NULL,
                      dropped_aes = NULL,
                      ...,
                      show.legend = NA,
                      inherit.aes = TRUE) {
   # Check arguments 
   if (!is.function(compute_fun)) stop ("compute_fun must be a function")
   
   # Pass dotted arguments to a list
   fun.args <- match.call(expand.dots = FALSE)$`...`
   
   if(!is.null(required_aes)){StatRasapanel$required_aes <- required_aes}
   if(!is.null(default_aes)){StatRasapanel$default_aes <- default_aes}
   if(!is.null(dropped_aes)){StatRasapanel$dropped_aes <- dropped_aes}

   ggplot2::layer(
      data = data,
      mapping = mapping,
      stat = StatRasapanel,
      geom = geom,
      position = position,
      show.legend = show.legend,
      inherit.aes = inherit.aes,
      check.aes = FALSE,
      check.param = FALSE,
      params = list(
         compute_fun = compute_fun, 
         fun.args = fun.args,
         na.rm = FALSE,
         ...
      )
   )
}
```





### stat_group(fun = summarize_xy)

```{r}
group_means <- function(data){
  
 data %>% 
    summarise(x = mean(x),
              y = mean(y))
  
}

geom_means <- function(...){
  stat_group(compute_fun = group_means, ...)
}

mtcars |>
  ggplot() + 
  aes(x = wt,
      y = mpg) + 
  geom_point() + 
  geom_means(size = 6)

last_plot() + 
  aes(color = factor(cyl))
```

```{r}
group_label_at_center <- function(data){
  
 data %>% 
    group_by(label) %>% 
    summarise(x = mean(x, na.rm = T),
              y = mean(y, na.rm = T))
  
}

geom_center_label <- function(...){
  stat_group(compute_fun = group_label_at_center, geom = GeomLabel, ...)
}


palmerpenguins::penguins |>
  ggplot() +
  aes(x = bill_length_mm,
      y = bill_depth_mm) +
  geom_point() +
  aes(label = "All") +
  geom_center_label()


last_plot() +
  aes(color = species, label = species)


geom_center_text <- function(...){
  stat_group(compute_fun = group_label_at_center, geom = GeomText, ...)
}


palmerpenguins::penguins |>
  ggplot() +
  aes(x = bill_length_mm,
      y = bill_depth_mm) +
  geom_point(x = 42, aes(color = bill_length_mm)) +
  aes(color = species) +
  geom_center_text(color = "Black",
    aes(label = species), 
                     
                    alpha = .8,
                   size = 5, 
                   fontface = "bold")

layer_data(i = 2)
```


```{r}
compute_post <- function(data){
  
  data %>% 
    mutate(xend = x,
           yend = 0)
  
}

geom_post <- function(...){
  stat_group(compute_fun = compute_post, geom = "segment", ...)
}


data.frame(outcome = 0:1, prob = c(.4, .6)) |>
  ggplot() + 
  aes(x = outcome,
      y = prob) + 
  geom_post() + 
  geom_point() + 
  labs(title = "probability by outcome")


```


```{r}
compute_xmean <- function(data){
  
  data %>% 
    summarize(xintercept = mean(x))
  
}

geom_xmean <- function(...){
  stat_group(compute_fun = compute_xmean, geom = "vline", 
             dropped_aes = c("x", "y"), ...)
}

mtcars |>
  ggplot() + 
  aes(x = wt,
      y = mpg) + 
  geom_point() + 
  geom_xmean(linetype = "dashed")

last_plot() + 
  aes(color = factor(cyl))
```

```{r}
compute_xy_quantile <- function(data, q = .25){
  
  data %>% 
    summarise(x = quantile(x, q),
              y = quantile(y, q)) 
  
}

geom_quantile <- function(...){
  stat_group(compute_fun = compute_xy_quantile, geom = "point", ...)
}

mtcars |>
  ggplot() +
  aes(x = wt,
      y = mpg) +
  geom_point() +
  geom_quantile(size = 8) +
  geom_quantile(size = 8, q = .75)

```





## Stat Rasa one-liners

Since were organize with variable function input in first position and geom in secton position, and we can do one-liners (or two) use positioning for arguments.

`geom_xmean_line()` in 137 characters

```{r}
library(tidyverse)
geom_xmean_line <- function(...){stat_group(function(df) df |> summarize(xintercept = mean(x)), "vline", dropped_aes = c("x", "y"), ...)}


ggplot(cars) +
  aes(speed, dist) + 
  geom_point() + 
  geom_xmean_line(linetype = 'dashed')
  
last_plot() + 
  aes(color = dist > 50)


geom_xmean <- function(...){stat_group(function(df) df |> summarize(x = mean(x), y = I(.05)), "point", ...)}


ggplot(cars) +
  aes(speed, dist) + 
  geom_point() + 
  geom_xmean(size = 8, shape = "diamond") 


```


`geom_post()` in 101 characters

```{r cars}
geom_post <- function(...){stat_group(function(df) df |> mutate(xend = x, yend = 0), "segment", ...)}

data.frame(prob = c(.4,.6), outcome = c(0, 1)) %>% 
ggplot(data = .) +
  aes(outcome, prob) + 
  geom_post() +
  geom_point() 

geom_expectedvalue <- function(...){stat_group(function(df) df |> summarise(x = sum(x*y), y = 0), "point", ...)} 

last_plot() + 
  geom_expectedvalue()

geom_expectedvalue_label <- function(...){stat_group(function(df) df |> summarise(x = sum(x*y), y = 0) |> mutate(label = round(x, 2)), "text", hjust = 1, vjust = 0, ...)} 

last_plot() +
  geom_expectedvalue_label()
```

### `geom_proportion()` and  `geom_proportion_label()`


```{r}
rep(1, 15) |> 
  c(0) %>% 
  data.frame(outcome = .) |>
  ggplot() + 
  aes(x = outcome) + 
  geom_dotplot()

geom_proportion <- function(...){stat_panel(function(df) df |> summarise(x = sum(x)/length(x), y = 0), "point", ...)}   # this should work for T/F too when rasa_p is in play

last_plot() + 
  geom_proportion()

geom_proportion_label <- function(...){stat_panel(function(df) df |> summarise(x = sum(x)/length(x), y = 0) |> mutate(label = round(x,2)), vjust = 0, "text", ...)}   # this should work for T/F too when rasa_p is in play

last_plot() + 
  geom_proportion_label()

# last_plot() + 
#   geom_proportion_label() + 
#   ggsample::facet_bootstrap()
# 
# layer_data(i = 2)
# 
# 
# 
# rep(0:1, 10000) %>% # very large 50/50 sample
#   data.frame(outcome = .) |>
#   ggplot() + 
#   aes(x = outcome) + 
#   geom_dotplot() +
#   geom_proportion() + 
#   geom_proportion_label() + 
#   ggsample::facet_sample(n_facets = 25,
#     n_sampled = 16) ->
# p; p 
#   
# 
# layer_data(p, i = 2) |>
#   ggplot() + 
#   aes(x = x) + 
#   geom_rug() + 
#   geom_histogram()


```


## `geom_means` in 131 characters

```{r}
geom_means <- function(...){stat_group(function(df) df |> summarize(x = mean(x, na.rm = T), y = mean(y, na.rm = T)), "point", ...)}

palmerpenguins::penguins %>% 
  ggplot() + 
  aes(bill_length_mm, bill_depth_mm) + 
  geom_point() + 
  geom_means(size = 7)

last_plot() +
  geom_means(size = 5, aes(color = bill_depth_mm > 17))
```
## `geom_grouplabel_at_means()`

```{r}
geom_grouplabel_at_means <-  function(...){stat_group(function(df) df |> group_by(label) |> summarize(x = mean(x, na.rm = T), y = mean(y, na.rm = T)), "label", ...)}

palmerpenguins::penguins %>% 
  ggplot() + 
  aes(bill_length_mm, bill_depth_mm, label = species, group = species) + 
  geom_point() + 
  geom_grouplabel_at_means(size = 7)
```

## `compute_panel_highlight_lines()`

```{r}
compute_panel_highlight_lines <- function(data, which_id = NULL){

data %>% 
  mutate(ind_id = id %in% which_id) %>% 
  arrange(ind_id) %>%
  mutate(group = fct_inorder(id))
  
}

geom_line_highlight <- function(...){stat_panel(compute_panel_fun = compute_panel_highlight_lines, 
             geom = "line", 
             default_aes = aes(color = after_stat(ind_id)),
             ...)
}

gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line_highlight(which_id = "Bolivia",
                      linewidth = 3)

```

# Attempting a unified approach - allowing compute_panel or compute_group to be defined.  

## then we'll write helpers stat_group and stat_panel from this larger function


```{r}
knitr::knit_exit()

```


```{r}
StatRasa <- 
  ggplot2::ggproto("StatRasa", 
                   ggplot2::Stat)

# stat function used in ggplot - we reorder from conventional
stat_rasa <- function(mapping = NULL, 
                      data = NULL,
                      compute_panel_fun = NULL, 
                      compute_fun = NULL, 
                      geom = "point", 
                      position = "identity",
                      required_aes = NULL,
                      default_aes = NULL,
                      dropped_aes = NULL,
                      ...,
                      show.legend = NA,
                      inherit.aes = TRUE) {
   
   if(!is.null(required_aes)){StatRasa$required_aes <- required_aes}
   if(!is.null(default_aes)){StatRasa$default_aes <- default_aes}
   if(!is.null(dropped_aes)){StatRasa$dropped_aes <- dropped_aes}

   if(!is.null(compute_fun)){StatRasa$compute_group <- compute_group_rasa}
   if(!is.null(compute_panel_fun)){StatRasa$compute_panel <- compute_panel_rasa}
   # Pass dotted arguments to a list
   fun.args <- match.call(expand.dots = FALSE)$`...`
   
   ggplot2::layer(
      data = data,
      mapping = mapping,
      stat = StatRasa,
      geom = geom,
      position = position,
      show.legend = show.legend,
      inherit.aes = inherit.aes,
      check.aes = FALSE,
      check.param = FALSE,
      params = list(
         compute_panel_fun = compute_panel_fun, 
         compute_fun = compute_fun,
         fun.args = fun.args,
         na.rm = FALSE,
         ...
      )
   )
}


group_means <- function(data){
  
 data %>% 
    summarise(x = mean(x),
              y = mean(y))
  
}

geom_means <- function(...){
  stat_rasa(compute_fun = group_means, ...)
}

mtcars |>
  ggplot() + 
  aes(x = wt,
      y = mpg) + 
  geom_point() + 
  geom_means(size = 6)

last_plot() + 
  aes(color = factor(cyl))
```



## And a sf targetted one; uses GeomSF as default and the list2. In


```{r, eval = F}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

geo_reference_northcarolina_county <- nc |>
  dplyr::select(county_name = NAME, fips = FIPS) |>
  sf2stat:::sf_df_prep_for_stat(id_col_name = "county_name")

compute_panel_sf_fun <- function(df, ref_data){
  
  data |> dplyr::left_join(ref_data)
  
}

nc %>% 
  sf::st_drop_geometry() %>% 
  select(fips = fips) %>% 
  compute_panel_sf_fun(ref_data = geo_reference_northcarolina_county)
```

```{r, eval = F}
StatSfrasa <- ggplot2::ggproto(`_class` = "StatSfrasa",
                               `_inherit` = ggplot2::Stat)

stat_region_rasa <- function(mapping = NULL, 
                       data = NULL,
                       ref_data = NULL, 
                       compute_panel_sf_fun = NULL, 
                       compute_group_fun = NULL, 

                      required_aes = NULL,
                      default_aes = NULL,
                      dropped_aes = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatSfrasa,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
    coord_sf(crs = crs,
            default_crs = sf::st_crs(crs),
            datum = crs,
            default = TRUE)
     )
  }
```



## Closing remarks, Other Relevant Work, Caveats


See Elio's blog post for details
