---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts

The Grammar of Graphics is a book by Leland Wilkinson that envisions data visualization systems (2000).   The book identifies independent components of data visualizations and advocates for systems that allows for relatively more independent declaration of components.  It describes how graphicial components can be combine to specify familiar or new graph types. 

ggplot2 in the R programming language (2010) is a notable implementation of the ideas elaborated in Wilkinson's book , created by Hadley Wickham, allows plot components to be written down in code.  




the orthogonalizes data visualization components - implementations of which lead to a powerful


 and further developed with co-authors including notable contributions from Winston Chang, Thomas Lin Peterson and Teun van Den Brand, 


## Status Quo

```{r}
library(tidyverse)

```

## Experiment

```{r}

```



## Closing remarks, Other Relevant Work, Caveats
