---
title: "ggedgelist proposal"
author: "Evangeline Reynolds"
date: "9/20/2023"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Intro {ggedgelist} idea

Here we think about a shortcut to a first looks at networks using a flat, edgelist input that leads straight to a ggplot2 plot (ggraph) space; this will uses tidygraph and ggraph under the hood...



## Step 00 Before getting into it, create an 'interesting' edge list

```{r cars, message=F, warning=F}
library(tidyverse)
library(tidygraph)
library(ggraph)

set.seed(12345)

edge_list <- 
  data.frame(node_to = sample(rep(LETTERS[1:10], 50), 
                              replace = T), 
             node_from = sample(rep(LETTERS[1:10], 50),
                                replace = T)) %>% 
  arrange(node_to, node_from) %>% 
  count(node_to, node_from) %>% 
  sample_n(18)

head(edge_list)
```

## Step 0. Whats the status quo edgelist -> network viz

```{r}
edge_list %>% 
  as_tbl_graph() %>% 
  ggraph() +
  geom_edge_link(color = "orange") +
  geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) + 
  geom_node_text(aes(label = name))

# aesthetic mapping for edge characteristics
last_plot() + 
  geom_edge_link(color = "red", 
                 alpha = .5 ,
                 aes(edge_width = n))
```

## Step 0.a And to also vizualize node atributes


```{r}
node_info <- data.frame(my_nodes = LETTERS[1:10], ind_child = sample(c(T,F), 10, replace = T))

edge_list %>% 
  as_tbl_graph() %>%
  left_join(node_info %>% 
              rename(name = my_nodes)) %>%
  ggraph() +
  geom_edge_link(color = "orange") +
  geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) + 
  geom_node_text(aes(label = name))

```



# Proposed functions, direct edgelist -> plotspace feel


```{r}
# get into ggplot2 plot space from edge list data frame 
ggedgelist <- function(edgelist, nodelist = NULL, ...)(
  
  # message("'name' a variable created in the 'nodes' dataframe")
  
    if(is.null(nodelist)){
    edgelist %>% 
    as_tbl_graph() %>% 
    ggraph(...) 
    
  }else{ # join on nodes attributes if they are available
    
    names(nodelist)[1] <- "name"
    
    edgelist %>% 
    as_tbl_graph() %>%
    left_join(nodelist) %>% 
    ggraph(...) 
    
  }
  
)

# get a fill viz w edgelist dataframe only
quick_ggedgelist <- function(edgelist, nodelist = NULL, include_names = F,  ...){
  

  p <- ggedgelist(edgelist = edgelist,
                  nodelist = nodelist, ...) +
  geom_edge_link(color = "orange") +
  geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) 
  
  if(include_names){p + geom_node_label(aes(label = name))}else{p}
  
}



geom_node_label_auto <- function(...){ 
  
  geom_node_label(aes(label = name), ...)
  
}

geom_node_text_auto <- function(...){ 
  
  geom_node_text(aes(label = name), ...)
  
}


```


# examples w/ proposed functions


## `quick_ggedgelist()`

```{r, out.width="33%", fig.show='hold', message=F}
head(edge_list)
head(node_info)
edge_list %>% 
  quick_ggedgelist() 

edge_list %>% 
  quick_ggedgelist(include_names = T) 

edge_list %>% 
  quick_ggedgelist(nodelist = node_info) + 
  geom_node_point(aes(color = ind_child), size = 10)


```

## `ggedgelist()` + `geom_node_label_auto()`


```{r}
edge_list %>% 
  ggedgelist(layout = "kk") + 
  geom_edge_link(arrow = arrow(), linetype = "dashed") +
  geom_node_tile(width = .18, height = .5) + 
  geom_node_label_auto() 
```

# using the ggflowchart example (100% inspiration)

```{r}
# flowcharter example
ggflowchart_example <- tribble(~from, ~to,
        "A", "B",
        "A", "C",
        "A", "D",
        "B", "E",
        "C", "F",
        "F", "G") 

ggflowchart_example %>% 
  as_tbl_graph()
```


## Start by using quick plot function 'qedgelist2ggraph'

```{r, out.width="33%", fig.show='hold', message=F}
ggflowchart_example %>% 
  quick_ggedgelist(layout = "stress", 
                   include_names = T)

ggflowchart_example %>% 
  quick_ggedgelist(layout = "tree", 
                   include_names = T)

# auto which is default also produces three in this case
ggflowchart_example %>% 
  quick_ggedgelist(layout = "auto") 

layer_data(last_plot(), i = 2)
```

## Use ggedgelist and geom_edge_* and geom_node_* functions to customize

```{r}
ggflowchart_example %>% 
  ggedgelist() +
  geom_edge_link(linetype = "dashed") + 
  geom_node_point(size = 12, alpha = .2) + 
  geom_node_label_auto(fill = "magenta")

```

## Capabilities end here.  If you need to access powerful network calculation capabilities, step back into the tidygraph world!  


```{r}
ggflowchart_example %>% 
  as_tbl_graph() %>% 
  mutate(dg_cent = centrality_degree()) %>% 
  ggraph("stress") +
  geom_edge_link(linetype = "dashed") + 
  geom_node_point(alpha = .2, aes(size = dg_cent)) + 
  scale_size(range = c(8, 15)) +
  geom_node_label_auto(fill = "magenta")

```


## use ggflowchart to make it even faster and prettier

allows for node info, which currently isn't in proposal...

```{r}
library(ggflowchart)
node_data <- tibble::tibble(
  name = c("A", "B", "C", "D", "E", "F", "G"),
  type = c("Type 1", "Type 1", "Type 1", "Type 1", 
           "Type 2", "Type 2", "Type 2")
  )
ggflowchart(ggflowchart_example, 
            node_data, fill = type)

```

