---
title: "ggsample and the CLT"
subtitle: "Using flipbookr and xaringan"
author: "Me"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = F, warning = FALSE, comment = "", cache = F)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping 
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 70%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```



---




```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
```



---

`r flipbookr::chunk_reveal("faithful1", left_assign = "detect")`



```{r faithful1, warning=FALSE, include = FALSE}
# population mean
ggplot(data = faithful) + 
  aes(x = waiting) + 
  geom_rug() + 
  geom_dotplot(dotsize = .5) + 
  ggxmean::geom_x_mean() + 
  ggxmean::geom_x_mean_label() 
```



---

`r flipbookr::chunk_reveal("single_sample", break_type = 10, left_assign = "detect")`



```{r single_sample, include = FALSE}
# A single sample mean
last_plot() + 
  scale_x_continuous(limits = c(40,100)) +
  ggsample::facet_sample(n_facets = 1, n_sampled = 20)
```



---

`r flipbookr::chunk_reveal("bunch_of_samples", left_assign = "detect")`



```{r bunch_of_samples, include = FALSE}
# A bunch of sample means
last_plot() + 
  ggsample::facet_sample(n_facets = 20, n_sampled = 20)
```



---

`r flipbookr::chunk_reveal("make_a_function", left_assign = "detect")`



```{r make_a_function, include = FALSE}
# How can we take many samples
sample_wait_times <- function(x){
  
  sample(x = faithful$waiting, size = 20)
  
}

# check that function is working
sample_wait_times()
```



---

`r flipbookr::chunk_reveal("normal_dist", left_assign = "detect")`



```{r normal_dist, include = FALSE}
set.seed(12345)
tibble(x = 1:10000) %>% # for 1000 trials
  mutate(sample = map(x, sample_wait_times)) %>% # sample
  mutate(sample_mean = map_dbl(sample, mean)) %>% # and compute means
  ggplot() + 
  aes(x = sample_mean) + 
  geom_rug() + 
  geom_histogram() + 
  ggxmean::geom_normal_dist(fill = "magenta",
                            height = 7000) + 
  ggxmean::geom_normal_dist_zlines(color = "magenta",
                            height = 7000) +
  ggxmean::geom_x_mean() + 
  ggxmean::geom_x_mean_label() 
```


---

`r flipbookr::chunk_reveal("faithfulsamples", left_assign = "detect")`



```{r faithfulsamples, include = FALSE}
# And so 
set.seed(123456)
ggplot(data = faithful) + 
  aes(x = waiting) + 
  geom_rug() +
  geom_vline(xintercept = rep(71, 20), color = "red") +
  geom_dotplot(dotsize = .5) + 
  ggxmean::geom_x_mean(linetype = "dashed") + 
  ggxmean::geom_x_mean_label() +
  ggsample::facet_sample(n_facets = 16, 
                         n_sampled = 20) + 
  ggxmean:::geom_tdist(height = 7, fill = "goldenrod") + 
  ggxmean:::geom_ttestconf(color = "darkgreen",
                           alpha = 1)
```


<!-- adjust font size in this css code chunk, currently 80 -->

---




```{css, eval = TRUE, echo = FALSE, include = FALSE}
.remark-code{line-height: 1.5; font-size: 100%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```




