---
title: "Exploded code"
subtitle: "Using flipbookr and xaringan"
author: "Me"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = F, warning = FALSE, comment = "", cache = F)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```



---


`r flipbookr::chunk_reveal("prep", left_assign = "detect")`


```{r prep, include = FALSE}
library(tidyverse)

Titanic %>% 
  data.frame() %>% 
  uncount(weights = Freq) ->
tidy_titanic ; tidy_titanic %>% head()

Titanic %>% 
  data.frame() ->
flat_titanic ; flat_titanic %>% head()
```


---


`r flipbookr::chunk_reveal("slow", left_assign = "detect")`


```{r slow, include = FALSE} 
tidy_titanic %>% 
  group_by(Sex, Survived) %>% 
  summarize(value = n()) %>% 
  pivot_wider(names_from = Survived)
```

---


`r flipbookr::chunk_reveal("count", left_assign = "detect")`


```{r count, include = FALSE}
pivot_count <- function(data, y, y0, y00, x = NULL, 
                          value = NULL, 
                          fun = sum, 
                          pivot = T #ifelse(is.null(x),F,T)
                          ){
  
  y00 <- enquo(y00)
  y0 <- enquo(y0)
  y <- enquo(y)
  x <- enquo(x)

    data <- data %>% mutate(value = 1)

  
tidy <- data %>% 
  group_by(!!y00, !!y0, !!y, !!x) %>% 
  summarize(value = fun(value)) %>% 
  ungroup()
  
  if(pivot){
  tidy %>% 
    pivot_wider(names_from = !!x)
  }else{
  tidy
  }

}
```


---


`r flipbookr::chunk_reveal("calc", left_assign = "detect")`


```{r calc, include = FALSE}
pivot_calc <- function(data, y, y0, y00, x = NULL, 
                          value, 
                          fun = sum, 
                          pivot = T #ifelse(is.null(x),F,T)
                          ){
  
  y00 <- enquo(y00)
  y0 <- enquo(y0)
  y <- enquo(y)
  x <- enquo(x)
  value <- enquo(value)
  
tidy <- data %>% 
  group_by(!!y00, !!y0, !!y, !!x) %>% 
  summarize(value = fun(!!value)) %>% 
  ungroup()
  
  if(pivot){
  tidy %>% 
    pivot_wider(names_from = !!x)
  }else{
  tidy
  }

}
```


---


`r flipbookr::chunk_reveal("examples", left_assign = "detect")`


```{r examples, include = FALSE}
tidy_titanic %>% 
   pivot_count(x = Sex)

tidy_titanic %>% 
  pivot_count(y = Survived, x = Sex) 

tidy_titanic %>% 
  pivot_count(y = Survived, x = Sex, y0 = Class)

tidy_titanic %>% 
  pivot_count(x = Sex, y = Survived, pivot = F)
```



---


`r flipbookr::chunk_reveal("calc2", left_assign = "detect")`


```{r calc2, include = FALSE}
flat_titanic %>%
  pivot_calc(x = Sex, y = Survived, 
             value = Freq, fun = sum)
```

---


`r flipbookr::chunk_reveal("style", left_assign = "detect")`


```{r style, include = FALSE}
tidy_titanic %>% 
  pivot_count(x = Sex, y = Survived, y0 = Class) %>% 
  group_by(Class) %>% 
  gt::gt()
```


---


`r flipbookr::chunk_reveal("style2", left_assign = "detect")`


```{r style2, include = FALSE}
tidy_titanic %>% 
  pivot_count(x = Sex, y = Survived, y0 = Class, y00 = Age) %>% 
  group_by(Age) %>% 
  gt::gt()
```


---


`r flipbookr::chunk_reveal("proportionslow", left_assign = "detect")`


```{r proportionslow, include = FALSE}
tidy_titanic %>% 
  group_by(Sex, Survived) %>% 
  summarize(value = n()) %>% 
  group_by(Sex) %>% 
  mutate(prop = value/sum(value)) %>% 
  select(-value) %>% 
  pivot_wider(values_from = prop, names_from = Sex)
```

---


`r flipbookr::chunk_reveal("propfunction", left_assign = "detect")`


```{r propfunction, include = FALSE}
two_way_prop <- function(data, y, y0, y00, x = NULL, value = NULL, fun = sum, within, pivot = !is.null(x)){
  
  y00 <- enquo(y00)
  y0 <- enquo(y0)
  x <- enquo(x)
  y <- enquo(y)
  within <- enquo(within)
  
  if(is.null(value)){
    data <- data %>% mutate(value = 1)
  }else{
    value <- enquo(value)
  }
  
  data %>% 
  group_by(!!y00, !!y0, !!y, !!x) %>% 
  summarize(value = fun(value)) %>% 
  group_by(!!within) %>% 
  mutate(prop = value/sum(value)) %>% 
  select(-value) %>%
  ungroup() ->
tidy
  
  if(pivot){
    
    tidy %>% 
      pivot_wider(values_from = prop, names_from = !!x) 
    
  }else{
    
    tidy
  }
  
}
```


---


`r flipbookr::chunk_reveal("propusage", left_assign = "detect")`


```{r propusage, include = FALSE}
tidy_titanic %>% 
  two_way_prop(y = Survived, x = Class, within = Class)

tidy_titanic %>% 
  two_way_prop(y0 = Survived, 
               y = Sex, 
               x = Class, 
               within = Survived)
```

---


`r flipbookr::chunk_reveal("propgraph", left_assign = "detect")`


```{r propgraph, include = FALSE}
tidy_titanic %>% 
  two_way_prop(y0 = Survived, 
               y = Sex, 
               x = Class, 
               within = Survived, pivot = F) %>% 
  ggplot() +
  aes(x = Class, y = Sex) +
  facet_grid(rows = vars(Survived)) +
  geom_tile() +
  aes(fill = prop) + 
  geom_text(aes(label = prop %>% round(3)))
```


