---
title: My First Closeread
format: closeread-html
---

Graphical poems with ggplot2.


```{r,include=F}
library(tidyverse)
cars <- tibble(cars)
knitr::opts_chunk$set(echo = T, message = F, warning = F)

```

:::{.cr-section}

Closeread enables scrollytelling.

ggplot2 allows you build up your plot bit by bit - to write 'graphical poems' (Wickham 2010).  It is easy to gain insights simply by defining what variation i variables in dataset via channels in data,  @cr-features

:::{#cr-features}
1. Highlighting
2. Zooming
3. Panning
:::


:::{focus-on="cr-myplot000"}
Now we add a point layer, ``r knitr::knit_code$get("myplot000") |> paste(collapse = "")``
:::

:::{#cr-myplot000}
```{r myplot000, echo = T}
library(ggplot2); 
cars
```
:::



:::{focus-on="cr-myplot00"}
Now we add a point layer, ``r knitr::knit_code$get("myplot00") |> paste(collapse = "")``
:::

:::{#cr-myplot00}
```{r myplot00, echo = T}
cars |>
ggplot(data = _) 
```
:::


:::{focus-on="cr-myplot0"}
Now we define positional mapping x and y, ``r knitr::knit_code$get("myplot0") |> paste(collapse = " ")``
:::

:::{#cr-myplot0}
```{r myplot0}
last_plot() + 
  aes(x = speed, y = dist)
```
:::



:::{focus-on="cr-myplot1"}
Now we add a point layer, ``r knitr::knit_code$get("myplot1") |> paste(collapse = " ")``
:::

:::{#cr-myplot1}
```{r myplot1}
last_plot() + 
  geom_point()
```
:::

:::{focus-on="cr-myplot2"}
Now we add a title, ``r knitr::knit_code$get("myplot2") |> paste(collapse = " ")``
:::

:::{#cr-myplot2}
```{r myplot2}
last_plot() + 
  labs(title = "A correlation study: speed to stopping distance")
```
:::


:::{focus-on="cr-myplot3"}
and smooth, ``r knitr::knit_code$get("myplot3") |> paste(collapse = " ")``
:::

:::{#cr-myplot3}
```{r myplot3}
last_plot() + 
  geom_smooth()
```
:::



:::{focus-on="cr-myplot4"}
And plot fitted, ``r knitr::knit_code$get("myplot4") |> paste(collapse = " ")``
:::

:::{#cr-myplot4}
```{r myplot4}
last_plot() + 
  stat_smooth(geom = GeomPoint, xseq = cars$speed, color = "blue")
```
:::






:::{focus-on="cr-myplot5"}
With subset of data, ``r knitr::knit_code$get("myplot5") |> paste(collapse = " ")``
:::

:::{#cr-myplot5}
```{r myplot5}
last_plot() %+%
  (cars |> filter(speed < 15))
```
:::




:::{focus-on="cr-myplot6"}
change to teaching theme, ``r knitr::knit_code$get("myplot6") |> paste(collapse = " ")``
:::

:::{#cr-myplot6}
```{r myplot6}
last_plot() +
  ggchalkboard::theme_chalkboard()
```
:::




:::







