---
title: "Easy Geom Recipes (Exploded version)"
subtitle: "Using flipbookr and xaringan"
author: "Gina Reynold, Morgan Brown"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = F, warning = FALSE, comment = "", cache = F)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 70%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```




Using ggplot2 has been described as writing 'graphical poems'.
<!-- and a system that lets us 'speak our plot into existence'.  -->
But we may feel at a loss for 'words' when functions we'd like to have don't exist.  The ggplot2 extension system allows us to build new 'vocabulary' for fluent expression.  

An exciting extension mechanism is inheriting from existing geoms.  Particularly important to statisticians and mathematicians is writing new geom_* that perform and visualize calculations.  

To get your feet wet in this world and give you a taste of patterns for geom extension, we provide three basic examples of the geoms_ that inherit from *existing* geoms (point, text, segment, etc) along with practice exercises.  With such geoms, calculation is done under the hood by the ggplot2 system; with these geom, you can write new graphical poems with exciting new graphical 'words'! 

This tutorial is intended for individuals who already have a working knowledge of the grammar of ggplot2, but may like to build a richer vocabulary for themselves.  

---


`r flipbookr::chunk_reveal("setup", left_assign = "detect")`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

---
class: inverse, middle, center

# Recipe #1: `geom_point_xy_medians()`

--

- This will be a point at the median of x and y

---
class: inverse, middle, center

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("penguins", left_assign = "detect")`


```{r penguins, include = FALSE}
library(tidyverse)
library(palmerpenguins)
penguins <- remove_missing(penguins)


penguins_medians <- penguins %>% 
  summarize(bill_length_mm_median = median(bill_length_mm),
            bill_depth_mm_median = median(bill_depth_mm))

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  geom_point(data = penguins_medians,
    color = "red", size = 4,
             aes(x = bill_depth_mm_median,
                 y = bill_length_mm_median))
```

---
class: inverse, middle, center

## Step 1: computation

--

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

---


`r flipbookr::chunk_reveal("compute_group_xy_medians", left_assign = "detect")`


```{r compute_group_xy_medians, include = FALSE}
compute_group_xy_medians <- function(data, scales){
  data %>% 
    summarize(x = median(x),
            y = median(y))
}
penguins %>%
  rename(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_xy_medians()
```


---
class: inverse, middle, center

## Step 2: define ggproto

--

- what's the naming convention for the proto object?
- which aesthetics are required as inputs?
- where does the function from above go?

---


`r flipbookr::chunk_reveal("StatXYMedians", left_assign = "detect")`


```{r StatXYMedians, include = FALSE}
StatXYMedians <- ggplot2::ggproto(`_class` = "StatXYMedians",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_xy_medians)
```

---
class: inverse, middle, center

## Step 3: define geom_* function


--

- define the stat and geom for your layer

---


`r flipbookr::chunk_reveal("geom_point_xy_medians", left_assign = "detect")`


```{r geom_point_xy_medians, include = FALSE} 
geom_point_xy_medians <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatXYMedians, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

---
class: inverse, middle, center

## Step 4: Enjoy! Use your function

---


`r flipbookr::chunk_reveal("enjoy_penguins", left_assign = "detect")`


```{r enjoy_penguins, include = FALSE}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, y = bill_length_mm)+
  geom_point()+
  geom_point_xy_medians(color = "red")
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_penguins", left_assign = "detect")`


```{r conditional_penguins, include = FALSE}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, 
      y = bill_length_mm, 
      color = species)+
  geom_point()+
  geom_point_xy_medians(size = 4)
```

## Now you ...

### Create the function `geom_point_xy_means()`.  

---
class: inverse, middle, center

# Recipe #2:  `geom_label_id()`

---
class: inverse, middle, center

## Step 1: computation



---


`r flipbookr::chunk_reveal("compute_group_row_number", left_assign = "detect")`


```{r compute_group_row_number, include = FALSE}
# you won't use the scales argument, but ggplot will later
compute_group_row_number <- function(data, scales){
  
  data %>% 
    # add an additional column called label
    # the geom we inherit from requires the label aesthetic
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs as columns
  rename(x = speed, y = dist) %>% 
  compute_group_row_number() %>% 
  head()
```

---
class: inverse, middle, center

## Step 2: define ggproto


---


`r flipbookr::chunk_reveal("StatRownumber", left_assign = "detect")`


```{r StatRownumber, include = FALSE}
StatRownumber <- ggplot2::ggproto(`_class` = "StatRownumber",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_row_number)
```

---
class: inverse, middle, center

## Step 3: define geom_* function

---


`r flipbookr::chunk_reveal("geom_label_row_number", left_assign = "detect")`


```{r geom_label_row_number, include = FALSE}
geom_label_row_number <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatRownumber, # proto object from Step 2
    geom = ggplot2::GeomLabel, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


---
class: inverse, middle, center

## Step 4: Enjoy! Use your function

---


`r flipbookr::chunk_reveal("enjoy_again", left_assign = "detect")`


```{r enjoy_again, include = FALSE}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label_row_number(hjust = 1.2) # function in action
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_compute", left_assign = "detect")`


```{r conditional_compute, include = FALSE}
last_plot() + 
  aes(color = dist > 60) # Computation is within group
```




---

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


Hint:



---

# Recipe #3:  `geom_point_lm_fitted()`

---

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("fitted_1", left_assign = "detect")`


```{r fitted_1, include = FALSE}
model <- lm(formula = bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins_w_fitted <- penguins %>% 
  mutate(fitted = model$fitted.values)


penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_point(data = penguins_w_fitted,
             aes(y = fitted),
             color = "blue")
```

---
class: inverse, middle, center

## Step 1: computation

---


`r flipbookr::chunk_reveal("fitted_2", left_assign = "detect")`


```{r fitted_2, include = FALSE}
compute_group_lm_fitted<- function(data, scales){
  model<-lm(formula= y ~ x, data = data)
  data %>% 
    mutate(y=model$fitted.values)
}

# test out the function
penguins %>%
  # rename to explicitly state the x and y inputs
  rename(x = bill_depth_mm, y = bill_length_mm)%>%
  compute_group_lm_fitted()
```


## Step 2: define ggproto

---


`r flipbookr::chunk_reveal("fitted_3", left_assign = "detect")`


```{r fitted_3, include = FALSE}
StatLmFitted<-ggplot2::ggproto(`_class` = "StatLmFitted",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_lm_fitted)
```


## Step 3: define geom_* function


---


`r flipbookr::chunk_reveal("fitted_4", left_assign = "detect")`


```{r fitted_4, include = FALSE}
geom_point_lm_fitted <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatLmFitted, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

## Step 4: Enjoy! Use your function

---


`r flipbookr::chunk_reveal("fitted_5", left_assign = "detect")`


```{r fitted_5, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F)+
  geom_point_lm_fitted(color="blue")
```

### And check out conditionality

---


`r flipbookr::chunk_reveal("fitted_6", left_assign = "detect")`


```{r fitted_6, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F) +
  geom_point_lm_fitted() + 
  facet_wrap(facets = vars(species))
```

