---
title: "Easy Geom Recipes (Exploded version)"
subtitle: "Using flipbookr and xaringan"
author: "Gina Reynold, Morgan Brown"
output:
  xaringan::moon_reader:
    lib_dir: libs
    seal: false
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


class: inverse, middle
background-image: url(https://images.unsplash.com/photo-1592173376801-185310a68dea?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=1286&q=80)
background-size: cover



# Easy geom_*() recipes, 2024

###Gina Reynolds and Morgan Brown

<br>
<br>
<br>

```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 7, fig.height = 5, message = F, warning = FALSE, comment = "", cache = T)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```

---


### Using ggplot2 has been described as writing 'graphical poems'.

--

<!-- and a system that lets us 'speak our plot into existence'.  -->
### But we may feel at a loss for words when functions we'd like to have don't exist.  The ggplot2 extension system allows us to build new 'vocabulary' for fluent expression.  

--

### An exciting extension mechanism that allows more plotting fluidity is to create new, custom 'Stats' and combine them with existing 'Geoms' (GeomPoint, GeomText, GeomSegment, etc)into new geom\_\* or stat\_\* functions.  

---

### By the way, you might only be familiar with the user-facing geom_point(), geom_text(), geom_segment() layer functions that use the more primitive GeomPoint, GeomText, and GeomSegment under the hood, and that's okay.  You might also have a feel for existing stats like StatCount and StatBin, specify computation in user-facing functions like geom_bar() and geom_histogram(). 

---

### To get your feet wet in this world and give you a taste of patterns for new-Stat-plus-existing-Geom extension, we provide three *introductory* recipes (A, B, and C) for geom\_\* functions that make use of newly defined 'Stats' and *existing* 'Geoms' along with three practice exercises: *Easy geom\_() recipes*

---

### With your new geom\_\*() function, calculation is done under the hood by the ggplot2 system; with these geom\_\*() functions, you can write graphical poems with concise new language you've tailored for goals! 

---

### This tutorial is intended for individuals who already have a working knowledge of the grammar of ggplot2, but may like to build a richer vocabulary for themselves. 

--

### Grab an .Rmd worksheet version [here](https://raw.githubusercontent.com/EvaMaeRey/mytidytuesday/master/2022-01-03-easy-geom-recipes/easy_geom_recipes_2024.Rmd).

### or you can browse the rendered version of that [here](https://evamaerey.github.io/mytidytuesday/2022-01-03-easy-geom-recipes/easy_geom_recipes_2024.html).

---

## *Recipes Form Preview:*


--

## **Step 0. Get the job done with 'base' ggplot2.**  



<!-- ## *It's a good idea to clarify what needs to happen without getting into the extension architecture* -->

--

## **Step 1. Use base build experience to help define compute and test.** 



<!-- ## *Define compute that will transform your input data before rendering it. Try it out.* -->

--

## **Step 2. Use compute definition to define a new Stat and test.** 



<!-- ## *Stats, which are ggproto objects, allow your  to get compute done within a ggplot2 build!  You'll use the computation function from Step 1 to build your Stat. Then, use the stat with the generic ggplot `layer()` function to test it out.  `layer()` lets you combine stats and geoms for very flexible expression and used to create the user-facing layer functions that you're familiar with geom_point, geom_histogram, etc.* -->

--

## **Step 3. Use your new Stat to define a user-facing geom\_\*() function and test.** 



<!-- ## *You're ready to write your target function!  We'll define a user-facing function using `layer()` function and the newly defined stat from Step 2.  Test/Enjoy!* -->




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

---
class: inverse, middle, center

# Recipe A.

--

## `geom_point_xy_medians()`



---
class: inverse, middle, center

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("penguins")`


```{r penguins, include = FALSE}
library(tidyverse)
library(palmerpenguins)

penguins <- remove_missing(penguins)

penguins %>% 
  summarize(bill_length_mm_median = median(bill_length_mm),
            bill_depth_mm_median = median(bill_depth_mm)) ->
penguins_medians

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  geom_point(data = penguins_medians,
    color = "red", size = 4,
             aes(x = bill_depth_mm_median,
                 y = bill_length_mm_median))
```

---
class: inverse, middle, center

## Step 1: computation

--

- define computation that should be done before plotting
- test that functionality

---


`r flipbookr::chunk_reveal("compute_group_xy_medians")`


```{r compute_group_xy_medians, include = FALSE}
compute_group_xy_medians <- function(data, 
                                     scales){
  data %>% 
    summarize(x = median(x),
              y = median(y))
}

# And use computation function to test it
penguins %>%
  # function requires dataframe
  # with columns named x and y
  select(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_xy_medians()
```


---
class: inverse, middle, center

## Step 2: define ggproto

--

# What's the naming convention for the proto object?
# Which aesthetics are required as inputs?
# Where does the function from above go?

---


`r flipbookr::chunk_reveal("StatXymedians", widths = c(1,.2), break_type = 1)`


```{r StatXymedians, include = F}
# define new Stat
StatXymedians <- ggplot2::ggproto(
  `_class` = "StatXymedians",    # declare new Stat
  `_inherit` = ggplot2::Stat,    # inherit behavior of generic stat
  required_aes = c("x", "y"),    # declare required variable mappings
  compute_group = compute_group_xy_medians  # pass computation definition for stat
  )

# test new stat w ggplot2 layer function
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  layer(geom = GeomPoint, 
        stat = StatXymedians, 
        position = "identity", 
        params = list(size = 8)) 
```



---
class: inverse, middle, center

## Step 3: define user-facing geom\_\* function

For ease of use, we'll wrap the layer function into something that looks more familiar.  We also take care to manage arguments like mapping, data, position, na.rm, show.legend, inherit.aes, etc, so that your new, user-facing function will behaive similarly to other user-facing geom_*s.

- define the stat and geom for your layer

---


`r flipbookr::chunk_reveal("geom_point_xy_medians",  break_type = 1)`


```{r geom_point_xy_medians, include = FALSE} 
geom_point_xy_medians <- function(
  mapping = NULL, # global aesthetics will be used if NULL
  data = NULL, # global data will be used if NULL
  position = "identity", 
  na.rm = FALSE, 
  show.legend = NA,
  inherit.aes = TRUE, 
  ...  ){ # many other arguments can be used specific to  the Geom thats used in layer and to computation definition
ggplot2::layer(
    stat = StatXymedians, 
    geom = ggplot2::GeomPoint, 
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
  
}
```

---
class: inverse, middle, center


## Step 3.b: Enjoy! Use your function


---


`r flipbookr::chunk_reveal("enjoy_penguins")`


```{r enjoy_penguins, include = FALSE}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, 
      y = bill_length_mm) +
  geom_point() +
  geom_point_xy_medians(size = 8)
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_penguins")`


```{r conditional_penguins, include = FALSE}
last_plot() +
  aes(color = species)
```

---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1628559225804-71f6e4643f44?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=837&q=80)
background-size: cover



---
class: inverse, middle, center

## Now you ...

--

### Create the function `geom_point_xy_means()`

---
class: inverse, middle, center

# Recipe B.

--

## `geom_label_id()`


---
class: inverse, center, middle

## Step 0: use base ggplot2 to get the job done

---

`r flipbookr::chunk_reveal("cars")`

```{r cars, include = F}
cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label(aes(label = id_number), 
             hjust = 1.2)
```


---
class: inverse, middle, center

## Step 1: define computation



---


`r flipbookr::chunk_reveal("compute_group_row_number")`


```{r compute_group_row_number, include = FALSE}
# you won't use the scales argument, but ggplot will later
compute_group_row_number <- function(data,
                                     scales){
  
  data %>% 
    # add an additional column called label
    # the geom we inherit from requires the label aesthetic
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs as columns
  select(x = speed, 
         y = dist) %>% 
  compute_group_row_number()
```

---
class: inverse, middle, center

## Step 2: define Stat


---


`r flipbookr::chunk_reveal("StatRownumber")`


```{r StatRownumber, include = FALSE}
StatRownumber <- ggplot2::ggproto(
  `_class` = "StatRownumber",
  `_inherit` = ggplot2::Stat,
  required_aes = c("x", "y"),
  compute_group = compute_group_row_number
  )


## Step 2.b: test
cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  layer(geom = GeomText,
        stat = StatRownumber,
        position = "identity")
```


---
class: inverse, middle, center

## Step 3: define user-facing geom\_\* function

---


`r flipbookr::chunk_reveal("geom_label_row_number", break_type = 1)`


```{r geom_label_row_number, include = FALSE}
geom_label_row_number <- function(
  mapping = NULL, 
  data = NULL,
  position = "identity", 
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  
  ggplot2::layer(
    stat = StatRownumber, # proto object from Step 2
    geom = ggplot2::GeomLabel, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
    
  )
}
```


---
class: inverse, middle, center


## Step 3.b: Enjoy! Use your function


---


`r flipbookr::chunk_reveal("enjoy_again")`


```{r enjoy_again, include = FALSE}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label_row_number(hjust = 1.2) # function in action
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_compute")`


```{r conditional_compute, include = FALSE}
last_plot() + 
  aes(color = dist > 60) 
```


---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1530751127259-074b0cdc0469?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=1413&q=80)
background-size: cover

---
class: inverse, middle, center

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


Hint: `paste0("(", 1, ", ",3., ")")`



---
class: inverse, middle, center

# Recipe C.  

--


## `geom_point_lm_fitted()`

---

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("fitted_1")`


```{r fitted_1, include = FALSE}
model <- lm(formula = 
              bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins_w_fitted <- penguins %>% 
  mutate(fitted = model$fitted.values)


penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_point(data = penguins_w_fitted,
             aes(y = fitted),
             color = "blue")
```

---
class: inverse, middle, center

## Step 1: define computation

---


`r flipbookr::chunk_reveal("fitted_2")`


```{r fitted_2, include = FALSE}
compute_group_lm_fitted <- function(data,
                                   scales){
  model<-lm(formula= y ~ x, data = data)
  data %>% 
    mutate(y=model$fitted.values)
}

# test out the function
penguins %>%
  # select to explicitly state the x and y inputs
  select(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_lm_fitted()
```


## Step 2: define new Stat

---


`r flipbookr::chunk_reveal("fitted_3", break_type = 1)`


```{r fitted_3, include = FALSE}
StatLmFitted <- ggplot2::ggproto(
  `_class` = "StatLmFitted",
  `_inherit` = ggplot2::Stat,
  required_aes = c("x", "y"),
  compute_group = compute_group_lm_fitted
  )
```


## Step 3: define user-facing geom\_\* function


---


`r flipbookr::chunk_reveal("fitted_4", break_type = 1)`


```{r fitted_4, include = FALSE}
geom_point_lm_fitted <- function(
  mapping = NULL, data = NULL,
  position = "identity", 
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  
  ggplot2::layer(
    stat = StatLmFitted, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
  
}
```

---
class: inverse, middle, center


## Step 3.b: Enjoy! Use your function



---


`r flipbookr::chunk_reveal("fitted_5")`


```{r fitted_5, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F)+
  geom_point_lm_fitted(color = "blue")
```

---
class: inverse, middle, center


### And check out conditionality

---


`r flipbookr::chunk_reveal("fitted_6")`


```{r fitted_6, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, 
      y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F) +
  geom_point_lm_fitted() + 
  facet_wrap(facets = vars(species))
```



---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1499636136210-6f4ee915583e?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=464&q=80)
background-size: cover


---
class: inverse, middle, center

## Now you... Create the function `geom_segment_lm_residuals()`. 

--

### Hint: read about what aesthetics are required for segments.  We'll give you Step 0 this time...

--

## Step 0: use base ggplot2 to get the job done

---

`r flipbookr::chunk_reveal("residuals")`

```{r residuals, include=FALSE}
model <- lm(formula = 
              bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins_w_fitted <- penguins %>% 
  mutate(fitted = model$fitted.values)

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_segment(data = penguins_w_fitted,
             aes(yend = fitted, xend = bill_depth_mm),
             color = "blue")
```

---

#### Photo Credits: Thalia Ruiz, Jennifer Pallian, Lisa Hanly, John Dancy, Piper Bchamp

#### More comprehensive extension resources: 

- [ggplot2-extension-cookbook: an example-rich Stat-focused resource](https://evamaerey.github.io/ggplot2-extension-cookbook/) Gina Reynolds

- [Extending your ability to extend ggplot2: StatCircle demo and high level overview](https://www.youtube.com/watch?v=uj7A3i2fi54) Thomas Lin Pederson

- [Extending ggplot2 package Vignette: walkthough many extension  mechanisms](https://cran.r-project.org/web/packages/ggplot2/vignettes/extending-ggplot2.html) 

- [Extending ggplot2 chapter](https://ggplot2-book.org/extensions.html)
