---
title: "easy geom recipes"
author: "Evangeline Reynolds"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Recipe #1:  `geom_label_id()`

---

## Step 0: use base ggplot2 to get the job done


```{r}
library(tidyverse)
cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label(aes(label = id_number), 
             hjust = 1.2)
```

---

## Step 1: computation function

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
compute_group_row_number <- function(data, scales){
  
  data %>% 
    # add an additional column called label
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs 
  rename(x = speed, y = dist) %>% 
  compute_group_row_number() %>% 
  head()
```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs
- where does the function from above go?

```{r}
StatRownumber <- ggplot2::ggproto(`_class` = "StatRownumber",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_row_number)
```


---

## Step 3: define geom_* function

```{r}
geom_label_row_number <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatRownumber, # proto object from step 2
    geom = ggplot2::GeomLabel, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label_row_number(hjust = 1.2) # function in action

last_plot() + 
  aes(color = dist > 60) # Computation is within group
```




---

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


---

# Recipe #2: `geom_point_xy_medians()`

--

- This will be a point at the median of x and y

## Step 0:

```{r}

```


## Step 1:

```{r}

```


## Step 2:

```{r}

```


## Step 3:


```{r}

```

## Now you


### Create the function `geom_point_xy_means()`.  

--

Recipe #2 modification:  `geom_point_xy_means()`

# Recipe #3:  `geom_point_lm_fitted()`

---

## Step 0:

```{r}

```


## Step 1:

```{r}

```


## Step 2:

```{r}

```


## Step 3:


```{r}

```

---

## Now you


### Create the function `geom_segment_lm_residuals()`.  
