---
title: "Easy Geom Recipes (Exploded version)"
subtitle: "Using flipbookr and xaringan"
author: "Gina Reynold, Morgan Brown"
output:
  xaringan::moon_reader:
    lib_dir: libs
    seal: false
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


class: inverse, middle
background-image: url(https://images.unsplash.com/photo-1592173376801-185310a68dea?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=1286&q=80)
background-size: cover



# Easy geom_*() recipes

###Gina Reynolds and Morgan Brown

<br>
<br>
<br>

```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 7, fig.height = 5, message = F, warning = FALSE, comment = "", cache = T)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```

---


### Using ggplot2 has been described as writing 'graphical poems'.

--

<!-- and a system that lets us 'speak our plot into existence'.  -->
### But we may feel at a loss for words when functions we'd like to have don't exist.  The ggplot2 extension system allows us to build new 'vocabulary' for fluent expression.  

--

### An exciting extension mechanism is inheriting from existing Geoms.  Particularly important to statisticians and mathematicians is writing new geom_* functions that perform and visualize calculations.  

---

### To get your feet wet in this world and give you a taste of patterns for geom extension, we provide three *introductory* examples of the geom_ functions that inherit from *existing* geoms (GeomPoint, GeomText, GeomSegment, etc) along with practice exercises.  

--

### (You may only be familiar with the user-facing geom_point(), geom_text(), geom_segment() layer functions that use these more primitive objects under the hood, and that's okay.) 

---

### With your new geom_ function, calculation is done under the hood by the ggplot2 system; with these geom_* functions, you can write graphical poems with consice new language you've tailored for yourself! 

---

### This tutorial is intended for individuals who already have a working knowledge of the grammar of ggplot2, but may like to build a richer vocabulary for themselves. 

--

### Grab an .Rmd worksheet version [here](https://raw.githubusercontent.com/EvaMaeRey/mytidytuesday/master/2022-01-03-easy-geom-recipes/easy_geom_recipes_2024.Rmd).

### or you can browse the rendered version of that [here](https://evamaerey.github.io/mytidytuesday/2022-01-03-easy-geom-recipes/easy_geom_recipes_2024.html).

---

# Preview

Our recipes will take form:

--

**Step 0. Get the job done with 'base' ggplot2.**  
--
It's a good idea to clarify what needs to happen without getting into the extension architecture

--

**Step 1. Write a computation function.** 
--
Wrap the necessary computation into a function that your target geom_*() function will perform.

--

**Step 2. Define a ggproto object.** 
--
ggproto objects allow your extension  to work together with base ggplot2 functions!  You'll use the computation function from step 1 to help define it.

--

**Step 3. Write your geom function!** 
--
You're ready to write your function.  You will incorporate the ggproto from step 2.  

--

**Step 4. Test/Enjoy!** 
--
Take your new geom for a spin! Check out group-wise computation behavior!



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

---
class: inverse, middle, center

# Recipe #1

--

##`geom_point_xy_medians()`

--

- This will be a point at the median of x and y


---
class: inverse, middle, center

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("penguins", left_assign = "detect")`


```{r penguins, include = FALSE}
library(tidyverse)
library(palmerpenguins)

penguins <- remove_missing(penguins)

penguins %>% 
  summarize(bill_length_mm_median = median(bill_length_mm),
            bill_depth_mm_median = median(bill_depth_mm)) ->
penguins_medians

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  geom_point(data = penguins_medians,
    color = "red", size = 4,
             aes(x = bill_depth_mm_median,
                 y = bill_length_mm_median))
```

---
class: inverse, middle, center

## Step 1: computation

--

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

---


`r flipbookr::chunk_reveal("compute_group_xy_medians", left_assign = "detect")`


```{r compute_group_xy_medians, include = FALSE}
compute_group_xy_medians <- function(data, 
                                     scales){
  data %>% 
    summarize(x = median(x),
              y = median(y))
}

penguins %>%
  # function requires dataframe
  # with columns named x and y
  select(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_xy_medians()
```


---
class: inverse, middle, center

## Step 2: define ggproto

--

- what's the naming convention for the proto object?
- which aesthetics are required as inputs?
- where does the function from above go?

---


`r flipbookr::chunk_reveal("StatXymedians", widths = c(1,.2), break_type = 1)`


```{r StatXymedians, include = F}
StatXymedians <- ggplot2::ggproto(
  `_class` = "StatXymedians",    # declare new Stat
  `_inherit` = ggplot2::Stat,    # inherit behavior of generic stat
  required_aes = c("x", "y"),    # declare required variable mappings
  compute_group = compute_group_xy_medians    # pass computation definition for stat
  )
```



---

Try using the stat in long-form, layer function. 

`r flipbookr::chunk_reveal("layerStatXymedians")`


```{r layerStatXymedians, include=F}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  layer(geom = GeomPoint, stat = StatXymedians, 
        position = "identity", params = list(size = 8)) 
```



---
class: inverse, middle, center

## Step 3: define user-facing geom_* function

For ease of use, we'll wrap the layer function into something that looks more familiar.  We also take care to manage arguments like mapping, data, position, na.rm, show.legend, inherit.aes, etc, so that your new, user-facing function will behaive similarly to other user-facing geom_*s.

- define the stat and geom for your layer

---


`r flipbookr::chunk_reveal("geom_point_xy_medians",  break_type = 1)`


```{r geom_point_xy_medians, include = FALSE} 
geom_point_xy_medians <- function(
  mapping = NULL, # global aesthetics will be used if NULL
  data = NULL, # global data will be used if NULL
  position = "identity", 
  na.rm = FALSE, 
  show.legend = NA,
  inherit.aes = TRUE, 
  ...  ){ # many other arguments can be used specific to  the Geom thats used in layer and to computation definition
ggplot2::layer(
    stat = StatXymedians, 
    geom = ggplot2::GeomPoint, 
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
  
}
```

---
class: inverse, middle, center


## Step 4: Enjoy! Use your function


---


`r flipbookr::chunk_reveal("enjoy_penguins", left_assign = "detect")`


```{r enjoy_penguins, include = FALSE}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, 
      y = bill_length_mm) +
  geom_point() +
  geom_point_xy_medians(color = "red")
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_penguins", left_assign = "detect")`


```{r conditional_penguins, include = FALSE}
penguins %>%
  ggplot() +
  aes(x = bill_depth_mm, 
      y = bill_length_mm, 
      color = species) +
  geom_point() +
  geom_point_xy_medians(size = 4)
```

---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1628559225804-71f6e4643f44?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=837&q=80)
background-size: cover



---
class: inverse, middle, center

## Now you ...

--

### Create the function `geom_point_xy_means()`

---
class: inverse, middle, center

# Recipe #2

--

## `geom_label_id()`


---
class: inverse, center, middle

## Step 0: use base ggplot2 to get the job done

---

`r flipbookr::chunk_reveal("cars")`

```{r cars, include = F}
cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label(aes(label = id_number), 
             hjust = 1.2)
```


---
class: inverse, middle, center

## Step 1: computation



---


`r flipbookr::chunk_reveal("compute_group_row_number", left_assign = "detect")`


```{r compute_group_row_number, include = FALSE}
# you won't use the scales argument, but ggplot will later
compute_group_row_number <- function(data,
                                     scales){
  
  data %>% 
    # add an additional column called label
    # the geom we inherit from requires the label aesthetic
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs as columns
  select(x = speed, 
         y = dist) %>% 
  compute_group_row_number()
```

---
class: inverse, middle, center

## Step 2: define ggproto


---


`r flipbookr::chunk_reveal("StatRownumber", break_type = 1)`


```{r StatRownumber, include = FALSE}
StatRownumber <- ggplot2::ggproto(
  `_class` = "StatRownumber",
  `_inherit` = ggplot2::Stat,
  required_aes = c("x", "y"),
  compute_group = compute_group_row_number
  )
```

---
class: inverse, middle, center

## Step 2.b: test



```{r}
cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  layer(geom = GeomText,
        stat = StatRownumber,
        position = "identity")
```


---
class: inverse, middle, center

## Step 3: define user-facing geom_* function

---


`r flipbookr::chunk_reveal("geom_label_row_number", break_type = 1)`


```{r geom_label_row_number, include = FALSE}
geom_label_row_number <- function(
  mapping = NULL, 
  data = NULL,
  position = "identity", 
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  
  ggplot2::layer(
    stat = StatRownumber, # proto object from Step 2
    geom = ggplot2::GeomLabel, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
    
  )
}
```


---
class: inverse, middle, center


## Step 4: Enjoy! Use your function


---


`r flipbookr::chunk_reveal("enjoy_again", left_assign = "detect")`


```{r enjoy_again, include = FALSE}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label_row_number(hjust = 1.2) # function in action
```

---
class: inverse, middle, center

### And check out conditionality!

---


`r flipbookr::chunk_reveal("conditional_compute", left_assign = "detect")`


```{r conditional_compute, include = FALSE}
last_plot() + 
  aes(color = dist > 60) 
```


---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1530751127259-074b0cdc0469?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=1413&q=80)
background-size: cover

---
class: inverse, middle, center

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


Hint: `paste0("(", 1, ", ",3., ")")`



---
class: inverse, middle, center

# Recipe #3:  

--


## `geom_point_lm_fitted()`

---

## Step 0: use base ggplot2 to get the job done

---


`r flipbookr::chunk_reveal("fitted_1")`


```{r fitted_1, include = FALSE}
model <- lm(formula = 
              bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins_w_fitted <- penguins %>% 
  mutate(fitted = model$fitted.values)


penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_point(data = penguins_w_fitted,
             aes(y = fitted),
             color = "blue")
```

---
class: inverse, middle, center

## Step 1: computation

---


`r flipbookr::chunk_reveal("fitted_2")`


```{r fitted_2, include = FALSE}
compute_group_lm_fitted <- function(data,
                                   scales){
  model<-lm(formula= y ~ x, data = data)
  data %>% 
    mutate(y=model$fitted.values)
}

# test out the function
penguins %>%
  # select to explicitly state the x and y inputs
  select(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_lm_fitted()
```


## Step 2: define ggproto

---


`r flipbookr::chunk_reveal("fitted_3", break_type = 1)`


```{r fitted_3, include = FALSE}
StatLmFitted <- ggplot2::ggproto(
  `_class` = "StatLmFitted",
  `_inherit` = ggplot2::Stat,
  required_aes = c("x", "y"),
  compute_group = compute_group_lm_fitted
  )
```


## Step 3: define geom_* function


---


`r flipbookr::chunk_reveal("fitted_4", break_type = 1)`


```{r fitted_4, include = FALSE}
geom_point_lm_fitted <- function(
  mapping = NULL, data = NULL,
  position = "identity", 
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  
  ggplot2::layer(
    stat = StatLmFitted, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
  
}
```

---
class: inverse, middle, center


## Step 4: Enjoy! Use your function



---


`r flipbookr::chunk_reveal("fitted_5", left_assign = "detect")`


```{r fitted_5, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F)+
  geom_point_lm_fitted(color = "blue")
```

---
class: inverse, middle, center


### And check out conditionality

---


`r flipbookr::chunk_reveal("fitted_6", left_assign = "detect")`


```{r fitted_6, include = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, 
      y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F) +
  geom_point_lm_fitted() + 
  facet_wrap(facets = vars(species))
```



---
class: inverse, middle, center
background-image: url(https://images.unsplash.com/photo-1499636136210-6f4ee915583e?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=464&q=80)
background-size: cover


---
class: inverse, middle, center

## Now you... Create the function `geom_segment_lm_residuals()`. 

--

### Hint: read about what aesthetics are required for segments.  We'll give you Step 0 this time...

--

## Step 0: use base ggplot2 to get the job done

---

`r flipbookr::chunk_reveal("residuals")`

```{r residuals, include=FALSE}
model <- lm(formula = 
              bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins_w_fitted <- penguins %>% 
  mutate(fitted = model$fitted.values)

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_segment(data = penguins_w_fitted,
             aes(yend = fitted, xend = bill_depth_mm),
             color = "blue")
```

---
class: inverse, center, middle
background-image: url(https://images.unsplash.com/photo-1630297169944-1316e5e011f8?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxzZWFyY2h8MTc1fHxiYWtlcnl8ZW58MHx8MHx8&auto=format&fit=crop&w=400&q=60)
background-size: cover


## Not interested in writing your own geoms?  

--

### Check out some ready-to-go geoms that might be of interest in the [{ggxmean}](https://evamaerey.github.io/ggxmean/) package... or [other extension packages](https://exts.ggplot2.tidyverse.org/).


--

## Or interested in working a bit more with geoms and making them available to more folks, but not interested in writing your own package?

--


### Join in on the development and validation of the ggxmean package for statistical educators and everyday analysis.

---



#### Photo Credits: Thalia Ruiz, Jennifer Pallian, Lisa Hanly, John Dancy, Piper Bchamp

#### More comprehensive extension resources: 

- https://www.youtube.com/watch?v=uj7A3i2fi54
- https://cran.r-project.org/web/packages/ggplot2/vignettes/extending-ggplot2.html

- https://ggplot2-book.org/extensions.html
