---
title: "easy geom recipes"
author: "Evangeline Reynolds"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Recipe #1:  `geom_label_id()`

---

## Step 0: use base ggplot2 to get the job done


```{r}
library(tidyverse)
cars %>% 
  mutate(id_number = row_number()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label(aes(label = id_number), 
             hjust = 1.2)
```

---

## Step 1: computation

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
compute_group_row_number <- function(data, scales){
  
  data %>% 
    # add an additional column called label
    mutate(label = row_number())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs 
  rename(x = speed, y = dist) %>% 
  compute_group_row_number() %>% 
  head()
```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs
- where does the function from above go?

```{r}
StatRownumber <- ggplot2::ggproto(`_class` = "StatRownumber",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_row_number)
```


---

## Step 3: define geom_* function

```{r}
stat_label_row_number <- function(mapping = NULL, data = NULL,
                                  geom = "label",
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatRownumber, # proto object from step 2
    geom = geom, # inherit other behavior; defaults to GeomLabel
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  stat_label_row_number(hjust = 1.2) # function in action
```

### Checkout using other geoms

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  stat_label_row_number(hjust = 1.2, geom = "text") # or `geom = GeomText`
```


### And check out conditionality!

```{r}
last_plot() + 
  aes(color = dist > 60) # Computation is within group
```


---

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


Hint:

```{r}
paste0("(", 1, ", ",3., ")")
```


---

# Recipe #2: `geom_point_xy_medians()`

--

- This will be a point at the median of x and y

## Step 0: use base ggplot2 to get the job done

```{r}
cars %>% 
  summarize(dist_median = median(dist),
            speed_median = median(speed)) ->
cars_medians

cars %>% 
  ggplot() + 
  aes(x = speed) + 
  aes(y = dist) + 
  geom_point() + 
  geom_point(data = cars_medians,
    color = "red", size = 4,
             aes(x = speed_median,
                 y = dist_median))
```


## Step 1: computation

```{r}
compute_group_xy_medians<- function(data, scales){
  data%>% 
    summarize(x = median(x),
            y = median(y)) ->
xy_medians
  xy_medians
}
cars%>%
  rename(x= speed, y=dist)%>%
  compute_group_xy_medians()
```


## Step 2: define ggproto

```{r}
StatXYMedians<-ggplot2::ggproto(`_class` = "StatXYMedians",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_xy_medians)
```


## Step 3: define geom_* function


```{r}
geom_point_xy_medians <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatXYMedians, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


## Step 4: Enjoy! Use your function

```{r}
cars%>%
  ggplot()+
  aes(x=speed, y=dist)+
  geom_point()+
  geom_point_xy_medians(color= "red")
```

### And check out conditionality!

```{r}
cars%>%
  ggplot()+
  aes(x=speed, y=dist, color= dist>40)+
  geom_point()+
  geom_point_xy_medians(size = 4)
```

## Now you ...


### Create the function `geom_point_xy_means()`.  

--

Recipe #2 modification:  `geom_point_xy_means()`

# Recipe #3:  `geom_point_lm_fitted()`

---

## Step 0: use base ggplot2 to get the job done

```{r}
model <- lm(formula = dist ~ speed, data = cars) 

cars %>% 
  mutate(fitted = model$fitted.values) ->
cars_w_fitted

cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_point(data = cars_w_fitted,
             aes(y = fitted),
             color = "blue")
```


## Step 1: computation

```{r}
compute_group_lm_fitted<- function(data, scales){
  model<-lm(formula= y~x, data=data)
  data%>% 
    mutate(y=model$fitted.values)-> data_w_fitted
  data_w_fitted
}
cars%>%
  rename(x= speed, y=dist)%>%
  compute_group_lm_fitted()
```


## Step 2: define ggproto

```{r}
StatLmFitted<-ggplot2::ggproto(`_class` = "StatLmFitted",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_lm_fitted)
```


## Step 3: define geom_* function


```{r}
geom_point_lm_fitted <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatLmFitted, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

## Step 4: Enjoy! Use your function

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method="lm", se= F)+
  geom_point_lm_fitted(color="blue")
```

---

## Now you...


### Create the function `geom_segment_lm_residuals()`.  
Hint: read about what aesthetics are required for segments.
