---
title: "easy geom recipes"
author: "Evangeline Reynolds, Morgan Brown"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Recipe #1:  `geom_label_id()`

---

## Step 0: use base ggplot2 to get the job done


```{r}
library(tidyverse)

cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label(aes(label = id_number), 
             hjust = 1.2)
```

---

## Step 1: computation

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
# you won't use the scales argument, but ggplot will later
compute_group_row_number <- function(data, scales){
  
  data %>% 
    # add an additional column called label
    # the geom we inherit from requires the label aesthetic
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs as columns
  rename(x = speed, y = dist) %>% 
  compute_group_row_number() %>% 
  head()
```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs
- where does the function from above go?

```{r}
StatRownumber <- ggplot2::ggproto(`_class` = "StatRownumber",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_row_number)
```


---

## Step 3: define geom_* function

```{r}
geom_label_row_number <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatRownumber, # proto object from Step 2
    geom = ggplot2::GeomLabel, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_label_row_number(hjust = 1.2) # function in action
```

### And check out conditionality!

```{r}
last_plot() + 
  aes(color = dist > 60) # Computation is within group
```




---

## Now you...

--

### Create the function `geom_text_coordinates()`.  

--

- geom should label point with its coordinates '(x, y)'
- geom should have behavior of geom_text (not geom_label)


Hint:

```{r}
paste0("(", 1, ", ",3., ")")
```


---

# Recipe #2: `geom_point_xy_medians()`

--

- This will be a point at the median of x and y

## Step 0: use base ggplot2 to get the job done

```{r}
library(palmerpenguins)
penguins <- remove_missing(penguins)
penguins %>% 
  summarize(bill_length_mm_median = median(bill_length_mm),
            bill_depth_mm_median = median(bill_depth_mm)) ->
penguins_medians

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm) + 
  aes(y = bill_length_mm) + 
  geom_point() + 
  geom_point(data = penguins_medians,
    color = "red", size = 4,
             aes(x = bill_depth_mm_median,
                 y = bill_length_mm_median))
```


## Step 1: computation

```{r}
compute_group_xy_medians <- function(data, scales){
  data %>% 
    summarize(x = median(x),
            y = median(y))
}
penguins %>%
  rename(x = bill_depth_mm, 
         y = bill_length_mm) %>%
  compute_group_xy_medians()
```


## Step 2: define ggproto

```{r}
StatXYMedians<-ggplot2::ggproto(`_class` = "StatXYMedians",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_xy_medians)
```


## Step 3: define geom_* function


```{r}
geom_point_xy_medians <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatXYMedians, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


## Step 4: Enjoy! Use your function

```{r}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, y = bill_length_mm)+
  geom_point()+
  geom_point_xy_medians(color = "red")
```

### And check out conditionality!

```{r}
penguins %>%
  ggplot()+
  aes(x = bill_depth_mm, 
      y = bill_length_mm, 
      color = species)+
  geom_point()+
  geom_point_xy_medians(size = 4)
```

## Now you ...


### Create the function `geom_point_xy_means()`.  

--

Recipe #2 modification:  `geom_point_xy_means()`

# Recipe #3:  `geom_point_lm_fitted()`

---

## Step 0: use base ggplot2 to get the job done

```{r}
model <- lm(formula = bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins %>% 
  mutate(fitted = model$fitted.values) ->
penguins_w_fitted

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_point(data = penguins_w_fitted,
             aes(y = fitted),
             color = "blue")
```


## Step 1: computation

```{r}
compute_group_lm_fitted<- function(data, scales){
  model<-lm(formula= y ~ x, data = data)
  data %>% 
    mutate(y=model$fitted.values)
}

# test out the function
penguins %>%
  # rename to explicitly state the x and y inputs
  rename(x = bill_depth_mm, y = bill_length_mm)%>%
  compute_group_lm_fitted()
```


## Step 2: define ggproto

```{r}
StatLmFitted<-ggplot2::ggproto(`_class` = "StatLmFitted",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_lm_fitted)
```


## Step 3: define geom_* function


```{r}
geom_point_lm_fitted <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatLmFitted, # proto object from step 2
    geom = ggplot2::GeomPoint, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

## Step 4: Enjoy! Use your function

```{r}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F)+
  geom_point_lm_fitted(color="blue")
```

### And check out conditionality

```{r}
penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method="lm", se= F) +
  geom_point_lm_fitted() + 
  facet_wrap(facets = vars(species))
```

---

## Now you... Create the function `geom_segment_lm_residuals()`. 

### Hint: read about what aesthetics are required for segments.  We'll give you Step 0 this time...

## Step 0: use base ggplot2 to get the job done

```{r}
model <- lm(formula = bill_length_mm ~ bill_depth_mm, 
            data = penguins) 

penguins %>% 
  mutate(fitted = model$fitted.values) ->
penguins_w_fitted

penguins %>% 
  ggplot() + 
  aes(x = bill_depth_mm, y = bill_length_mm) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  geom_segment(data = penguins_w_fitted,
             aes(yend = fitted, xend = bill_depth_mm),
             color = "blue")
```


