---
title: "latinR proposal: Flipping the script with new geography-targeted Stats"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

ggplot2 is a beloved graphical system because of it's elegance and flexibility.  Usually a graph is composed of three elements (and ggplot2 can take care of the rest).  These elements are data (which dataframe will be the basis of the graphic), variable aesthetic mapping (should variables be represented by x position, y position, color, linewidth etc), and geom, (the mark that should take on these aesthetics).

ggplot2 added more geographic capabilities with the addition of it's geom_sf_*() functions. However, geographic mapping with geom_sf usage may feel a little out-of-step with ggplot2's other plot builds.  Let's consider creating a choropleth of characteristic that varies with greographic area:  

```{r, df}
nc_shapes <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

my_nc_data_w_shapes <- nc_shapes %>% left_joint(my_nc_data, by = "county_name")  
  
ggplot(nc) +
  geom_sf() + 
  aes(fill = my_var)
```

A first step is to find shape data for the areas of interest, ideally in a data frame with one column identifying the geographic area, and another list-column often named `geometry` with the geographic information included. Then the actual data of interest is joined to the geography data frame via the identifying column.  Then this data can be input into ggplot().  To render a map, geom_sf() can be added with no aesthetic mapping required, as under the hood, this layer is directed to look for a column named `geometry`. To make the map a choropleth, aesthetic mapping instruction `aes(fill = my_var)` would be required. 

This talk will propose creating a Stat to allow for a workflow that's more similar to the 'classic' ggplot2 build.  







## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)

```

## Experiment

```{r}

```



## Closing remarks, Other Relevant Work, Caveats
