---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Step 0. Status Quo: base ggplot2

```{r cars}
library(tidyverse)

gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  mutate(is_argentina = country == "Argentina") %>% 
  arrange(is_argentina) %>%
  mutate(country = fct_inorder(country)) %>% 
  ggplot() + 
    aes(x = year, y = lifeExp, group = country) + 
  geom_point() + 
  geom_line(lwd = 3) + 
  aes(color = is_argentina) + 
  scale_color_manual(values = c("grey50", "magenta"))
```

## Step 1. define compute. test.

```{r}
compute_panel_highlight_lines <- function(data, scales, which_id = "Argentina"){

data %>% 
  mutate(ind_id = id %in% which_id) %>% 
  arrange(ind_id) %>%
  mutate(group = fct_inorder(id))
  
}

# test
gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  select(x = year, y = lifeExp, id = country) %>% 
  compute_panel_highlight_lines
```

# step 2 pass to ggproto. test

```{r}
statHighlight <- ggproto("statHighlight",
                         Stat,
                         compute_panel = compute_panel_highlight_lines,
                         default_aes = aes(color = after_stat(ind_id)),
                         required_aes = c("x", "y", "id")
                         )

# test.
gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line(color = "grey") + 
  layer(geom = "line", stat = statHighlight, position = "identity",
        params = list(linewidth = 3, alpha = 1, which_id = "Peru"))
```

# Step 3. make user-facing function & test. (I'm not being very careful here)

```{r}
geom_line_highlight <- function(mapping = NULL, # global aesthetics will be used if NULL
  data = NULL, # global data will be used if NULL
  position = "identity", 
  na.rm = FALSE, 
  show.legend = NA,
  inherit.aes = TRUE, 
  ...  ) {
  
  layer(geom = "line", stat = statHighlight, 
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
  
  
}

# test
gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line_highlight(which_id = "Bolivia") 
```

# Step 4. Specific request: preset colors - an geom + scale approach

# Step 4.aa Create custom 2 color manual scale function.

```{r}
scale_color_manual_2 <- function(val1 = "grey", val2 = "red"){
  
  scale_color_manual(values = c(val1, val2))
  
}

```
  
# Step 4.ab. test both components together
  
```{r}
gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line_highlight(which_id = "Bolivia") + 
  scale_color_manual_2()
```
  
# Step 4.ba. bundle into pseudo geom (layer+scale)

```{r}
geom_line_highlight_w_scale <- function(val1 = "grey", highlight_color = "pink", ...){
  
  list(geom_line_highlight(...),
      scale_color_manual_2(val1 = val1, val2 = highlight_color)
  )
      
}
```

# Step 4bb test the geom-scale
  
```{r}  
gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line_highlight_w_scale(which_id = "Colombia",
                              linewidth = 4,
                              highlight_color = "goldenrod3")

gapminder::gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot() + 
  aes(x = year, y = lifeExp, id = country) + 
  geom_point() + 
  geom_line_highlight_w_scale(which_id = "Chile", 
                         val1 = "white", 
                         highlight_color = "darkred",
                         linewidth = 2)
```

# 4bb test overwrite scale

```{r}
last_plot() + 
  scale_color_manual(values = c("gray", "red"), 
                     breaks = c(F,T), 
                     labels = c("Peers", "Chile"))
```

---

# Other data.


```{r}
data <- tibble::tribble(
    ~x,  ~group,        ~y, ~measure,  ~base,
"FY18",   "aaa", 0.5603448,      260,    464,
"FY15",   "aaa", 0.5081269,     1313,   2584,
"FY19",   "aaa", 0.5799373,      185,    319,
"FY16",   "aaa", 0.5225225,      580,   1110,
"FY13",   "aaa", 0.4779116,      595,   1245,
"FY17",   "aaa", 0.5502471,      334,    607,
"FY14",   "aaa", 0.5339007,     1882,   3525,
"FY20",   "aaa", 0.4960998,      318,    641,
"FY21",   "aaa", 0.4765840,      173,    363,
"FY21",   "bbb", 0.5135802,      208,    405,
"FY20",   "bbb", 0.5384615,      259,    481,
"FY17",   "bbb", 0.4628099,       56,    121,
"FY18",   "bbb", 0.5474453,       75,    137,
"FY15",   "bbb", 0.4705882,       72,    153,
"FY19",   "bbb", 0.6083916,       87,    143,
"FY14",   "bbb", 0.5097087,      105,    206,
"FY16",   "bbb", 0.5395683,       75,    139,
"FY13",   "bbb", 0.5432099,      132,    243,
"FY14",   "ccc", 0.5326591,     2528,   4746,
"FY21",   "ccc", 0.4320038,      899,   2081,
"FY13",   "ccc", 0.5420887,     3265,   6023,
"FY16",   "ccc", 0.5075188,      810,   1596,
"FY15",   "ccc", 0.5116469,     1252,   2447,
"FY18",   "ccc", 0.4206349,      477,   1134,
"FY17",   "ccc", 0.4564565,      456,    999,
"FY19",   "ccc", 0.4651429,      407,    875,
"FY20",   "ccc", 0.4759398,      633,   1330
)

```


```{r}
data %>% 
  mutate(group = factor(group)) %>% 
  mutate(x = factor(x)) %>% 
ggplot() +
  aes(x = x, y = y, id = group) + 
  geom_line_highlight_w_scale(which_id = "ccc", 
                              linewidth = 3,
                         highlight_color = "darkred")

  
```




## Closing remarks, Other Relevant Work, Caveats
