---
title: "Indicator recode"
author: "Evangeline Reynolds"
date: "5/23/2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# context: indicator variables well understood

- https://emilyriederer.netlify.app/post/column-name-contracts/

# problem: content loss using indicator variables directly in data products

```{r cars}
library(tidyverse)

tidytitanic::passengers %>% 
  ggplot() + 
  aes(x = survived) + 
  geom_bar()

tidytitanic::passengers %>% 
  count(survived)

tidytitanic::passengers %>% 
  janitor::tabyl(sex, survived)

tidytitanic::passengers %>% 
ggplot() + 
  aes(x = sex) + 
  geom_bar() + 
  facet_grid(~ survived)
```

# one-off solution: manual indicator recode, but is repetitive

addition downside is unanticipated reordering depending on value of true case v. false case category 

```{r}
tidytitanic::passengers %>% 
  tibble() %>% 
  mutate(cat_survived = ifelse(survived, "survived", "not survived"), 
         .before = 1)

tidytitanic::passengers %>% 
ggplot() + 
  aes(x = sex) + 
  geom_bar() + 
  facet_grid(~ ifelse(survived, "survived", "not survived"))
```



# proposed many-times solution: ind_recode() function uses variable name as starting point for creating true and false human understandable categories
  
```{r}
ind_recode <- function(var, var_prefix = "ind_", negator = "not", cat_true = NULL, cat_false = NULL, rev = FALSE){
  
  if(is.null(cat_true)){
    cat_true = deparse(substitute(var)) %>% 
      str_remove(var_prefix) %>% 
      str_replace_all("_", " ")
  }
  
  if(is.null(cat_false)){
    cat_false = paste(negator, cat_true)
  }

  if(rev){
    ifelse({{var}}, cat_true, cat_false) %>% 
      factor(levels = c(cat_true, cat_false))
  }else{
    ifelse({{var}}, cat_true, cat_false) %>% 
      factor(levels = c(cat_false, cat_true))
  }
  
}

tibble(ind_female = c(0,0,1,1,1 ,0 ,0)) %>% 
  mutate(cat_female  = ind_recode(ind_female))


```

# Examples

```{r}
tidytitanic::passengers %>% 
  tibble() %>% 
  mutate(cat_survived = ind_recode(survived))

tidytitanic::passengers %>% 
ggplot() + 
  aes(x = ind_recode(survived)) + 
  geom_bar()

# or
last_plot() +
  aes(x = ind_recode(survived, cat_false = "perished"))

  
# or
last_plot() +
  aes(x = ind_recode(survived, cat_false = "didn't", rev = T)) + 
  labs(x = FALSE)

tidytitanic::passengers %>% 
ggplot() + 
  aes(x = sex) + 
  geom_bar() + 
  facet_grid(~ ind_recode(survived))
```

# Known Limitations: not for use with magrittr pipe (but base pipe works!)

```{r}
tidytitanic::passengers %>% 
ggplot() + 
  aes(x = sex) + 
  geom_bar() + 
  facet_grid(~ survived %>% ind_recode())

tidytitanic::passengers %>% 
ggplot() + 
  aes(x = sex) + 
  geom_bar() + 
  facet_grid(~ survived |> ind_recode())
```
