---
title: "Towards incremental table specification"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

# Part 0.  **Objective** take many argument function, where arguments are orthogonal, and make a pipable interface

## `pivot_count()` create count table based on direct user inputs

```{r, echo = T}
library(tidyverse)
pivot_count <- function (data, rows = NULL, cols = NULL) {
    
  cols_quo <- rlang::enquo(cols)
  
  grouped <- data %>% dplyr::group_by(
    dplyr::across(c({{cols}}, {{rows}})))
  
  summarized <- grouped |> 
    dplyr::count() |> 
    dplyr::ungroup()

    if (rlang::quo_is_null(cols_quo)) {
        summarized
    }
    else {
        summarized %>% tidyr::pivot_wider(names_from = {{cols}}, 
                                    values_from = n)
    }
}
```

## status quo usage: enter arguments at once

```{r}
Titanic |> 
  data.frame() |>
  uncount(Freq) ->
tidy_titanic

pivot_count(data = tidy_titanic,
            rows = Survived,
            cols = c(Sex, Age))
```

## Desired usage: Piped

```{r eval = F}
ggtable(mtcars) |> # a count of observations
  set_rows(cyl) |> # a count by cyl (cly name each row)
  set_cols(c(gear, am)) |> # count by cyl and gear (gears head cols)
  set_value(mpg) |> # sum of mpgs by cyl and gear
  set_fun(mean) |> # average mpg by cyl and gear
```

# Part. 1. **Implementation**

## Step 1. **New 3S object** Create `new_tidypivot()` which defines components in '`tp`' a 3S object.

```{r}
library(tidyverse)

new_tidypivot <- function(data = data.frame(), 
                          rows = NULL, 
                          columns = NULL, 
                          value = NULL, 
                          weight = NULL) {
  
  # table specification components !
  tp <- list(
    data = data,
    rows = rows,
    columns = columns,
    value = value,
    weight = weight  
    # more 'slots' to be added
  )
  
  # declare class 'tidypivot'
  class(tp) <- "tidypivot"
  
  # Return the created object
  invisible(tp)
  
}
```

## Step 2. **Print**: Once specification is defined, actually do  manipulation to produce the table - we use a routine from the {tidypivot} package.



### `return_specified_table()` create table based on table *specification* (`tp`)

```{r}
return_specified_table = function(tp){

      out <- 'pivot_count(thedata, rows, cols)'

      str_replace_or_null <- function(x, pattern, replacement){
        
        if(is.null(replacement)){x |> str_replace(pattern, 'NULL')}
        else{x |> str_replace(pattern, replacement)}
        
      }
      
      # out <- str_replace_or_null(out, "data",  tp$data)
      out <- str_replace_or_null(out, "rows",  tp$rows)
      out <- str_replace_or_null(out, "cols",  tp$cols)

      eval(parse(text = out))         
      
}
```

###  `print.tidypivot` declare print method for `tp`, the 3S objects

```{r}
print.tidypivot <- function(tp){
  
  return_specified_table(tp)
  
}
```
  
  
## Step 3. **user-facing functions!**

### `ggtable()` *initiate table*
  
```{r}  
ggtable <- function(data){
  
  thedata <<- data # don't love this

  tp <- new_tidypivot(deparse(substitute(thedata)))

  print(return_specified_table(tp))

  invisible(tp)
  
  # tp
  
  last_tp <<- tp

}

ggtable(cars)

```

### `set_rows()`, (the other 'set' functions are hidden - set up the same way...)

```{r}
set_rows <- function(tp, rows = NULL){
  
tp$rows <- deparse(substitute(rows))
 
  print(return_specified_table(tp))
  
  invisible(tp)
  
    last_tp <<- tp
}
```

```{r, echo = F}
set_cols <- function(tp, cols = NULL){
  
tp$cols <- deparse(substitute(cols))

  print(return_specified_table(tp))
  
  invisible(tp)
  
    last_tp <<- tp
}


set_fun <- function(tp, fun = sum){

tp$fun <- deparse(substitute(fun))


  print(return_specified_table(tp))

  invisible(tp)

    last_tp <<- tp

  
}

set_value <- function(tp, value = NULL){
  
tp$value <- deparse(substitute(value))

  print(return_specified_table(tp))
  
  invisible(tp)
  
  last_tp <<- tp
  
}

set_weight <- function(tp, weight = NULL){
  
tp$weight <- deparse(substitute(weight))

  print(return_specified_table(tp))
  
  invisible(tp)
  
  last_tp <<- tp
  
}

```

### `last_table()`

```{r}
last_table <- function(){
  
  last_tp
  
}
```

# Part 2. **Test functionality!**

## Slow build with `last_table()`

```{r}
Titanic |> 
  data.frame() |>
  uncount(Freq) ->
tidy_titanic

tidy_titanic |> tibble()

tidy_titanic |>
  ggtable() 

last_table() |> 
  set_rows(Survived)

last_table() |> 
  set_cols(c(Sex))

last_table() |>
  set_rows(Class)
```


## fast, piped `|>` build *uh-oh, we get to see the in-between tables even when we don't use last table!  Ah!!!!*

```{r}
ggtable(mtcars) |> 
  set_rows(cyl) |> 
  set_cols(am) |>
  set_cols(c(gear, am)) 
```


# Part 3. **Dispatch** Define behavior when tps are piped to other environmenst 

Future work
