---
title: "ggplot2 extension as an analytic care task"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

'Care tasks describes any task, chore, or errand that is required to care for self and keep life going.'

'Everyone is worthy of self-care (ggplot2 extension), including you!'

## Intro Thoughts

Attention to systems for mundane and routine tasks is important. 

'Care tasks exist for one reason only... to make your body and space functional enough for you to easily experience the joy this world has to offer.'
― K.C. Davis, How to Keep House While Drowning: A Gentle Approach to Cleaning and Organizing


Housekeeping can be really inspirational and focus on being aesthetically pleasing rather than functional.  (flashier extension packages maybe divert our attention from a practical approach)

care tasks are morally neutral  (if I want to write geom_pie...) 



Baskets

Anything that is worth doing is worth doing half-assed.


## Status Quo

```{r cars}
library(tidyverse)

```

## Experiment

```{r}

```



## Closing remarks, Other Relevant Work, Caveats
