---
title: "janitor::tabyl()"
subtitle: ""
author: ""
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = T)
library(flipbookr)
library(tidyverse)
```


class: center, inverse, middle

# Do you know the [{janitor}](https://github.com/sfirke/janitor) package?

--

# and `janitor::tabyl()`?

--

## 'A fully-featured alternative to table()'?

---
class: center, inverse middle

# Let's have a look at how to produce some report-ready tables!


---

`r chunk_reveal("cars_multi")`

```{r cars_multi, include = FALSE}
library(tidyverse)
datasets::Titanic %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  tidyr::uncount(weights = freq) %>% 
  janitor::tabyl(survived, class) %>% 
  janitor::adorn_totals(c("row", "col")) %>% 
  janitor::adorn_percentages(denominator = "col") %>% 
  janitor::adorn_pct_formatting() %>% 
  # add Ns also known as counts
  janitor::adorn_ns(position = "front")
```

---

`r chunk_reveal("cars_multi2", break_type = 2)`

```{r cars_multi2, include = FALSE}
datasets::Titanic %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  tidyr::uncount(weights = freq) %>% 
  janitor::tabyl(survived, class, sex) %>% #<<
  janitor::adorn_totals(c("row", "col")) %>% 
  janitor::adorn_percentages(denominator = "col") %>% 
  janitor::adorn_pct_formatting() %>% 
  janitor::adorn_ns(position = "front")
```

---
class: inverse, middle, center

# A bit more about how it works.

--

## Let's see tabyl carrying along the raw counts as meta data using `str()` on each step in the same pipeline. 

--

## This allows tabyl() to calculate other quantities of interest, like percentages and percentages on totals wtih its adorn_*() functions.

---

`r chunk_reveal("cars_multi3", left_assign_add = "str()", left_assign = T)`

```{r cars_multi3, include = FALSE}
output <- datasets::Titanic %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  tidyr::uncount(weights = freq) %>% 
  janitor::tabyl(survived, class) %>% 
  janitor::adorn_totals(c("row", "col")) %>% 
  janitor::adorn_percentages(denominator = "col") %>% 
  janitor::adorn_pct_formatting() %>% 
  janitor::adorn_ns(position = "front") 
```

---

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```

---
class: center, middle

# [https://github.com/sfirke/janitor](https://github.com/sfirke/janitor)

---
