---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r cars}
library(tidyverse)

```

## Experiment

1. What are the three components of a function?

> The formals(), the list of arguments that control how you call the function.

> The body(), the code inside the function.

> The environment(), the data structure that determines how the function finds the values associated with the names.

2. What does the following code return?


```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}


f1(x = 1)()
f1(1)
```
3. How would you usually write this code?

```{r}
`+`(1, `*`(2, 3))

2*3 + 1
```

4. easier to read?

```{r}
# trim, na.rm, x
mean(, TRUE, x = c(1:10, NA))

# explicit
mean(trim = 0, na.rm = TRUE, x = c(1:10, NA))

# moving to usual positions
mean(c(1:10, NA), na.rm = T)
```

# Does the following code throw an error when executed? Why or why not?

```{r, error = T}
f2 <- function(a, b) {
  a * 10
}

f2(a = 10, b = stop("This is an error!"))

f2(a = stop("This is an error"))

```

# What is an infix function? How do you write it? What’s a replacement function? How do you write it?

> Infix functions get their name from the fact the function name comes inbetween its arguments, and hence have two arguments. R comes with a number of built-in infix operators

> replacement: functions that replace values by assignment, like names(df) <- c("a", "b", "c") . They actually look like prefix functions.

> replace() function in R Programming Language is used to replace the values in the specified string vector x with indices given in the list by those given in values.


---


```{r}
# bind function to a name
myfun <- function(x){x + 1}

# don't

# fun in a list
funs <- list(
  half = function(x) x / 2,
  double = function(x) x * 2
)

funs$double(10)
```


# do.call

```{r}
args <- list(x = 1:10, na.rm = TRUE)


do.call(mean, args)


```

- Given a name, like "mean", match.fun() lets you find a function. Given a function, can you find its name? Why doesn’t that make sense in R?


- It’s possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?
# 

```


# How do you ensure that cleanup action occurs regardless of how a function exits?




## Closing remarks, Other Relevant Work, Caveats
