---
title: "ggwhere? ggplot2 extenders community of practice"
author: "Evangeline Reynolds"
date: "3/21/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library(tidyverse)
ggplot() + 
  coord_equal() + 
  ggstamp::stamp_circle(alpha = .3, fill = 'blue',
                        x0 = -1) + 
  ggstamp::stamp_text(label = "programming\ngods", x = -1) + 
  ggstamp::stamp_circle(alpha = .3, fill = 'magenta', 
                       x0 = 1) + 
  ggstamp::stamp_text(label = "stats\ngeniuses", x = 1) +
  ggstamp::stamp_label(label = "ggplot2\nextenders") + 
  ggstamp::stamp_label(label = "me", y = -2, x = 1)
```


---

'This is what I'm thinking, this is how I'm picturing this plot. Like you're the computer 'Go and do it'.'

---

'And, you know, I'd get a dataset. And, in my head I could very clearly kind of picture, I want to put this on the x-axis. Let's put this on the y-axis, draw a line, put some points here, break it up by this variable. And then, like, getting that vision out of my head, and into reality, it's just really, really hard. Just, like, felt harder than it should be. Like, there's a lot of custom programming involved,'

---

Spoiled by ggplot2.  Used to flying...




---

2020 Thomas Lin Pederson:   * **You** can extend ggplot2 *

```{r}


```

---

'This is what I'm thinking, this is how I'm picturing this plot. Like you're *ggplot2* 'Go and do it'.'


---

- Experiential learning, freedom of play (try a lot of stuff) - #tidytuesday model
- Document successes process (tutorials)
- Building safety nets, catch buddies, and harnesses (ggplot2 extenders)

---

ggplot2 extenders

```{r}




```


---






