---
title: "ggdandalion case study: Melodically placed song lyrics in a round"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
options(tidyverse.quiet = TRUE)
```



## Intro, a lyrics-melody data frame

```{r}
library(tidyverse)


beluga_song <- ggdoremi:::join_phrases_drm_lyrics()

beluga_song %>% head()
```


## Dandalion song w/o extension

```{r, fig.height=12, fig.width=12 }

compute_panel_dandalion <- function(data, scales, start_angle =  -90){
  
  spacer_df <- tibble(spacer = 1)
  
  data <- bind_rows(spacer_df, data, spacer_df)
  
  data$id <- 1:nrow(data)
  # nrow(data)
  
  single_wedge_angle <- 1/nrow(data)*360
  half_wedged_angle <- single_wedge_angle/2
  
  data$around <- (nrow(data)):1/nrow(data) 

  data$circ_angle <- data$around*2*pi + 
    (start_angle -half_wedged_angle)*2*pi/360 
  data$x <- data$radius * cos(data$circ_angle)
  data$xend <- 0
  data$yend <- 0
  data$y <- data$radius * sin(data$circ_angle)
  data$angle <- data$around*360 + 180 + 90 + 
    start_angle - half_wedged_angle
  
  data %>% 
    slice(-1) %>% 
    slice(-nrow(.))
  
}

beluga_song %>% head(4) %>% 
  # head(30) %>% 
  mutate(radius = as.numeric(drm) + 10) %>% 
  compute_panel_dandalion() %>% 
  ggplot() + 
  ggstamp::stamp_circle(radius = 8+10, alpha = 0, size = .25, linetype = "dotted") +
  ggstamp::stamp_circle(radius = 8+12, alpha = 0, size = .25, linetype = "dotted") +
  ggstamp::stamp_circle(radius = 8+14, alpha = 0, size = .25, linetype = "dotted") +
  aes(x = x, radius = radius, y = y, label = lyric, angle = angle) + 
  geom_segment(xend = 0, yend = 0, color = "black", linewidth = .5) +
  geom_point(shape = 21, size = 16, fill = "white") +
  # aes(color = doremi) +
  geom_text() +
  coord_equal()  + 
  theme_void()
```

# statexpress enabled extension

## write extension

```{r statexpress}

stat_dandalion <- function(geom = "segment", ...){
  statexpress::stat_panel(compute_panel_dandalion, geom, ...)
}

geom_dandalion_segment <- stat_dandalion
geom_dandalion_text <- function(...){stat_dandalion(geom =  "text", ...)}
geom_dandalion_point <- function(...){stat_dandalion(geom =  "point", ...)}


```

```{r stamp}
stamp_staff_dms <- function(radius_addition = 10 ){
  
  list(
  ggstamp::stamp_circle(radius = 8+0+radius_addition, alpha = 0, size = .25, linetype = "dotted"),
  ggstamp::stamp_circle(radius = 8+2+radius_addition, alpha = 0, size = .25, linetype = "dotted"),
  ggstamp::stamp_circle(radius = 8+4+radius_addition, alpha = 0, size = .25, linetype = "dotted")
  )
  
  
}


stamp_staff_dmst <- function(radius_addition = 10 ){
  
  list(stamp_staff_dms(radius_addition = radius_addition), 
      ggstamp::stamp_circle(radius = 8+6+radius_addition, 
                            alpha = 0, size = .25, 
                            linetype = "dotted")
  )
  
  
}

stamp_staff_dmstr <- function(radius_addition = 10 ){
  
  list(stamp_staff_dmst(radius_addition = radius_addition), 
      ggstamp::stamp_circle(radius = 8+8+radius_addition, 
                            alpha = 0, size = .25, 
                            linetype = "dotted")
  )
  
  
}

```


## Use extension

### Beluga

```{r beluga, fig.height=12, fig.width=12 }
ggdoremi:::join_phrases_drm_lyrics() ->
beluga_song

head(beluga_song) 

beluga_song %>% 
  ggplot() + 
  stamp_staff_dms(3) +
  aes(radius = as.numeric(drm)+3) +
  geom_dandalion_segment() +
  geom_dandalion_point(size = 21,
                       shape = 21) +
  aes(label = lyric) +
  geom_dandalion_text(size = 5) +
  coord_equal() + 
  theme_void() + 
  aes(fill = doremi) + 
  scale_fill_brewer("blues")
  
  


```

## ABCs

To allow some melodic plotting, first of all thinks about dataframes. Here we see lyrics and tones (drmfslt) are entered independently as text.  Then a function is used to pair up lyrics with the associated note.  

Tones are positioned in three octaves.  DRM, drm, 123 so single character represents relative position.

Looking at twinkle twinkle - Mozart-Salzburg connection!

```{r abc}
lyrics_abc <- "A B C D E F G
H I J K L M N O P
Q R S T U V
W X Y Z
Now I know my A B Cs
Next time won't you sing with me"


drm_abc <- "ddsslls
ffmmrrrrd
ssfmmr
sfmr
ddsslls
ffmmrrd"


abc_df <- ggdoremi:::join_phrases_drm_lyrics(drm_phrases = drm_abc, 
                                   lyrics_phrases = lyrics_abc)  

abc_df

abc_df %>% 
  ggplot() + 
  aes(radius = as.numeric(drm) - 2) +
  stamp_staff_dms(-2) +
  geom_dandalion_segment(color = "black", 
                         size = .5) +
  geom_dandalion_point(size = 13, 
                       fill = "white",
                       shape = 21) +
  aes(label = lyric) +
  geom_dandalion_text() +
  coord_equal() + 
  theme_void() + 
  aes(size = 5) + 
  scale_size(range = c(3,5))

last_plot() %+%
  (ggdoremi:::join_phrases_drm_lyrics(drm_abc, 
                                   lyrics_abc)[1:26,]) + 
  aes(size = 8) + 
  theme(legend.position = "none")
```

### twinkle twinkle

```{r}
lyrics_abcd <- '一 闪 一 闪 亮 晶 晶'

drm_abcd <- 'ddsslls'


ggdoremi:::join_phrases_drm_lyrics(drm_abcd, 
                                   lyrics_abcd) %>% 
  ggplot() + 
  aes(radius = as.numeric(drm) - 2) +
  stamp_staff_dms(-2) +
  geom_dandalion_segment(color = "black", 
                         size = .5) +
  geom_dandalion_point(size = 13, 
                       fill = "white",
                       shape = 21) +
  aes(label = lyric) +
  geom_dandalion_text(angle = 0) +
  coord_equal() + 
  theme_void() + 
  aes(size = 5)
  theme(legend.position = "none")
```

```{r, fig.height=10, fig.width=10 }
lyrics_hb <- 
'Hap-py birth-day to you
Hap-py birth-day to you
Hap-py birth-day dear X-X
Hap-py birth-day to you'

drm_hb <- 
'SSLSdT
SSLSrd
SSsmdTL
ffmdrd'


hb_df <- ggdoremi:::join_phrases_drm_lyrics(drm_hb, lyrics_hb) 



hb_df %>%
  # head(30) %>%
  ggplot() +
  stamp_staff_dms(10) +
  aes(radius = as.numeric(drm) +10 ) +
  geom_dandalion_segment(xend = 0, yend = 0) +
  geom_dandalion_point(shape = 21, size = 20, fill = "white") +
  # aes(color = doremi) +
  aes(label = lyric) +
  geom_dandalion_text(size = 5) +
  coord_equal() + 
  theme_void() + 
  guides(color = "none", 
         label = "none")

```

```{r, fig.height=12, fig.width=12 }
lyrics_doe_a_deer <-
'Doe a deer a fe-male deer
ray a drop of gol-den sun
me a name I call my-self
far a long long way to run
sew a need-le pull-ing thread
la a note to fol-low sol
tea a drink with jam and bread
that will bring us back to do o o o'

drm_doe_a_deer <-
'drmrmrm
rmffmrf
mfsmsms
fsllsfl
sdrmfsl
lrmfslt
tmfslt1
1tlsts1smr'


ggdoremi:::join_phrases_drm_lyrics(drm_doe_a_deer, lyrics_phrases = lyrics_doe_a_deer) %>%
  # head(30) %>%
  ggplot() +
  stamp_staff_dmst(0) +
  aes(radius = as.numeric(drm) ) +
  geom_dandalion_segment(xend = 0, yend = 0) +
  geom_dandalion_point(shape = 21, size = 19, fill = "white") +
  aes(color = doremi, label = lyric) +
  geom_dandalion_text(size = 7) +
  coord_equal() + 
  theme_void() + 
  guides(color = "none", label = "none")

last_plot() + 
  aes(label = doremi)

```



## Closing remarks, Other Relevant Work, Caveats
