---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```


@EvaMaeRey I am glad to be able to contribute! Thanks for the invitation to talk!
I give here some examples using only 'ggrepel', starting from Teun's code. As you already noticed, repulsion in 'ggrepel' uses a random displacement, so running the same code can result in a different position of labels in a plot. In addition, as the repulsion is applied during rendering, the position of repulsed labels depends on the size used to render the plot. Repulsion can disabled or restricted to one direction. Force can be adjusted as well and maximum number of overlaps at which the repulsion algorithm gives up.

By setting labels to `""`, 'ggrepel' does not attempt to change their positions, however, it repulses other labels away from the unlabelled points. Is this slower? yes to some extent, but not much, I think. It is at least much faster that repulsing actuals labels for every point. (If the labels are set to `NA` the whole row is removed from `data` before the plot is rendered. This is why `""` was chosen as the marker for no label.)

While position functions from 'ggplot2' only modify `x` and `y`, the `position_nudge_repel()` retains in data a copy of the original values of `x` and `y`.  (I copied this approach in 'ggpp', but using a slightly different naming scheme, so that the position functions in 'ggpp' could remain compatible with normal geoms from 'ggplot2' and extensions).

```{r}
library(tidyverse)
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  filter(continent == "Americas") ->
gap2002_americas_df

select_countries <- c("Chile" = "Chile", "United States" = "The US is here")

select_countries[as.character(gap2002_americas_df$country)]
```



```{r}
library(tidyverse)
library(ggrepel)

# sometimes labels overlap the points, depending on the random numbers seed
gapminder::gapminder |>
  filter(year == 2002) |>
  ggplot() +
  aes(gdpPercap, lifeExp) +
  geom_point(colour = "darkgrey") ->
base_plot  

aes_label_nas <- aes(label = select_countries[as.character(country)])
  
geom_text_repel_favs <- function(...){
    geom_text_repel(na.rm = TRUE, 
    box.padding = unit(1, "cm"),
    point.padding = unit(5, "mm"),
    arrow = arrow(length = unit(2, "mm"), 
                  type = "closed"), ...)
}

base_plot + 
  aes_label_nas +
  geom_text_repel_favs()

```



```{r}
# labels never overlap the points as long as there is space
aes_label_empty_string <- aes(label = select_countries[as.character(country)] %>% 
        replace_na(""))

base_plot + 
  aes_label_empty_string +   # New label
  geom_text_repel_favs(
        max.overlaps = 30    # overlaps
  )

```


```{r}
# In this case we can restrict repulsion to the vertical axis

base_plot + 
  aes_label_empty_string + 
  geom_text_repel_favs(
    direction = "y",       #<< direction y
    max.overlaps = 30
  )
```


```{r}
# Using nudging to start movement upwards
base_plot + 
  aes_label_empty_string + 
  geom_text_repel_favs(
    nudge_y = 3,           #<< New
    direction = "y",
    max.overlaps = 30
  )
```


```{r}
base_plot + 
  aes_label_empty_string + 
  geom_text_repel_favs(
    nudge_y = 4,
    direction = "y",
    force = 0,          #<< force changes
    force_pull = 0,     #<< force changes
  )
```

---

In this case, the best approach seems to be the 4th: nudging + repulsion along _y_.

Tomorrow I should have a bit more time, and I will post here some variations using 'ggpp'.


---


Hi. 

Here are some examples with 'ggpp' and with 'ggrepel'. I do not think any automatic approach can provide plots as nice as those than be obtained by manual placement. One important difference between 'ggrepel' and 'ggpp' is that `geom_text_repel()` avoids clipping of labels at the edge of the plotting area, while `geom_text_s()` behaves like `geom_text()` and may clip part  of labels. Of course, the solution is to manually expand the axis limits.

Padding 'ggpp' does not support 'grid' units, differemtly to 'ggrepel'.

```{r}
library(tidyverse)
library(ggpp)
#> Registered S3 methods overwritten by 'ggpp':
#>   method                  from   
#>   heightDetails.titleGrob ggplot2
#>   widthDetails.titleGrob  ggplot2
#> 
#> Attaching package: 'ggpp'
#> The following object is masked from 'package:ggplot2':
#> 
#>     annotate
library(ggrepel)

# Using 'ggpp' ------------------------------------------------------------

# I label the same two countries as in the original example

geom_text_s_favs <- function(...){
  geom_text_s(
    na.rm = TRUE, 
    box.padding = 0.5,
    point.padding = 0.3,
    arrow = arrow(length = unit(2, "mm"), type = "closed"), ...
  )
}


# Using non-repulsive geom
base_plot +
  aes_label_nas + 
  geom_text_s_favs(
    nudge_y = 4
  )

```


```{r}

# Justification changes the attachment of the arrow
base_plot +
  aes_label_nas +
  geom_text_s_favs(
    aes(label = c("Chile" = "Chile", "United States" = "The US is here")[as.character(country)]),
    nudge_y = 4, nudge_x = 1000,
    hjust = "left"
  )
```


```{r}

# Nudge can be a vector (it will be recycled if necessary)
base_plot +
  aes_label_nas + 
  geom_text_s_favs(
    nudge_y = c(-4, 4), nudge_x = 1000,
    hjust = "left",
  )
```


```{r}
# Nudging to a target position
base_plot +
  aes_label_nas +
  geom_text_s_favs(
    position = position_nudge_to(y = 82),
    hjust = "left"
  )
```


```{r}

# Here I label more countries than in the original example

some.countries <- c("Chile" = "Chile",
                    "Argentina" = "Argentina is an interesting case" %>% str_wrap(20),
                    "Paraguay" = "Paraguay",
                    "Bolivia" = "Bolivia",
                    "Peru" = "Perú",
                    "Somalia" = "Somalia",
                    "United States" = "The US is here")

aes_label_some_countries <- aes(label = some.countries[as.character(country)])

# Nudging away from the cloud of points, which does not work well without repulsion
base_plot +
  geom_point(aes(alpha = country %in% names(some.countries)), show.legend = FALSE) ->
base_plot2

base_plot2 +
  aes_label_some_countries +
  geom_text_s_favs(
    position = position_nudge_line(y = 4, x = 1000, method = "spline"),
    hjust = "outward"
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}


# Nudging away from the cloud of points, increasing nudging helps
base_plot2 +
  aes_label_some_countries +
  geom_text_s_favs(
    position = position_nudge_line(y = 6, x =2500, method = "spline")
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Nudging away from the cloud of points, increasing nudging and setting hjust
base_plot2 +
  aes_label_some_countries +
  geom_text_s(
    position = position_nudge_line(y = 6, x =2500, method = "spline"),
    hjust = "outward"
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Nudging away from the cloud of points, setting nudging direction to "split"
base_plot2 +
  aes_label_some_countries +
  geom_text_s(
    position = position_nudge_line(y = 4, x = 1000, method = "spline",
                                   direction = "split")
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Using 'ggpp' + 'ggrepel' ------------------------------------------------

# The examples above tend to work, but there is no guarantee of no overlaps,
# so combining them with repulsion is safer. The more dispersion there is
# in the data points the more necessary repulsion becomes.

# I set max.overlaps = 30 so that 'ggrepel' does not give up too early

# Nudging away based on a spline fitted to the cloud of points
base_plot2 +
  aes_label_some_countries +
  geom_text_repel_favs(
    position = position_nudge_line(y = 5, x = 3000, method = "spline"),
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Nudging away based on a spline fitted to the cloud of points with repulsion and hjust
# The effect of justification is different in the geoms from 'ggrepel' and 'ggpp'!
aes_label_some_countries_empty_string <- aes(label = ifelse(country %in% names(some.countries),
                       some.countries[as.character(country)],
                       ""))

base_plot2 +
  aes_label_some_countries_empty_string  +
  geom_text_repel_favs(
    position = position_nudge_line(y = 5, 
                                   x = 3000, method = "spline"),
    hjust = "outward" # <- added
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Nudging away from the cloud of points with nudging with direction away from
# fitted spline.
base_plot2 +
  aes_label_some_countries_empty_string  +
  geom_text_repel_favs(
    position = position_nudge_line(y = 5, x = 3000, method = "spline",
                                  direction = "split"), # <-- added
    hjust = "outward",
    max.overlaps = 50
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}
# Nudging away from the cloud of points with repulsion, a different option for
# direction.
base_plot2 +
    aes_label_some_countries_empty_string  +
  geom_text_repel_favs(

    position = position_nudge_line(y = 5, x = 3000, method = "spline",
                                   direction = "automatic"), # <-- added
    max.overlaps = 50,
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}

# Nudging away from the cloud of points with repulsion, adding a minimum
# distance from the fitted line for the nudging that is 1.5 times the
# nudging of observations.
base_plot2 +
    aes_label_some_countries_empty_string +
  geom_text_repel_favs(
    position = position_nudge_line(y = 5, x = 3000, method = "spline",
                                   direction = "automatic",
                                   line_nudge = 1.5),
    max.overlaps = 50
  )
#> Warning: Using alpha for a discrete variable is not advised.
```


```{r}
# Nudging away from the cloud of points with repulsion, adding a minimum
# distance from the fitted line that is 2 times the nudging for individual
# observations, while reducing the amount of nudging for observations.
base_plot2 +
  aes_label_some_countries_empty_string +
  geom_text_repel_favs(
    position = position_nudge_line(y = 3, x = 2000, method = "spline",
                                   direction = "split",
                                   line_nudge = 2),
    max.overlaps = 50,
    min.segment.length = 0,
  )
#> Warning: Using alpha for a discrete variable is not advised.
```




