---
title: 'Companion Guide: "Extending your Ability to Extend" by Thomas Lin Pederson'
author: "Evangeline Reynolds"
date: "3/7/2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 00: Press play on 'Extending your Ability to Extend' https://www.youtube.com/watch?v=uj7A3i2fi54

Step 0: Do it with ggplot2.

# Step 0: get it done with base ggplot2/tidyverse

```{r step0}
library(tidyverse)

create_circle <- function(n = 10){
  
    angles <- seq(from = 0,
                to = 2 * pi,
                length.out = n)

  data.frame(
    x = c(cos(angles)),
    y = c(sin(angles))
  )
  
}



five_cars %>% 
  mutate(which_car = row_number()) %>% 
  crossing(create_circle() %>%
             mutate(order = row_number())) %>% 
  arrange(which_car, order) %>% 
  # scaling
  mutate(x = speed * x) %>% 
  mutate(y = speed * y) %>%  
  # centering
  mutate(x = x + speed) %>%
  mutate(y = y + dist) %>%
  ungroup() ->
five_cars_circles
  
five_cars %>% 
  ggplot() + 
  aes(x = speed, dist) +
  geom_point() + 
  geom_polygon(data = five_cars_circles, 
            aes(x = x, y = y, group = which_car),
            alpha = .5) + 
  coord_equal()
```

# step 1a:  write `setup_data` function

```{r cars}



setup_data_circle <- function(data, params) {
  
    if (data$group[1] == -1) {
      nrows <- nrow(data)
      data$group <- seq_len(nrows)
    }
  
    data  # return data with a group variable

}
```

# step 1b: test `setup_data` data function

```{r step1b}
cars %>% 
  slice(1:5) %>% 
  mutate(group = -1) %>% # no grouping neg one is default in ggplot2
  setup_data_circle() # setup makes each row defines a group

cars %>% 
  slice(5:20) %>% 
  mutate(group = 2) %>% # if a group is already defined
  setup_data_circle() # setup data does not do anything
```

---

# step 2a: write a `compute_group` function

We write a routine that will act on each group in the data (in this case each row)

---

```{r step_2a}
compute_panel_circles <- function(data, scales, n = 5){

  create_circle <- function(n = 10){
  
    angles <- seq(from = 0,
                to = 2 * pi,
                length.out = n)

  data.frame(
    x = c(cos(angles)),
    y = c(sin(angles))
  )
  
}



data %>% 
  # mutate(which_car = row_number()) %>% 
  crossing(create_circle() %>%
             mutate(order = row_number())) %>% 
  arrange(order) %>% 
  # scaling
  mutate(x = r * x) %>% 
  mutate(y = r * y) %>%  
  # centering
  mutate(x = x + x0) %>%
  mutate(y = y + y0)
  
  
}




# compute_panel_circles <- function(data, scales, n = 20){
# 
# data %>% 
#   mutate(diamond = 
#            purrr::map(.x = r, 
#                        create_circle)) %>% 
#   mutate(my_group = row_number()) %>% 
#   unnest() %>% 
#   mutate(x = x + x0) %>% 
#   mutate(y = y + y0)
# 
# }

```


# step 2b: test `compute_panel` row processing function

```{r 2a2}
mtcars %>%
  select(disp, mpg, cyl) %>% 
  rename(x0 = disp, y0 = mpg, r = cyl) %>% 
  compute_panel_circles(n = 6)

cars %>%
  rename(x0 = dist, y0 = speed) %>% 
  mutate(r = x0) %>%
  .[5,] %>% 
  compute_panel_circle(n = 6) %>% 
  ggplot() + 
  aes(x = x, y = y) +
  geom_polygon(alpha = .5) + 
  coord_equal()

```



# Step 3: write ggproto to create StatCircle; `setup_data` and `compute_group` functions will be inputs

```{r 3}
StatCircle <- ggproto(`_class` = "StatCircle", 
                      `_inherit` = Stat, 
                      # setup_data = setup_data_circle,
                      compute_panel = compute_panel_circles,
                      required_aes = c("x0", "y0", "r")
                      )
```



# Step 4: write `geom_circle`, inheriting from `GeomPolygon`

```{r 4}
geom_circle <- function(mapping = NULL, 
                        data = NULL, 
                        stat = "circle", 
                        position = "identity", 
                        ..., 
                        n = 8, 
                        lineend = "butt", 
                        linejoin = "round", 
                        na.rm = FALSE, 
                        show.legend = NA, 
                        inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPolygon, 
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(
      n = n, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}
```

# Step 5: Enjoy! Test out `geom_circle`

```{r 5}
test_df <- data.frame(
  
  x0 = c(-5, 5),
  y0 = c(5, -5),
  r = c(5, 4),
  class = c("a", "b")
  
)


cars %>% 
  slice(1:5) %>% 
  ggplot() + 
  aes(x = speed, y = dist) +
  geom_point() +
  aes(x0 = speed, 
      y0 = dist, 
      r = speed/6) + 
  geom_circle() +
  coord_equal() 
```

# Keep testing; second guess everything; dispare

```{r}
baseplot + 
  aes(fill = speed == 6) +
  geom_circle(n = 7, alpha = .2) + 
  facet_wrap(~class)
```


# Look at `geom_circle` from ggforce; hope renewed

```{r}
baseplot + 
  aes(fill = speed > 6) +
  ggforce::geom_circle(n = 5)
```



# Step 6: Post Mordem using `layer_data` to look at the data frame


```{r 6}
baseplot +
  geom_circle(n = 7, alpha = .2) ->
p1

layer_data(p1, 2)

baseplot + 
  aes(fill = speed > 6) +
  geom_circle(n = 7, alpha = .2) ->
p2

layer_data(p2, 2)
```


