---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      error = T)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)

```

## Experiment

```{r}
library(ggplot2)
data_filter <- function(.keep, .by) {
  structure(list(keep_specification = rlang::enquo(.keep), 
                 by_specification = rlang::enquo(.by)), 
            class = "filterobs")
}

ggplot_add.filterobs <- function(object, plot, object_name) {
  
  new_data <- dplyr::filter(plot$data, 
                            !!object$keep_specification, 
                            .by = !!object$by_specification)
  plot$data <- new_data
  plot

}

mtcars
plot(mtcars)

ggplot(mtcars) + 
  aes(factor(cyl), fill = factor(cyl)) + 
  geom_bar() -> p1;p1

ggplot(mtcars) + 
  aes(id = cyl, fill = factor(cyl)) + 
  ggcirclepack::geom_circlepack() + 
  coord_equal() -> p2

ggplot(mtcars) + 
  aes(wt, mpg, color = factor(cyl)) + 
  geom_point(size = 4) -> p3
  
library(patchwork)
p1 + p2 + p3

(p1 + p2 + p3) * data_filter(cyl != 4)

(p1 + p2 + p3) * aes(alpha = cyl != 4)



```


However, w/ `/` some repetition is required.

```{r}

(p1 + p2) * aes(alpha = cyl != 4) / p3 * aes(alpha = cyl != 4)


(p1 + p2) * data_filter(cyl != 4)  / p3 * data_filter(cyl != 4)


p1 * data_filter(cyl != 4)  / ((p2 + p3) * data_filter(cyl != 4))

```


This can be shorted by naming specification.


```{r}
cyl4spotlight <- aes(alpha = cyl == 4)

(p1 + p2) * cyl4spotlight / p3 * cyl4spotlight

rm4 <- data_filter(cyl != 4)

(p1 + p2) * rm4 / p3 * rm4


```


There is an `&` operator, that is demonstrated with 'theme' in the documentation.

> & will add the element to all subplots in the patchwork, and * will add the element to all the subplots in the current nesting level. As with | and /, be aware that operator precedence must be kept in mind.

Which sounds promising, but isn't working with data_filter() as written but *does* work with the opaqueness spotlighting. Exciting!

```{r}
patchwork <- p3 / (p1 | p2)
patchwork & theme_minimal()

patchwork & aes(alpha = cyl == 4) & guides(alpha = "none")
patchwork & data_filter(cyl != 4)
```


```{r}
file.info("programmatic-crosstalk.Rmd")$ctime -
  Sys.time()
```

## Closing remarks, Other Relevant Work, Caveats
