---
title: "penguinseval"
subtitle: ""
author: ""
output:
  xaringan::moon_reader:
    seal: FALSE
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## ggplot_build() steps revealed

---

```{r, fig.height=3}
library(ggplot2)
set.seed(1234)
ggplot(palmerpenguins::penguins |> 
         dplyr::sample_n(10), aes(sex)) +
  stat_count(position = "stack") +
  facet_wrap(~species) + 
  aes(fill = island) ->
p;p
```

---

```{r}
# ggplot2::ggplot_build()
# function (plot) 
# {
  plot <- ggplot2:::plot_clone(p); 
plot |> str() # step 1.```
```
    
---
    
```{r}
  if (length(plot$layers) == 0) {  plot <- plot + geom_blank()}; 
plot |> str() # step 2.
```

---
    
```{r}
  layers <- plot$layers; 
layers # step 3.
```

---
    
```{r}
  data <- rep(list(NULL), length(layers)); 
data # step 4.
```

---
    
```{r}
  scales <- plot$scales; 
data # step 1.
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$layer_data(plot$data), 
                             layers, data, "computing layer data"); data # step 5.
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$setup_layer(d, plot), 
                             layers, data, "setting up layer"); data # step 6.
```

---
    
```{r}
  layout <- ggplot2:::create_layout(plot$facet, 
                                    plot$coordinates, plot$layout); data # step 7.
```

---
    
```{r}
  data <- layout$setup(data, plot$data, plot$plot_env); 
data # step 8.
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$compute_aesthetics(d, plot), 
                             layers, data, "computing aesthetics"); data # step 9.
```

---
    
```{r}
  data <- .ignore_data(data); 
data # step 10.
```

---
    
```{r}
  data <- lapply(data, scales$transform_df); 
data  # step 11
```

---
    
```{r}
  scale_x <- function() scales$get_scales("x"); 
data # step 12
```

---
    
```{r}
  scale_y <- function() scales$get_scales("y"); 
data # step 13
```

---
    
```{r}
  layout$train_position(data, scale_x(), scale_y()); 
data # step 14
```

---
    
```{r}
  data <- layout$map_position(data); 
data # step 15
```

---
    
```{r}
  data <- .expose_data(data); 
data # step 16
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$compute_statistic(d, layout), 
                             layers, data, "computing stat"); data # step 17
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$map_statistic(d, plot), 
                             layers, data, "mapping stat to aesthetics"); data # step 18
```

---
    
```{r}
  plot$scales$add_missing(c("x", "y"), plot$plot_env); 
data # step 19
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$compute_geom_1(d), 
                             layers,  data, "setting up geom"); data # step 20
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$compute_position(d, layout),  
                             layers, data, "computing position"); data # step 21
```

---
    
```{r}
  data <- .ignore_data(data); 
data # step 22 
```

---
    
```{r}
  layout$reset_scales(); 
data # step 23
```

---
    
```{r}
  layout$train_position(data, scale_x(), scale_y()); 
data #layout # step 24
```

---
    
```{r}
  layout$setup_panel_params(); 
data #layout # step 25
```

---
    
```{r}
  data <- layout$map_position(data); 
data # step 26
```

---
    
```{r}
  layout$setup_panel_guides(plot$guides, plot$layers); 
data #layout # step 27
```

---
    
```{r}
  npscales <- scales$non_position_scales(); 
data #npscales # step 28

```  

---

    
```{r}
  if (npscales$n() > 0) {lapply(data, npscales$train_df); plot$guides <- plot$guides$build(npscales, plot$layers,  plot$labels, data);     data <- lapply(data, npscales$map_df)}else { plot$guides <- plot$guides$get_custom() };
data# plot$guides  step 29
```

---
    
```{r}
  data <- .expose_data(data); 
data # step 30
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$compute_geom_2(d), layers,  
                             data, "setting up geom aesthetics"); data # step 31
```

---
    
```{r}
  data <- ggplot2:::by_layer(function(l, d) l$finish_statistics(d),  layers, 
                             data, "finishing layer stat"); data # step 32
```

---
    
```{r}
  data <- layout$finish_data(data); 
data # step 33
```

---
    
```{r}
  plot$labels$alt <- get_alt_text(plot); 
data # step 34
```

---
    
```{r}
  structure(list(data = data, layout = layout, plot = plot),   class = "ggplot_built"); 
data # step 35
```





## Closing remarks, Other Relevant Work, Caveats
