---
title: "return to dim red after a while"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```


2026-06-28 @corybrunson I've mentioned that I wanted to share with you on this, and I've thought about rehashing and emailing or putting on discussions but on my walk today, I felt like @ - ing might be a less resource intense way.  So just doing that. 🤷‍♀️  Can @ be pushed to github?  As I am doing?  Another way?  

Here, exploring a pivot to preprocessing or different in-processing features() instead.  

## Intro Thoughts

ggdims feels a little more engineered than it might be. See https://github.com/EvaMaeRey/ggdims.  So  Let's look at a slightly different approach:

Target:

```{r, eval = F}
data |> 
  make_dims_column() |>
  ggplot() + 
  aes(dims = .dims_column) + # list column... each row has unique data for the 
  geom_pca()
```


## Experiment

```{r}
library(tidyverse)
declare_features <- function(data, ...){

data$features <- list(NA, length(data))

features_df <- data |> select(...)
# names(features_df) <- paste0(".f", 1:ncol(features_df))

for (i in 1:nrow(data)){
  
  data$features[[i]] <- features_df |> slice(i)
  
}             

data |> as_tibble()

}


iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) 

iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |>
  ggplot() + 
  aes(features = features) + ## do features() or do geometry = geometry...
  geom_blank()
```

# make it interact with compute via a stat

```{r}
compute_panel_pca <- function(data, scales){
  
  dims <- data |> select(features) |> unnest(cols = features)
    
  pca_ed <- dims |>
    ordr::ordinate(model = ~ prcomp(., scale. = TRUE)) |> 
    _[[5]] |> 
    as_tibble()
  
  pca_ed |> 
    bind_cols(data)
  
}

iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |> 
  select(features) |>
  compute_panel_pca() 


library(statexpress)
iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |> 
  ggplot() + 
  aes(features = features) + 
  qlayer(stat = qstat_panel(compute_panel_pca,
                      default_aes = aes(x = after_stat(PC1), 
                                        y = after_stat(PC2)))) + 
  aes(color = Species)


p <- last_plot()

features <- function(){
  
  aes(features = features)
  
}


iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |> 
  ggplot() + 
  features() + 
  qlayer(stat = qstat_panel(compute_panel_pca,
                      default_aes = aes(x = after_stat(PC1), 
                                        y = after_stat(PC2)))) + 
  aes(color = Species)

```


# write a proper Stat and user facing function...

```{r}
StatPCA <- ggproto(`_class` = "StatPCA", `_inherit` = Stat,
                   compute_panel = compute_panel_pca,
                   required_aes = "features",
                   default_aes = aes(x = after_stat(PC1), 
                                        y = after_stat(PC2)))


geom_pca0 <- make_constructor(GeomPoint, stat = StatPCA)

iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |> 
  ggplot() + 
  aes(features = features) +
  geom_pca0()

last_plot() +
  aes(color = Species)
```

# automatically look for the column like geom_sf looks for geometry (how does that managed?)

```{r}
geom_pca <- function(...){
  
  list(
  geom_pca0(...),
  aes(features = features)
  )
  
}

iris |> 
  declare_features(Sepal.Length:Petal.Length, Petal.Width) |> 
  ggplot() + 
  geom_pca()

last_plot() + 
  aes(color = Species)
```


## Preferred ... wish list:  add a `features()` function that can be declared within - fails with selections with commas (other tidy select moves?)

```{r, eval = T, error = T}
#' @export
features0 <- function(select0) {

  structure(
    list(
      select_spec = rlang::enquo(select0)
    ), 
    class = "features0"
    )

}


#' @import ggplot2
#' @importFrom ggplot2 ggplot_add
#' @export
ggplot_add.features0 <- function(object, plot, object_name) {
  
  plot$data <- plot$data |> 
    declare_features(!!object$select_spec)

  plot <- plot + aes(features = features)

  plot

}


features <- function(select){
  
  features0({{select}})
  
}


library(statexpress)
iris |> 
  ggplot() + 
  features(Sepal.Length:Petal.Length) + 
  geom_pca0() + 
  aes(color = Species)




library(statexpress)
iris |> 
  ggplot() + 
  features(Sepal.Length:Petal.Length, Petal.Width) + 
  geom_pca0() + 
  aes(color = Species)


```



# not preferred, though I'm having trouble articulating why...




```{r}
ggdims <- function(data, mapping, ...){
  
  
  
  
  
}




```


# a little venn exploration

```{r}
library(ggVennDiagram)
y <- list(
  A = sample(letters, 8),
  B = sample(letters, 8),
  C = sample(letters, 8),
  D = sample(letters, 8)
) 

as.data.frame(y)

ggVennDiagram(y)

process_region_data(Venn(y))


genes <- paste("gene",1:1000,sep="")
set.seed(20231214)
x <- list(A=sample(genes,300),
          B=sample(genes,525),
          C=sample(genes,440),
          D=sample(genes,350))


ggVennDiagram


ggVennDiagram::plot_venn



compute_venn_list <- function(data){
  
  df <- data |>
    select(features) |> 
    unnest(features) 
  
  venn_list <- list(NA)
  # names(venn_list) <- names(df)
    
  for(i in 1:ncol(df)){
    
    venn_list[[i]] <- (1:nrow(df))[df[,i] |> pull() |> as.logical()]
    
  }
  
  names(venn_list) <-  names(df)

  venn_list
  
  }


# raw TF data
recipes_characteristics_logical <- data.frame(
  tasty = sample(0:1, 100, replace = T), 
  easy = sample(0:1, 100, replace = T)) |>
  mutate(id = paste("item", row_number()), .before = 1)


recipes_characteristics_logical |> 
  # declare features and select only features
  declare_features(tasty, easy) |> 
  select(features) |>
  # compute that happens internally in an geom_venn()
  compute_venn_list() |> 
  ggVennDiagram() + 
  labs(title = "Recipe characteristics: What percent are easy and tasty?")

ggVennDiagram:::plot_shapes()

# data |> 
#   declare_features |> 
#   ggplot() + 
#     aes(features = features) + 
#     geom_venn()


```




## Closing remarks, Other Relevant Work, Caveats
