---
title: "ggtrace_highjack_return"
subtitle: "A flipbook treatment"
author: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 12, fig.height = 10, message = FALSE, 
                      warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
set.seed(1293812)
theme_set(new = theme_gray(base_size = 15))
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Usage

```{r, eval = F}
ggtrace_highjack_return(
  
  x, 
  # A ggplot object
  
  method, 
  # A function or a ggproto method. The ggproto method may be specified 
  # using any of the following forms: ggproto$method, namespace::ggproto$method,
  # namespace:::ggproto$method
  
  cond = 1L, 
  # When the return value should be replaced. Defaults to 1L.
  
  value = quote(returnValue()), 
  # What the method should return instead. Defaults to quote(returnValue()).

  draw = TRUE 
  # Whether to draw the modified graphical output from evaluating x. 
  # Defaults to TRUE.

)
```







---

`r chunk_reveal("start", break_type = 1, widths = c(2,3), title = "# base plot to highjack")`

```{r start, include = F}
library(tidyverse)
library(ggtrace)
p1 <- ggplot(diamonds, aes(cut)) +
  geom_bar(aes(fill = cut)) +
  facet_wrap(~ clarity, ncol = 2)

p1
```




---

`r chunk_reveal("start1", widths = c(2,3), break_type = 1, title = "# first panel to zero")`

```{r start1, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel, #<<
  value = quote({ #<<
    returnValue() %>% #<<
      mutate(count = 0) #<<
  })
)
```


---

`r chunk_reveal("start12", widths = c(2,3), break_type = 1, title = "# first panel times 20")`

```{r start12, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel,
  value = quote({
    returnValue() %>%
      mutate(count = count*20) #<<
  })
)
```


---

`r chunk_reveal("start11", widths = c(2,3), break_type = 1, title = "# back to original")`

```{r start11, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel,
  value = quote({
    returnValue() %>%
      mutate(count = count) #<<
  })
)
```


---

`r chunk_reveal("start2", widths = c(2,3), break_type = 1, title = "# mix up 4th panel x positions?")`

```{r start2, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel,
  cond = quote(data$PANEL[1] == 4), #<<
  value = quote({
    returnValue() %>%
      mutate(x = sample(x)) #<<
  })
)
```


---

`r chunk_reveal("start27", widths = c(2,3), break_type = 1, title = "# mix up 4th panel counts?")`

```{r start27, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel,
  cond = quote(data$PANEL[1] == 4), #<<
  value = quote({
    returnValue() %>%
      mutate(count = sample(count)) #<<
  })
)
```

---

`r chunk_reveal("start111", widths = c(2,3), break_type = 1, title = "# back to original")`

```{r start111, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = Stat$compute_panel,
  cond = quote(data$PANEL[1] == 4), 
  value = quote({
    returnValue() %>%
      mutate(count = count) #<<
  })
)
```


---

`r chunk_reveal("start3", widths = c(2,3), break_type = 1, title = "# very special highlighting for panel 4")`

```{r start3, include = F}
ggtrace_highjack_return(
  x = p1, 
  method = GeomBar$draw_panel, #<<
  cond = quote(data$PANEL[1] == 4),
  value = quote({
    editGrob( #<<
      returnValue(),
      gp = gpar( #<<
        col = "black",  #<<
        alpha = seq(0.2, 1,  #<<
                    length.out =  #<<
                      nrow(data) #<<
    )))
  })
)
```


