---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts

Below, I'm copying and pasting a first cut at ggpie - one that mostly just borrows from geom_bar and the recommended approach in the ggplot2 documentation.

But there is a point of friction in the y or x needs to be specified to build the pie.  Which doesn't really seem like relevant to building a pie chart. 

The bigger issue is be that overall area of the pie could be relevant, and this, I believe is not possible with geom_bar/stat_count.  So it seems worthwhile to try to build a custom geom.  


---





# ggpie  🦄 (is mostly mythical), probably changing to ggwedge

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of ggpie is to allow this interface:

```{r, eval = T}
library(ggplot2)
library(ggpie)

ggpie(diamonds) + 
  aes(fill = color) + 
  geom_wedge() 

last_plot() + 
  facet_wrap(facets = vars(cut))
```


Observations:  the `ggplot()` function start point has a particular set of defaults that might not be the best suited for final plot.

Assertion: other defaults can be bundled up and serve as grammatically-equivalent alternative start points.


### Alternative without ggpie

```{r}
library(tidyverse)

diamonds %>%
ggplot() + 
  aes(y = 0, fill = cut) + 
  geom_bar(position = "fill") + 
  coord_polar() + 
  theme_void() 
```

#### adding labels

```{r, error=T}
last_plot() + 
 stat_count(
    geom = "text", position = "fill",
    color = "white", size = 8,
    aes(x = .25,
        label = after_stat(
          count*100/sum(count)) |>
          round(1) |>
          paste0("%"),
        y = stage(start = NULL,
                  after_stat = count,
                  after_scale = (ymin +
                                   ymax)/2)))
  


```

#### case where you already have summary on which to proportion...


```{r}
library(tidyverse)

diamonds %>%
  count(cut) %>% 
  ggplot() + 
  aes(x = 0, fill = cut, weight = n) + 
  geom_bar(position = "fill") + 
  coord_polar(theta = "y") + 
  theme_void() 
```


# Developing the new API.

```{r, pie_functions}
#' Title
#'
#' @return
#' @export
#'
#' @examples
plot_pie <- function(...){
  
  list(
    ggplot2::coord_polar(theta = "y", ...),
    ggplot2::theme_void(),
    ggplot2::aes(x = 0) # hacky; grammar problem
  )
  
}

#' Title
#'
#' @param data 
#'
#' @return
#' @export
#'
#' @examples
ggpie <- function(data){
  
  ggplot2::ggplot(data = data) + 
  plot_pie()
  
}


# just aliasing to be nice to ourselves
# probably a better way 
# is doing more re-writing so that x is not a required aesthetic
#' Title
#'
#' @param ... 
#'
#' @return
#' @export
#'
#' @examples
geom_wedge <- function(...){
  
  ggplot2::geom_bar(position = "fill", ...)
  
}


# some very preliminary and messy ideas for new Stat Wedge approach
# but wondering if this is worth it.  
# I think it could be if geom is used with ggplot() start point.
# StatWedge <- ggproto(`_class` = StatCount2, 
#                      `_inherit` = ggplot2::Stat,
#                      compute_group )
# 
# ggplot2::StatCount$compute_group %>% 
#   mutate(x)
# 
# stat_count

# geom_wedge <- function (mapping = NULL, data = NULL, geom = "bar", 
#                         position = "fill", 
#                         ..., width = NULL, na.rm = FALSE, 
#                         orientation = NA, show.legend = NA, 
#                         inherit.aes = TRUE) 
# {
#     params <- list2(na.rm = na.rm, orientation = orientation, 
#         width = width, ...)
#     layer(data = data, mapping = mapping, stat = StatCount, geom = geom, 
#         position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
#         params = params)
# }

```


## Test it out


```{r}
ggpie(diamonds) + 
  aes(fill = cut) +
  geom_wedge() 

# a ggdonut() function could also be written
ggpie(diamonds) + 
  aes(fill = cut) +
  geom_wedge() +
  xlim(-2, 1)
  
```
