---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)
```

## Experiment

```{r}
library(tmap)
data(NLD_prov)
data(NLD_muni)
```


## Create reference data. 

Now we're looking at a dependency free method (w/o sf2stat), it's a bit of code. 

I believe this is a work around and it's just a matter of time before figuring out how to get this more automated using what's alread in ggplot2.

```{r}
# lots of work to build sf2stat
# NLD_prov |>
#   dplyr::select(name_prov = name, code_prov = code, geometry) |>
#   sf2stat:::sf_df_prep_for_stat(id_col_name = "name_prov") ->
# nld_prov_geo_reference0

# a fun moment
# # build a plot with sf type data
# NLD_prov %>% 
#   ggplot() + 
#   geom_sf() + 
#   geom_sf_text(aes(label = name))
# 
# # extract xmin ymin, x y, 
# NLD_prov %>% 
#   select(name_prov = name, code_prov = code, geometry) %>%
#   bind_cols(layer_data() %>% select(xmin, xmax, ymin, ymax)) %>% 
#   bind_cols(layer_data(i = 2) %>% select(x, y)) %>% 
#   as.data.frame() %>% 
#   mutate(id_col = name_prov) ->
# nld_prov_geo_reference

# # the way - but now I put it below
# NLD_prov %>% 
#   StatSf$compute_panel(coord = CoordSf) ->
# 
# 
# NLD_prov %>% 
#   StatSfCoordinates$compute_group(coord = CoordSf)


NLD_prov %>% 
   select(name_prov = name, code_prov = code, geometry) %>% 
   mutate(id_col = name_prov)  %>% 
    StatSf$compute_panel(coord = CoordSf) %>%  # adds individual bounding boxes, xmin, ymin etc
    StatSfCoordinates$compute_group(coord = CoordSf) %>% # adds xy coordinates
  tibble() -> # remove sf info
nld_prov_geo_reference


```

# write compute routine

```{r}
compute_panel_nl_prov <- function(data, scales, keep_id = NULL, drop_id = NULL, stamp = FALSE){
  
  if(!stamp){data <- dplyr::inner_join(data, nld_prov_geo_reference)}
  if( stamp){data <- nld_prov_geo_reference }
  
  if(!is.null(keep_id)){ data <- filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- filter(data, !(id_col %in% drop_id)) }
  
  data
}
```

# write stat

```{r}
StatNlprov <- ggplot2::ggproto(`_class` = "StatNlprov",
                                `_inherit` = ggplot2::StatSf,
                               required_aes = "name_prov|code_prov",
                               compute_panel = compute_panel_nl_prov,
                               default_aes = 
                                 ggplot2::aes(label = after_stat(id_col)))
```

# try out the stat with geom_sf

```{r}
NLD_prov |>
  sf::st_drop_geometry() |>
  slice(1:3) |>
  ggplot() + 
  aes(name_prov = name) + 
  geom_sf(stat = StatNlprov, stamp = T) + 
  geom_sf(stat = StatNlprov, fill = "darkred") +
  coord_sf(crs = "EPSG:28992") + 
  geom_sf_label(stat = StatNlprov, keep_id = "Groningen")
```


# write user-facing wrappers

```{r}
geom_prov <- function(...){
  
  list(
  geom_sf(stat = StatNlprov, ...),
  coord_sf(crs = "EPSG:28992")
  )
  
}


geom_prov_text <- function(...){
  
  list(
  geom_sf_text(stat = StatNlprov, ...),
  coord_sf(crs = "EPSG:28992")
  )
  
}
```


# use user-facers.

```{r}
NLD_prov |>
  sf::st_drop_geometry() |> # so we can test on flat file (no geometry )
  ggplot() + 
  aes(name_prov = name) + 
  geom_prov() + 
  geom_prov_text(size = 2) 

last_plot() + 
  aes(fill = population) + 
  scale_fill_viridis_c()
  
```



```{r}
NLD_prov |>
  sf::st_drop_geometry() |>
  ggplot() + 
  aes(name_prov = name) + 
  geom_prov(stamp = T) + 
  geom_prov_text(size = 2) 


NLD_prov |>
  sf::st_drop_geometry() |>
  ggplot() + 
  aes(name_prov = name) + 
  geom_prov(stamp = T) + 
  geom_prov_text(size = 5, keep_id = "Groningen") 

```

# ho! new reality in 30 lines of code...

```{r}
NLD_prov |>
  sf::st_drop_geometry() |>
  ggplot() + 
  aes(name_prov = name) +
  geom_prov() + 
  geom_prov_text(check_overlap = T)

last_plot() + 
  aes(fill = pop_65plus)




```



```{r}
knitr::knit_exit()


```

## Closing remarks, Other Relevant Work, Caveats



```{r, eval = F}
library(geobr)
library(tidyverse)


br_state_raw <- geobr::read_state(simplified = T)
str(br_state_raw)

state <- geobr::read_state(simplified = T) %>% 
  rename(geometry = geom) %>% 
  sf2stat:::sf_df_prep_for_stat(id_col_name = "name_state")


compute_panel_state <- function(data, scales){
  
  data <- inner_join(data, state)
  
  data
  
}


sf::st_crs( geobr::read_state(simplified = T))


state %>% 
  sf::st_drop_geometry() %>% 
  compute_panel_state()

StatBrasilstate <- ggproto("StatBrasilstate", 
                           Stat,
                           compute_panel = compute_panel_state)


state %>% 
  ggplot() +
  aes(geometry = geometry) +  # why is it needed here
  geom_sf() + 
  geom_sf_text(aes(label = name_state))


state %>% 
  sf::st_drop_geometry() %>% 
  ggplot() + 
  aes(code_state = code_state) +
  geom_sf(stat = StatBrasilstate) + 
  coord_sf(crs = "SIRGAS2000")

```

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE) %>% 
  sf2stat:::sf_df_prep_for_stat(id_col_name = "NAME")


compute_panel_nc <- function(data, scales){
  
  nc %>% inner_join(data)

}


StatNc <- ggproto("StatNc", 
                  Stat,
                  compute_panel = compute_panel_nc,
                  default_aes = aes(geometry = after_stat(geometry)))

nc %>% 
  sf::st_drop_geometry() %>% 
  ggplot() + 
  aes(FIPS = FIPS) + 
  geom_sf(stat = StatNc) + 
  coord_sf(crs = "NAD27") 
  
                  
nc %>% 
  sf::st_drop_geometry() %>% 
  slice(1:3) %>% 
  ggplot() + 
  aes(FIPS = FIPS) + 
  geom_sf(stat = StatNc)

```
