---
title: "after_stat, after_scale"
subtitle: "A flipbook and layer_data treatment"
author: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = FALSE, 
                      warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
theme_set(new = theme_gray(base_size = 15))
```







---

`r chunk_reveal("start", widths = c(2,3))`

```{r start, include = F}
ggplot(faithful, aes(x = waiting)) +
  geom_rug() + 
  geom_histogram(bins = 12)



layer_data(last_plot(), 2) %>% 
  .[1:6] 
```


---


`r chunk_reveal("after", break_type = 1, widths = c(2,3))`


```{r after, include = F}
ggplot(faithful, aes(x = waiting)) +
  geom_rug() + 
  geom_histogram(bins = 12, 
    aes(y = after_stat(density))
    )

layer_data(last_plot(), 2) %>% 
  .[1:6] 
```


---


`r chunk_reveal("after2", break_type = 1, widths = c(2,3))`


```{r after2, include = F}
ggplot(faithful, aes(x = waiting)) +
  geom_rug() + 
  geom_histogram(bins = 12, 
    aes(y = ..density..)
    )

layer_data(last_plot(), 2) %>% 
  .[1:6] 
```


---



```{r}
# faithful %>% 
#   rename(x = waiting, y) 
```


---

`r chunk_reveal("ok", break_type = 1, widths = c(2,2))`


```{r ok, include = F}
# The exact colour is known after scale transformation
ggplot(mpg, 
       aes(cty, colour = factor(cyl))) +
  geom_density()

layer_data(last_plot(), 1)[1:5, 1:3]

```

---

`r chunk_reveal("ok2", break_type = 1, widths = c(2,2))`


```{r ok2, include = F}
# We re-use colour properties for the fill without a separate fill scale
ggplot(mpg, 
       aes(cty, colour = factor(cyl))) +
  geom_density(aes(
    fill = after_scale(alpha(colour, 0.3))
    ))

layer_data(last_plot(), 1)[1:5, 1:3]
```

---

`r chunk_reveal("ok222", break_type = 1, widths = c(2,2))`


```{r ok222, include = F}
# We re-use colour properties for the fill without a separate fill scale
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot(
    aes(fill = class)
    )

layer_data(last_plot(), 1)[1:5, 1:3]
```


---

`r chunk_reveal("ok223", break_type = 1, widths = c(2,2))`


```{r ok223, include = F}
# We re-use colour properties for the fill without a separate fill scale
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot(
    aes(fill = class),
    alpha = 0.4
    )

layer_data(last_plot(), 1)[1:5, 1:3]
```

---

`r chunk_reveal("ok300", break_type = 1, widths = c(2,2))`


```{r ok300, include = F}
# The exact colour is known after scale transformation
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot(
    aes(fill = 
          stage(class,
                after_scale = 
                  alpha(fill, 0.4)))
    )


layer_data(last_plot(), 1)[1:5, 1:3]

```


---

`r chunk_reveal("ok38", break_type = 1, widths = c(2,2))`


```{r ok38, include = F}
# The exact colour is known after scale transformation
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot(
    aes(fill = 
          stage(class,
                after_scale = 
                  alpha(fill, 0.4)))
    )


layer_data(last_plot(), 1)[1:5, 1:3]

```



---

`r chunk_reveal("ok232", break_type = 1, widths = c(2,2))`


```{r ok232, include = F}
# We re-use colour properties for the fill without a separate fill scale
ggplot(mpg) +
  aes(x = class, 
      y = displ) +
  geom_violin() + 
  stat_summary(
    # layer is text
    geom = "text", 
    # compute variables on-the-fly  
    fun.data = 
      ~ round(data.frame(my_mean = mean(.x),
                         my_sd = sd(.x)), 2),
    # use those vars in plot
    mapping = 
      aes(y = stage(displ, after_stat = 8),
          label = after_stat(
            paste(my_mean, "±", my_sd)
            )
          )
  )

layer_data(last_plot(), 2)[1:5, 1:7]
```





Thanks for joining!  

This flipbook will look at the topic of fine-tuned aesthetic evaluation using after_stat, after_scale, and stage.  

Are you the type of person that learned ggplot2 before these mechanisms were put together, and just ignore them (that's me!) 

--

This book is for you!

---

Are you the kind of person that can copy and paste after_stat, after_scale, and stage code but really has no idea what's going on? (that'd be me if I actually used the mechanisms)

--

This book is for you!

---

Are you the kind of person that admires really beautifully and thoughtfully crafted data vizualizations, and have noticed the authors use of after_stat, after_scale, and stage, but have thought, I'm not sure it's worth all the fuss.

--

This book is for you!





The source material is basically an expansion of https://ggplot2.tidyverse.org/reference/aes_eval.html


