---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r, warning=F, message=F}
library(tidyverse)
library(sportyR)

sportyR::geom_baseball("mlb", display_range = "full")

```

From sportyR field plot:
 
- grab plot layers 
- pull out x,y data from alll layer
- convert to sf (simple features)
- add a name to each ‘region’ (or line) - somewhat ad hoc, and I select baseball example because it feels relatively familiar and there are fewer elements.

Having done this, use ggregions framework:

- declare the dataset to be the region of interest with set_regions()
- Use aes(regions = my_region_id) and geom_region() and friends (e.g. stamp_region, stamp_region_label) to plot data and context.

```{r}
get_layer_data <- function(layer){

    layer$data
  
}

to_sf_routine <- function(data){
  
  data |>
  mutate(y = -y) |>
  sf::st_as_sf(coords = c("x", "y"), agr = "constant") |>
  group_by(id, group) |>
  summarize(do_union = F) |> 
  ungroup() |> 
  group_by(id, group) |>
  summarise() |>
  mutate(geometry = geometry |> sf::st_cast("POLYGON")) |> 
  mutate(geometry = geometry |> sf::st_cast("MULTIPOLYGON")) |> 
  ungroup() 
  
}



out <- geom_baseball("mlb", display_range = "full")


out@layers |> 
  tibble(layers = _) |> 
  mutate(name = names(layers)) |>
  mutate(data = purrr::map(layers, .f = get_layer_data)) |> 
  unnest(data) |> 
  mutate(id = name, group = name) |>
  to_sf_routine() |> 
  mutate(name = c("infeild", 
                  "batting right",
                  "catcher's box",
                  "left foul line",
                  "right foul line",
                  "dugout",
                  "diamond",
                  "pitcher's mound",
                  "batter's mound",
                  "first base",
                  "second base",
                  "third base",
                  "pitcher's line",
                  "batting left")) |> 
  select(name, geometry) ->
baseball_reference

library(ggregions)
set_regions(baseball_reference)
head(baseball_reference)
```


```{r, warning=F, message=F}
tribble(~zone, ~info,
        "right foul line", "good",
        "infeild", "bad",
        "diamond", "ugly",
        "pitcher's mound", "fine") |>
  ggplot() +
  aes(region = zone,
      color = zone) + 
  stamp_region() + 
  geom_region_border() + 
  stamp_region_label(keep = "left foul line")
```


```{r}

library(gganatogram)
gganatogram(data=other_key[["bos_taurus"]], outline = T, fillOutline='white', organism="bos_taurus", sex='female', fill="colour")  +
        theme_void() +
        ggtitle("bos_taurus") + 
        theme(plot.title = element_text(hjust=0.5)) + 
        coord_fixed()

organs <- other_key[["bos_taurus"]]$organ

out <- last_plot()


out@layers |> 
  tibble(layers = _) |> 
  mutate(name = names(layers)) |>
  mutate(data = purrr::map(layers, .f = get_layer_data)) |> 
  slice(-1) |>
  unnest(data) |> 
  mutate(id = name, group = group) |>
  to_sf_routine() |> 
  mutate(organ = c("hide", "head", "liver",
                  "semicircle", "foot",
                  NA, "foot2", "underbelly",
                  "sliver", "arc","u-shape",
                  "below", "13", 14:64 )) ->
cow_sf


set_regions(cow_sf)

cow_sf |>
  head() |> 
  sf::st_drop_geometry() |>
  ggplot() + 
  aes(region = organ,
      fill = organ) +
  stamp_region() +
  geom_region()




```



## Closing remarks, Other Relevant Work, Caveats
