---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
options(tidyverse.quiet = TRUE)
```

```{r}
library(tidyverse)
```

## Status Quo:  "A pie chart is just a bar chart with polar coordinates!*" 

*But you do need to make sure that you do a couple of things with your standard geom_bar specification to make sure you you have a good experience.   

1. You've got to supply x or y constant
2. Ideally, for #1, it's y that you supply, so you don't need to mess with the theta specification in coord_polar
3. If you do any faceting, you'll want to have position set to "fill" in the geom_bar.
4. Also (not addressed here) reverse direction?  clockwise from the top?

```{r}
mpg_three <- mpg |> 
  filter(manufacturer %in% c("audi", "chevrolet", "dodge")) |>
  mutate(cyl = factor(cyl))

mpg_three |>
  ggplot() + 
  aes(y = 0) + # remember, use constant y (not x)
  aes(fill = cyl) + 
  geom_bar(
    position = "fill" # remember, if you want to facet, you'll need to reset this argument so your pies are all complete
    ) + 
  coord_polar() + 
  facet_wrap(~manufacturer) + 
  scale_x_reverse() 

last_plot() + 
  aes(fill = fct_infreq(cyl))
```

# Let's look at a high-level, first cut of `geom_pie()` that takes care of some of these finicky "*" requirements with `make_constructor`!

```{r}
geom_pie <- make_constructor(GeomBar, 
                             stat = "count", 
                             position = "fill", 
                             mapping = aes(y = 1))



ggplot(mpg_three) + 
  aes(fill = cyl) + 
  geom_pie() + 
  coord_polar()

last_plot() +
  facet_wrap(~manufacturer)
```

# but does not play nice with declaring aes locally 😔

```{r} 
mpg_three |>
  ggplot() + 
  geom_pie(aes(fill = cyl)) + # uh-oh
  coord_polar() + 
  facet_wrap(~manufacturer)
```

#  maybe we can use modify list to specify aes(y = 0) by default

```{r}
geom_pie


geom_pie <- function (mapping = NULL, data = NULL, stat = "count", 
    position = "fill", ..., lineend = "butt", linejoin = "mitre", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) 
{
  
  the_usual_required_mapping <- aes(y = 0)  #< new
  
  full_mapping <- if(is.null(mapping)){
      the_usual_required_mapping
    }else{
      modifyList(the_usual_required_mapping, mapping) #< new
    }
    

  
    layer(mapping = full_mapping, #< new
          data = data, geom = "bar", stat = stat, 
        position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
        params = rlang::list2(na.rm = na.rm, lineend = lineend, linejoin = linejoin, 
            ...))
}


# or a bit problematic (like what if you really wanted to set y yourself) but easier to think about 
# geom_pie <- function (mapping = NULL, data = NULL, stat = "count",
#     position = "fill", ..., lineend = "butt", linejoin = "mitre",
#     na.rm = FALSE, show.legend = NA, inherit.aes = TRUE)
# {
# 
#   list(
#     layer(mapping = mapping,
#           data = data, geom = "bar", stat = stat,
#         position = position, show.legend = show.legend, inherit.aes = inherit.aes,
#         params = rlang::list2(na.rm = na.rm, lineend = lineend, linejoin = linejoin,
#             ...)),
#   aes(y = 0))
# 
#   }

```


# Hooray!  A `geom_pie()` that's *really* like, 1) bar chart + 2) polar coords!

```{r, out.width="33%", fig.show='hold'}
mpg_three |>
  ggplot() + 
  geom_pie(aes(fill = cyl)) 

last_plot()  + 
  facet_wrap(~manufacturer)

last_plot() + 
  coord_polar()
```

```{r}
# demonstrate w/ global aes.
mpg_three |>
  ggplot() + 
  aes(fill = cyl) +
  geom_pie() + 
  coord_polar() 
```







# but then we might also want to know relative size of pies - I would!

And is there any hacky way to do this with GeomBar?  Seems like with width possibly but unsure...

```{r}
mpg_three |>
  ggplot() + 
  aes(alpha = manufacturer) +
  geom_pie() + 
  coord_polar()

ggforce::GeomArc
```

```{r, eval = F, echo = F}

# geom_pie <- function (mapping = NULL, data = NULL, stat = "count", 
#     position = "fill", ..., lineend = "butt", linejoin = "mitre", 
#     na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) 
# {
#   
#   quick_redefault_mapping <- aes(y = 1, 
#                                  width = after_stat(count |> sum()))
#   
#   full_mapping <- modifyList(quick_redefault_mapping, mapping)
#   
#     layer(mapping = full_mapping, data = data, geom = "bar", stat = stat, 
#         position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
#         params = rlang::list2(na.rm = na.rm, lineend = lineend, linejoin = linejoin, 
#             ...))
# }
# 
# mpg |> 
#   head(70) |>
#   ggplot() + 
#   geom_pie(aes(fill = paste("cyl:", cyl))) + 
#   coord_polar() + 
#   facet_wrap(~manufacturer)
# 
# ## this does not get to 
# mpg |> 
#   head(70) |>
#   count(cyl, manufacturer) |>
#   mutate(total_manufacturer = sum(n), .by = manufacturer) |>
#   ggplot() + 
#   aes(weight = n, fill = factor(cyl), y = 1, width = total_manufacturer) + 
#   geom_bar(position = "fill")
# 
# last_plot() + 
#   facet_wrap(~manufacturer)
# 
# last_plot() + 
#   coord_polar()

```


