---
title: "brain"
author: "Evangeline Reynolds"
date: "11/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem: positioning with SF feels spooky and data prep can feel effortful

W/ ggplot2, we are able to say 'where' with x and y.  These are positional aesthetics.  With SF the 'where' is implied with the geometry column. This feels non-declarative and spooky to me. 

Also, it often requires pre-processing to work with data you'd like to represent.

When we ask the question 'where' in a geographical context, a human being will respond with the name of a region; not coordinates and not a set of coordinates that define a region.  Is it possible to create an interface where you define 'where' with geographical names or other ids?  For example: `aes(country = my_countries_variable)` could be used with a geom/stat like geom_polygon_countries().





```{r}
library(tidyverse)
nato_names <- c("Albania", "Belgium", "Bulgaria", "Canada", "Croatia", "Czech Republic", "Denmark", "Estonia", "France", "Germany", "Greece", "Hungary",  
                "Iceland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Montenegro", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Turkey", "United Kingdom", "United States")  
library(gapminder)  
gapminder %>%  
  filter(year == 2002) %>%  
  rename(name = country) ->  
gapminder_2002_prepped  

rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%  
  select(name, pop_est, gdp_md_est,  
         continent, geometry) %>%  
  filter(name %in% nato_names) %>%  
  left_join(gapminder_2002_prepped, by = "name") ->  
  nato_countries  

ggplot(data = nato_countries) +  
  # This part feels spooky to me 
  ggplot2::geom_sf(data = nato_countries) +  
    aes(fill = lifeExp) +  
    scale_fill_viridis_c(option = "magma", direction = -1) +  
  ggplot2::coord_sf(xlim = c(-175, 47.5),  
           ylim = c(23, 85),  
           expand = FALSE) +  
  labs(title = "Life Expectancy, 2002",  
         subtitle = "NATO Member States",  
         fill = "Share of GDP (%)") +  
  theme_bw() +  
  theme(plot.title =  
            element_text(hjust = 0.5)) +  
  theme(plot.subtitle =
            element_text(hjust = 0.5))
```

See step-by-step:
https://evamaerey.github.io/ggplot2_grammar_guide/geom_sf.html#1

# Wait, why is this project called 'brain'?

One inspiration for this is ggseg too. Very cool project that uses sf spatial approach.  I wonder if there'd be a version where regions of the brain could be simply named, and other aesthetics defined. Motivated by how I'd like to learn more about the regions of the brain!  :-)


## A different API?

Here is what I think might be an easy to use API.

```
ggworld(data = gapminder_2002, projection = 'a good one') + 
  aes(country = country, # positioning aesthetic
      fill = lifeExp) + 
  geom_polygon_country()
```

Note: country names are really hard. Other codes could be used.  One could try to use the wonderful countrycodes as well when 

# Proof of concept? (without SF)

Because of my limited knowledge of SF, let's start w/ the x and y inputs approach to mapping with ggplot2 and see if we can cook something up. 

## Step 00: See how someone gets it done

I was here:
https://remiller1450.github.io/s230s19/Intro_maps.html

## Step 0: get it done w/ ggplot2

Look at mapping w/ ggplot2, no sf, we use geom_polygon and the x, y and group aesthetics).  It is not very efficient (lots of redundant info in the ), but we won't worry about that for this demonstration.  A little rewrite to get inspired.

```{r}
library(ggplot2) 
library(dplyr)
library(maps)
MainStates <- map_data("state")
StatePopulation <- read.csv("https://raw.githubusercontent.com/ds4stats/r-tutorials/master/intro-maps/data/StatePopulation.csv", as.is = TRUE)

MergedStates <- inner_join(MainStates, StatePopulation, by = "region")

StatePopulation %>% # becomes ggus()
  right_join(MainStates) %>% # becomes ggus()
  ggplot() + # becomes ggus()
  geom_polygon() + # becomes geom_polygon state
  aes(x = long, # becomes state = region
      y = lat, # becomes state = region
      group = group, # becomes state = region
      fill = population/1000000) + 
  coord_map(projection = "orthographic", 
            orientation = c(41, -74, 0))
```


# Step 1: compute function and test

```{r}
library(maps)
MainStates <- map_data("state")
# AllCounty <- map_data("county")
# MainCities <- filter(us.cities, long>=-130)

StatePopulation <- read.csv("https://raw.githubusercontent.com/ds4stats/r-tutorials/master/intro-maps/data/StatePopulation.csv", as.is = TRUE)

compute_state <- function(data, scales){
 
  data %>% 
    inner_join(map_data("state") %>% rename(state = region)) %>% 
    mutate(x = long, y = lat) %>% 
    mutate(group = group) 

}

StatePopulation %>% 
  rename(state = region) %>% 
  compute_state() ->
computed

str(computed)


computed %>% 
  ggplot() + 
  aes(x = x, y = y) + 
  geom_point()

my_setup_data <- function(data, params){
                                    if(data$group[1] == -1){
                                      nrows <- nrow(data)
                                      data$group <- seq_len(nrows)
                                    }
                                    data
}
```


# Step 2: pass to ggproto

```{r}


StatPolygonstate <- ggplot2::ggproto(`_class` = "StatPolygonstate",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("state"),
                                  compute_panel = compute_state#,
                                   # setup_data = my_setup_data
                                  # default_aes = aes(group = after_stat(state))
                                  )
```


# Step 3: write geom_* function

```{r}
geom_polygon_state <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatPolygonstate,  # proto object from step 2
    geom = ggplot2::GeomPolygon,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

# step 4: test geom_* function, celebrate!

```{r, eval = T}
StatePopulation %>% 
  mutate(region = as.factor(region)) %>% 
ggplot(data = .) + 
  aes(state = region, # state indicates position instead of x and y 
      fill = population) +
  geom_polygon_state() + 
  coord_map() + 
  geom_polygon_state(data = . %>% 
                       filter(region == "alabama"),
                     fill = "darkred" ) + 
  coord_map(projection = "orthographic", 
            orientation = c(41, -74, 0))
```


# Step 5: What's not working and taking a break.

1. Not sure what's going on w/ merge so that we only get the 4 states.  
2. state = var, var must be factor (not character).  I've seen this before and I'd like that not to be the case, but fine for now. 
3. We are celebrating in spite of 1 and 2!  And putting aside.

```{r}
StatePopulation %>% distinct(region) %>% pull()
```


---

# 0 County

```{r}
AllCounty <- map_data("county")
ggplot() + geom_polygon( data=AllCounty, aes(x=long, y=lat, group=group),
                color="darkblue", fill="lightblue", size = .1 ) +
  
          geom_polygon( data=MainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue",  size = 1, alpha = .3)
```

# Circle pack seems similar...


https://evamaerey.github.io/mytidytuesday/2022-05-28-circle-pack/circle_pack_stipple.html

