---
title: "+ aes(dims = dims(flexibly list vars))"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts

So, ggplot2 seems to be not well suited to taking a dataframe and doing dimensionality reduction within.  The assumption is that you will specify each and every variable to be visualized (which feels very intuitive and is very nice to use).  But there's a whole world of dimension reduction and visualization that accompanies it, and everybody is left to their own devices to try to figure out how that should happen when using ggplot2.  But 

## goal

<details> 
 
### so we can certainly use aes to refer to some variables.

```{r}
library(tidyverse)

dims <- function(...){}

iris |> 
  ggplot() + 
  aes(dims = dims(Sepal.Length:Sepal.Width, Petal.Width))

last_plot()$mapping
```


### but we need these expanded out

Our syntax will actually use mvars, specifying each variable individually and vars_unpack within our computation.

```{r, eval = F}
iris |> 
  ggplot() + 
  aes(dims = 
        mvars(Sepal.Length, Sepal.Width, 
                Petal.Length, Petal.Width),
      fill = Species) +
  geom_tsne0()

```


```{r}
mvars <- function(...) {
  
  varnames <- as.character(ensyms(...))
  vars <- list(...)
  listvec <- asplit(do.call(cbind, vars), 1)
  structure(listvec, varnames = varnames)

  }

vars_unpack <- function(x) {
  pca_vars <- x
  df <- do.call(rbind, pca_vars)
  colnames(df) <- attr(pca_vars, "varnames")
  as.data.frame(df)
  
}
```


### so let's use some ggplot_add to try to expand, and have these vars listed out individually

```{r}
iris |> 
  ggplot() + 
  aes(dims = dims(Sepal.Length:Petal.Length, Petal.Width))

p <- last_plot()

p$mapping$dims[[2]]  # the unexpanded expression

p$mapping$dims |> 
  as.character() |> 
  _[2] |> 
  stringr::str_extract("\\(.+") |> 
  stringr::str_remove_all("\\(|\\)") -> 
selected_var_names_expr

selected_var_names <- 
  selected_var_names_expr |> 
  str_split(", ") |> 
  _[[1]]
  
var_names <- c()

for(i in 1:length(selected_var_names)){

  new_var_names <- select(last_plot()$data, !!!list(rlang::parse_expr(selected_var_names[i]))) |> names()
  
var_names <- c(var_names, new_var_names)
  
}

expanded_vars <- var_names |> paste(collapse = ", ") 

new_dim_expr <- paste("mvars(", expanded_vars, ")")

p$mapping$dims[[2]] <- rlang::parse_expr(new_dim_expr)

p$mapping$dims[[2]]


```


```{r}
#' @export
dims_expand <- function() {

  structure(
    list(
      # data_spec = data,
         # vars_spec = rlang::enquo(vars)
         ), 
    class = "dims_expand"
    )

}

#' @import ggplot2
#' @importFrom ggplot2 ggplot_add
#' @export
ggplot_add.dims_expand <- function(object, plot, object_name) {
  
plot$mapping$dims |> 
  as.character() |> 
  _[2] |> 
  stringr::str_extract("\\(.+") |> 
  stringr::str_remove_all("\\(|\\)") -> 
selected_var_names_expr

selected_var_names <- 
  selected_var_names_expr |> 
  str_split(", ") |> 
  _[[1]]
  
var_names <- c()

for(i in 1:length(selected_var_names)){

  new_var_names <- select(plot$data, !!!list(rlang::parse_expr(selected_var_names[i]))) |> names()
  
var_names <- c(var_names, new_var_names)
  
}

expanded_vars <- var_names |> paste(collapse = ", ") 

new_dim_expr <- paste("mvars(", expanded_vars, ")")

plot$mapping$dims[[2]] <- rlang::parse_expr(new_dim_expr)

plot

}



iris |> 
  ggplot() + 
  aes(dims = dims(Sepal.Length:Petal.Length, Petal.Width)) + 
  dims_expand()

last_plot()$mapping

```


```{r}
# compute_tsne0 allows individually listed variables that are all of the same type
compute_tsne0 <- function(data, scales, perplexity = 20){
  
  set.seed(1345)
  
# identify duplicates just based on tsne data
data |>
  select(dims) |>
  mutate(vars_unpack(dims)) |>
  select(-dims) ->
data_unpacked ; data_unpacked

names_predictors <- names(data_unpacked); names_predictors

data_unpacked |>
   duplicated() ->
dups ; dups
# #
# # #
data_unpacked |>
    bind_cols(data) |>
     _[!dups,] |> 
  remove_missing() ->
clean_data ; clean_data
# # # 
clean_data |>
  _[names_predictors] |>
  as.matrix() |>
  Rtsne::Rtsne(perplexity = perplexity) |>
  _$Y |>
  as_tibble() |>
 rename(x = V1, y = V2) |>
 bind_cols(clean_data)
#   

}


iris |> 
  mutate(dims = mvars(Sepal.Length, Sepal.Width, 
                Petal.Length, Petal.Width)) |>
  select(dims) |>
  compute_tsne0()
  

StatTsne0 <- ggproto("StatTsne0", Stat, 
                     compute_panel = compute_tsne0)

GeomPointFill <- ggproto("GeomPointFill", 
                         GeomPoint,
                         default_aes = 
                           modifyList(GeomPoint$default_aes, 
                                      aes(shape = 21, 
                                          color = from_theme(paper),
                                          size = from_theme(pointsize * 2.5),
                                          alpha = .7,
                                          fill = from_theme(ink))))

geom_tsne0 <- make_constructor(GeomPointFill, stat = StatTsne0, perplexity = 30)

iris |> 
  ggplot() + 
  aes(dims = 
        mvars(Sepal.Length, Sepal.Width, 
                Petal.Length, Petal.Width),
      fill = Species) +
  geom_tsne0()


p$mapping$dims
p + 
  geom_tsne0() + 
  aes(fill = Species)
```



```{r}
geom_tsne <- function(...){
  list(
    dims_expand(),
    geom_tsne0(...)
  )
}
```

</details> 


```{r}
iris |> 
  ggplot() + 
  aes(dims = dims(Sepal.Length:Petal.Length, Petal.Width)) 

last_plot() +
  geom_tsne()

last_plot() + 
  aes(fill = Species)

last_plot() + 
  dims_expand() # stable...
```

# Now UMAP

<details>

```{r}
compute_umap <- function(data, scales, n_components = 2, random_state = 15){
  
set.seed(1345)
  
# identify duplicates just based on tsne data
data |>
  select(dims) |>
  mutate(vars_unpack(dims)) |>
  select(-dims) ->
data_unpacked ; data_unpacked

names_predictors <- names(data_unpacked); names_predictors

data_unpacked |>
    bind_cols(data) |>
  remove_missing() ->
clean_data ; clean_data

# # # 
clean_data |>
  _[names_predictors] |>
  umap::umap(n_components = n_components, random_state = random_state)  |>
  _$layout |>
  as_tibble() |>
 rename(x = V1, y = V2) |>
 bind_cols(clean_data)
#   

}

iris |> 
  mutate(dims = 
        mvars(Sepal.Length, Sepal.Width, 
                Petal.Length, Petal.Width)) |>
  select(color = Species, dims) |>
  compute_umap()

StatUmap <- ggproto("StatUmap", Stat, 
                     compute_panel = compute_umap)

geom_umap0 <- make_constructor(GeomPointFill, stat = StatUmap, random_state = 15, n_components = 4)

geom_umap <- function(...){
  
  list(dims_expand(), 
       geom_umap0())
  
}
```

</details>


```{r}
iris |> 
  ggplot() + 
  aes(dims = dims(Sepal.Length:Petal.Width)) + 
  geom_umap()

last_plot() + 
  aes(fill = Species)

palmerpenguins::penguins |> 
  sample_n(size = 200) |>
  remove_missing() |> 
  ggplot() + 
  aes(dims = dims(bill_length_mm:body_mass_g)) + 
  geom_umap() 

last_plot() + 
  aes(fill = species)
```



```{r}
knitr::knit_exit()

```



