---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)
library(palmerpenguins)

beak_means_df <- penguins |>
  dplyr::group_by(species) |>
  # I just need to change this name...
  dplyr::summarise(mean_length = mean(bill_length_mm, na.rm = TRUE))

penguins |>
  ggplot(aes(x = bill_length_mm, y = species)) +
  # ... and this accordingly
  geom_segment(
    data = beak_means_df,
    aes(x = mean_length, xend = mean_length, y = -Inf, yend = species),
    linetype = 3
  ) +
  geom_jitter(
    aes(
      fill = species,
      size = body_mass_g
    ),
    shape = 21,
    width = 0,
    height = 0.15,
    colour = "#1A242F",
    stroke = 0.5
  ) +
  ggtext::geom_textbox(
    data = beak_means_df,
    aes(
     # ... and this
      x = mean_length,
      y = species,
      label = paste0(
        species,
        " mean<br>**",
        janitor::round_half_up(mean_length),
        "mm**"
      )
    ),
    hjust = 0,
    fill = NA,
    nudge_y = -0.3,
    box.colour = NA
  )





```


## Experiment
```{r}
# install.packages("renv")
renv::init()
```
    This will create a `renv/` folder and an `.Rprofile` file in your project directory. It also creates a 
project-local library.
3.  **Install the desired version of the package within *this project*:**
    *   **From CRAN archives:** Use `remotes::install_version()`
        ```R
        # If you don't have remotes installed globally:
        # install.packages("remotes")
        remotes::install_version("ggplot2", version = "3.3.0", repos = "http://cran.us.r-project.org")
        ```
    *   **From GitHub (if it's a dev version or specific tag):** Use `remotes::install_github()`
        ```R
        remotes::install_github("tidyverse/ggplot2", ref = "v3.3.0") # Use a specific tag or commit
        ```
    R will install this version into your project's `renv/library` directory.
4.  **Load the package:**
    ```R
    library(ggplot2)
    packageVersion("ggplot2")
    ```
    This will now load the version installed in your project's `renv` library.
5.  **Snapshot (optional but recommended):**
    ```R
    renv::snapshot()


## Closing remarks, Other Relevant Work, Caveats
