---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

# convert papers into podcasts...

# failure as an academic... 

# conversation/communication as motivation for language emerging?

# how powerful is conversation, how improbable is it.  Encoding meaning into grunts, utterances, and phrases, prose...? Why does a large contingent of the world's population agree that '

# ggplot2 2017.. This is it... last_plot() 

# and then + aes()

# Disputed goodness... In language, convention is very strong...


# ggplot2 has this *really* conversational dimention (if we can stomach it)

# 'Epic' conversations

# Longform statistical narratives... 

# ggplyr

# ggregions... Texas

# ggdims ... dimension reduction

# ggplot2 extenders, stamtisch, bate papo, Johnny's of Newton.


# data science for *everyone*... Pretty literal with this... what if we translate ggplot



# Statistics 










## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)

```

## Experiment

```{r}

```



## Closing remarks, Other Relevant Work, Caveats
