---
title: "geom smooth"
author: "Evangeline Reynolds"
date: "9/1/2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
```

# Background 

## geom_smooth old hat


```{r stat-smooth}
library(tidyverse, )

mtcars %>% 
  ggplot() +
  aes(wt, mpg) + 
  geom_point() + 
  stat_smooth()
```

## a less common move:  stat_smooth(geom = "point", color = "blue")

Specific predictions (80 evenly spaced between xmin and xmax). A bit of under-the-hood thinking

```{r}
mtcars %>% 
  ggplot() +
  aes(wt, mpg) + 
  stat_smooth(geom = "point", color = "blue")
```

## almost surely new to you (and probably more interesting to stats instructors): predicting at observed values of x

xseq is not advertised, but possibly of interest.. https://ggplot2.tidyverse.org/reference/geom_smooth.html

```{r}
# fit where the support is in the data... 
mtcars %>% 
  ggplot() +
  aes(wt, mpg) + 
  stat_smooth(geom = "point", 
              color = "blue", 
              xseq = mtcars$wt)
```

# Proposal: Can we make this available in a bit 'smoother' way? 

## Target code

```{r, eval = F}
mtcars %>% 
  ggplot() + 
  aes(wt, mpg) + 
  geom_smooth_fit() + 
  geom_smooth_error() 
```

##  And also explicitly visualize error (residuals as segments)

```{r}
last_plot() + 
  geom_smooth_error() 
```

## Details of implementation...


```{r}
# This function exists to silence an undeclared import warning
gam_method <- function() mgcv::gam
data_frame0 <- function(...) data_frame(..., .name_repair = "minimal")
predictdf <- function(model, xseq, se, level) UseMethod("predictdf")
```

# prediction mechanisms from ggplot2..

```{r, echo = F}
#' @export
predictdf.default <- function(model, xseq, se, level) {
  pred <- stats::predict(
    model,
    newdata = data_frame0(x = xseq),
    se.fit = se,
    level = level,
    interval = if (se) "confidence" else "none"
  )

  if (se) {
    fit <- as.data.frame(pred$fit)
    names(fit) <- c("y", "ymin", "ymax")
    base::data.frame(x = xseq, fit, se = pred$se.fit)
  } else {
    base::data.frame(x = xseq, y = as.vector(pred))
  }
}

#' @export
predictdf.glm <- function(model, xseq, se, level) {
  pred <- stats::predict(
    model,
    newdata = data_frame0(x = xseq),
    se.fit = se,
    type = "link"
  )

  if (se) {
    std <- stats::qnorm(level / 2 + 0.5)
    base::data.frame(
      x = xseq,
      y = model$family$linkinv(as.vector(pred$fit)),
      ymin = model$family$linkinv(as.vector(pred$fit - std * pred$se.fit)),
      ymax = model$family$linkinv(as.vector(pred$fit + std * pred$se.fit)),
      se = as.vector(pred$se.fit)
    )
  } else {
    base::data.frame(x = xseq, y = model$family$linkinv(as.vector(pred)))
  }
}

#' @export
predictdf.loess <- function(model, xseq, se, level) {
  pred <- stats::predict(
    model,
    newdata = data_frame0(x = xseq),
    se = se
  )

  if (se) {
    y <- pred$fit
    ci <- pred$se.fit * stats::qt(level / 2 + .5, pred$df)
    ymin <- y - ci
    ymax <- y + ci
    base::data.frame(x = xseq, y, ymin, ymax, se = pred$se.fit)
  } else {
    base::data.frame(x = xseq, y = as.vector(pred))
  }
}

#' @export
predictdf.locfit <- function(model, xseq, se, level) {
  pred <- stats::predict(
    model,
    newdata = data_frame0(x = xseq),
    se.fit = se
  )

  if (se) {
    y <- pred$fit
    ci <- pred$se.fit * stats::qt(level / 2 + .5, model$dp["df2"])
    ymin <- y - ci
    ymax <- y + ci
    base::data.frame(x = xseq, y, ymin, ymax, se = pred$se.fit)
  } else {
    base::data.frame(x = xseq, y = as.vector(pred))
  }
}
```

# compute_group_smooth: verbatim from StatSmooth definition...

```{r, echo = F}
# would be intersted in changeing the x sequence to observed values of x, for drawing fitted and residuals
compute_group_smooth <- function(data, scales, method = NULL, formula = NULL,
                           se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
                           xseq = NULL, 
                           level = 0.95, method.args = list(),
                           na.rm = FALSE, flipped_aes = NA) {
    data <- flip_data(data, flipped_aes)
    if (vctrs::vec_unique_count(data$x) < 2) {
      # Not enough data to perform fit
      return(data_frame0())
    }

    if (is.null(data$weight)) data$weight <- 1

    
    # which values of x should we predict for ? 
    if (is.null(xseq)) {
      if (is.integer(data$x)) {
        if (fullrange) {
          xseq <- scales$x$dimension()
        } else {
          xseq <- sort(unique0(data$x))
        }
      } else {
        if (fullrange) {
          range <- scales$x$dimension()
        } else {
          range <- range(data$x, na.rm = TRUE)
        }
        xseq <- seq(range[1], range[2], length.out = n)
      }
    }

    # Special case span because it's the most commonly used model argument
    if (identical(method, "loess")) {
      method.args$span <- span
    }

    if (is.character(method)) {
      if (identical(method, "gam")) {
        method <- gam_method()
      } else {
        method <- match.fun(method)
      }
    }
    # If gam and gam's method is not specified by the user then use REML
    if (identical(method, gam_method()) && is.null(method.args$method)) {
      method.args$method <- "REML"
    }

    model <- rlang::inject(method(
      formula,
      data = data,
      weights = weight,
      !!!method.args
    ))

    prediction <- predictdf(model, xseq, se, level)
    prediction$flipped_aes <- flipped_aes
    flip_data(prediction, flipped_aes)
}
```


```{r}
compute_group_smooth_error <- function(data, scales, method = NULL, formula = NULL,
                           se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
                           level = 0.95, method.args = list(),
                           na.rm = FALSE, flipped_aes = NA){
  
  
compute_group_smooth(data = data, scales = scales, 
                       method = method, formula = formula, 
                       se = FALSE, n= n, span = span, fullrange = fullrange,
                       xseq = data$x, 
                       level = .95, method.args = method.args, 
                       na.rm = na.rm, flipped_aes = flipped_aes) %>% 
    mutate(xend = data$x,
           yend = data$y,
           ymin = y,
           xmin = x,
           ymax = yend,
           xmax = x + (ymax - ymin)
           )
  
}
```



# test compute group 

```{r}
library(dplyr)
mtcars %>%
  rename(x = wt, y = mpg, cat = am) %>%
  compute_group_smooth(method = lm, formula = y ~ x)

mtcars %>% 
  rename(x = wt, y = mpg, cat = am) %>% 
  compute_group_smooth_error(method = lm, formula = y ~ x, n = 7)
```


```{r}
setup_smooth <- function(data, params) {
    params$flipped_aes <- has_flipped_aes(data, params, ambiguous = TRUE)
    msg <- character()
    if (is.null(params$method) || identical(params$method, "auto")) {
      # Use loess for small datasets, gam with a cubic regression basis for
      # larger. Based on size of the _largest_ group to avoid bad memory
      # behaviour of loess
      max_group <- max(table(interaction(data$group, data$PANEL, drop = TRUE)))

      if (max_group < 1000) {
        params$method <- "loess"
      } else {
        params$method <- "gam"
      }
      msg <- c(msg, paste0("method = '", params$method, "'"))
    }

    if (is.null(params$formula)) {
      if (identical(params$method, "gam")) {
        params$formula <- y ~ s(x, bs = "cs")
      } else {
        params$formula <- y ~ x
      }
      msg <- c(msg, paste0("formula = '", deparse(params$formula), "'"))
    }
    if (identical(params$method, "gam")) {
      params$method <- gam_method()
    }

    if (length(msg) > 0) {
      cli::cli_inform("{.fn geom_smooth} using {msg}")
    }

    params
  }



#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
StatSmooth <- ggplot2::ggproto("StatSmooth", ggplot2::Stat,
  setup_params = setup_smooth,
  extra_params = c("na.rm", "orientation"),
  compute_group = compute_group_smooth,
  dropped_aes = c("weight"),
  required_aes = c("x", "y")
)


StatSmooth2 <- StatSmooth

StatSmoothError$compute_group <- compute_group_smooth_error

StatSmoothError <- ggplot2::ggproto("StatSmoothError", ggplot2::Stat,
  setup_params = setup_smooth,
  extra_params = c("na.rm", "orientation"),
  compute_group = compute_group_smooth_error,
  dropped_aes = c("weight"),
  required_aes = c("x", "y")
)


stat_smooth <- function(mapping = NULL, data = NULL,
                        geom = "smooth", position = "identity",
                        ...,
                        method = NULL,
                        formula = NULL,
                        se = TRUE,
                        n = 80,
                        span = 0.75,
                        fullrange = FALSE,
                        level = 0.95,
                        method.args = list(),
                        na.rm = FALSE,
                        orientation = NA,
                        show.legend = NA,
                        inherit.aes = TRUE) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatSmooth,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = rlang::list2(
      method = method,
      formula = formula,
      se = se,
      n = n,
      fullrange = fullrange,
      level = level,
      na.rm = na.rm,
      orientation = orientation,
      method.args = method.args,
      span = span,
      ...
    )
  )
}

stat_smooth_obs <- function(mapping = NULL, data = NULL,
                        geom = "smooth", position = "identity",
                        ...,
                        method = NULL,
                        formula = NULL,
                        se = TRUE,
                        n = 80,
                        span = 0.75,
                        fullrange = FALSE,
                        level = 0.95,
                        method.args = list(),
                        na.rm = FALSE,
                        orientation = NA,
                        show.legend = NA,
                        inherit.aes = TRUE) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatSmoothError,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = rlang::list2(
      method = method,
      formula = formula,
      se = se,
      n = n,
      fullrange = fullrange,
      level = level,
      na.rm = na.rm,
      orientation = orientation,
      method.args = method.args,
      span = span,
      ...
    )
  )
}





```


# Basically working!

```{r}


mtcars %>% 
  ggplot() +
  aes(wt, mpg) + 
  geom_point() +
  geom_smooth(alpha = .2, se = FALSE) 


last_plot() +
  stat_smooth_obs(geom = "point", color = "blue") +  # wrap as geom_smooth_fit()
  stat_smooth_obs(geom = "segment") # geom_smooth_error()
```

## even squared error (best w/ stdz and coord_equal)

```{r}
stdz <- function(x){
  
  var_mean <- mean(x) 
  var_sd <- sd(x)
  
  (x-var_mean)/var_sd
  
}

last_plot() + 
  stat_smooth_obs(geom = "rect", alpha = .1)  + # geom_smooth_error_sq() +
  aes(stdz(wt), stdz(mpg)) + 
  coord_equal()


mtcars %>% 
  ggplot() +
  aes(wt, mpg) + 
  geom_point() +
  geom_smooth(alpha = .2, se = FALSE, method = lm) + 
  stat_smooth_obs(geom = "point", color = "blue", method = lm) + # wrap as geom_smooth_fit()
  stat_smooth_obs(geom = "segment", method = lm)



```


# issues

- Squares hanging out in the background might not be desirable.  Maybe break up the computation.
