---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F)
options(tidyverse.quiet = TRUE)
```


```{r, fig.height=20, fig.width=20}
library(tidyverse)
library(ggwordcloud)
library(ggstamp)

extension_exported <- "https://raw.githubusercontent.com/EvaMaeRey/mytidytuesday/refs/heads/main/2024-11-19-gg-prefixes/exported_funs_exts_ggplot2_tidyverse_org.csv" |> read_csv()


extension_exported %>% 
  distinct(user)
```

## Intro Thoughts


# Description for Figure

## website


The ggplot2 extension gallery helps people explore the ggplot2 extension ecosystem.  Package developers can submit their extension packages to help others find their work.  About 140 packages have been submitted 



## star

The star contains packages names, with larger sized names corresponding to packages with greater numbers of exported functions.

## branches

Branches and background contain exported function names.  There are more than 5000 exported functions in the submitted packages. 

## trunk

The tree trunk is made up of package authors (github user). More than 100 unique package authors have contributed to the ecosystem.  




```{r}
lightparser::split_to_tbl("tree-w-ggwordcloud.Rmd") %>% 
  filter(section %in% c("website", "star", "branches", "trunk")) %>% 
  filter(type == "inline") %>% 
  unnest(cols = text) %>% 
  filter(text != "") %>% 
  select(section, text) ->
inline_text
  
inline_text$text
```



```{r, fig.width=20, fig.height=20}
ggplot(cars) +
  aes(speed, dist) + 
  geom_point() + 
  annotate(ggtext::GeomTextBox, 
           x = I(.05), 
           y = I(c(.95, .75, .5, .25)),
           label = inline_text$text,
           hjust = 0, vjust = 1, size = 8, 
           width = unit(5, "inch"), fill = "lavender"
           ) + 
  theme_void()
```









```{r wordcloud, fig.height=20,fig.width=20}
# background
ggplot() +    
  geom_text_wordcloud(
    aes(label = fun_exported, size = 2),
    data = extension_exported %>% 
      sample_n(1100),
    shape = "square", alpha = .2) +
  ggstamp::theme_void_fill(fill = "whitesmoke") +
  NULL ->
background; background


# Star
ggplot() + 
  geom_text_wordcloud(
    aes(label = repo, size = n),
    data = extension_exported %>% count(repo),
    shape = "star", color = "goldenrod4", y = .87) +
  theme_void() +
  # labs(title = "ggplot2 extensions on") +
  # annotate("text", x = I(.1), y = I(.87),
  #          label = "More than Extension packages...", size = 8) +
  NULL ->
star; star

# branches
ggplot() +    
  geom_text_wordcloud(
    aes(label = fun_exported, size = 2),
    data = extension_exported %>% 
      sample_n(500) %>%
      filter(str_count(fun_exported) < 18),
    shape = "triangle-upright",
                      color = "darkseagreen4", y = .4) +
  theme_void() +
  NULL ->
branches; branches


# base
ggplot() +
  geom_text_wordcloud(
    data = extension_exported %>% count(user),
    aes(label = user),
    shape = "square", color = "burlywood4", x = .5, y = .1) +
  theme_void() +
  annotate(ggtext::GeomTextBox,
           x = I(c(.05, .65, .12, .1)),
           y = I(c(.95, .85, .6, .2)),
           label = inline_text$text,
           hjust = 0, vjust = 1, 
           size = c(10,8,8,8),
           width = unit(5, "inch"), fill = "lavender"
           ) +
  NULL           ->
base; base

ggsave("background.png", background, width = 20, height = 20)
ggsave("star.png", star, width = 20, height = 20)
ggsave("branches.png", branches, width = 20, height = 20)
ggsave("base.png", base, width = 20, height = 20)

```

```{r}
library(magick)
c(image_read("background.png"),
  image_read("star.png"),
  image_read("branches.png"),
  image_read("base.png")
  ) %>% 
  magick::image_flatten() %>% 
  magick::image_write("tree.png")

image_read("tree.png")

```


```{r}
knitr::knit_exit()
```





```{r, eval=FALSE}

download.file("https://github.com/CHOP-CGTInformatics/ggswim/raw/main/man/figures/logo.png", destfile = "ggswim_logo.png")

download.file("https://github.com/teunbrand/ggh4x/raw/main/man/figures/logo_300px.png", destfile = "ggh4x_logo.png")

download.file("https://github.com/yjunechoe/ggtrace/raw/main/man/figures/logo.png",
              destfile = "ggtrace_logo.png")

download.file("https://github.com/pharmaverse/ggsurvfit/raw/main/man/figures/logo.png", destfile = "ggsurvfit_logo.png")

# ggstats, GGally


logo_urls <- c("https://github.com/CHOP-CGTInformatics/ggswimraw/main/man/figures/logo.png",#swimlogo
"https://github.com/corybrunson/ordr/raw/main/man/figures/logo.png"         ,#
"https://github.com/tidyverse/ggplot2/raw/main/man/figures/logo.png"        ,#3
"https://github.com/larmarange/ggstats/raw/main/man/figures/logo.png"       ,#
"https://github.com/jamesotto852/ggdensity/raw/main/man/figures/logo.png"   ,#5
"https://github.com/animint/animint2raw/main/man/figures/logo.png"          ,# 
"https://github.com/eliocamp/ggnewscale/raw/main/man/figures/logo.png"      ,#7
"https://github.com/yonicd/ggedit/raw/main/man/figures/logo.png"             ,#
"https://github.com/ggobi/ggally/raw/main/man/figures/logo.png"             ,#
"https://github.com/r-causal/ggdag/raw/main/man/figures/logo.png"           ,#10
"https://github.com/nrennie/ggflowchart/raw/main/man/figures/logo.png"       ,# 
"https://github.com/wjschne/arrowheadr/raw/main/man/figures/logo.png"        ,#
"https://github.com/jonocarroll/ggeasy/raw/main/man/figures/logo.png"       ,#
"https://github.com/pharmaverse/ggsurvfit/raw/main/man/figures/logo.png"    ,#14
"https://github.com/EvaMaeRey/ggcirclepack/raw/main/man/figures/logo.png"    ,#
"https://github.com/yjunechoe/ggtrace/raw/main/man/figures/logo.png"        ,# 16
"https://github.com/teunbrand/ggh4x/raw/main/man/figures/logo.png"          ,#17
"https://github.com/AllanCameron/geomtextpath/raw/main/man/figures/logo.png",#18
"https://github.com/davidsjoberg/ggbump/raw/main/man/figures/logo.png"     )  

```


```{r, eval = F}
for (i in 1:length(logo_urls)){
  
  try(download.file(logo_urls[i] #|> str_replace("main", "master")
                    , 
                    destfile = paste0("logo", i, ".png")
                    )
      )
  
}

for (i in c(7,13)){
  
  try(download.file(logo_urls[i] |> str_replace("main", "master")
                    , 
                    destfile = paste0("logo", i, ".png")
                    )
      )
  
}

```



## Closing remarks, Other Relevant Work, Caveats
