---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r}
library(tidyverse)
knitr::opts_chunk$set(eval = F)

```

## 

We want to look at the packages submitted to the ggplot2 extension gallery at https://exts.ggplot2.tidyverse.org/gallery/.

To add an ggplot2 extension, there is this instruction

> Add an entry for your extension in the _config.yml file of this repository with the meta data for your extension (copy another entry and modify). Please see below for guidance on the meta data.

```{r, eval = F}
 read_lines("https://raw.githubusercontent.com/ggplot2-exts/gallery/refs/heads/gh-pages/_config.yml") |> 
  write_lines("ext_config_yml.txt")
```

Here is one of the entries.

```{r}
read_lines("ext_config_yml.txt")[8:23]
```

Maybe we can easily collect all this info if we remove the returns before indentations... (the description fields)

```{r}
read_lines("ext_config_yml.txt") |>
  paste(collapse = "\n") |>
  str_replace_all("\\>\n    ", "") |>
  str_replace_all("\n    ", "") |>
  str_split("\n") ->
config_initial_parse

config_initial_parse |> head(20)

```




```{r}
read_lines("ext_config_yml.txt") %>% 
  str_detect("ghuser:|ghrepo:") %>% 
  read_lines("ext_config_yml.txt")[.] ->
info_github

info_github |> head(12)

info_github %>% 
  matrix(ncol = 2, byrow = T) %>% 
  as_tibble() %>% 
  rename(user = V1, repo = V2) %>% 
  mutate(user = str_remove(user, "ghuser: ") %>% str_trim()) %>% 
  mutate(repo = str_remove(repo, "ghrepo: ") %>% str_trim()) ->
df_gh # data frame github repository info

df_gh |> head(12)
```


To get the exported functions, we'll collect all...

```{r, fig.width=20, fig.height=20}

"https://raw.githubusercontent.com/lionel-/ggstance/refs/heads/main/NAMESPACE"

author <- "lionel-"
package <- "ggstance"
type <- "main"
sep = "/"

paste("https://raw.githubusercontent.com/lionel-/ggstance/refs/heads/main/NAMESPACE")
paste("https://raw.githubusercontent.com", author, package, 
      "refs/heads", type, "NAMESPACE", sep = sep)

namespace_main <- paste("https://raw.githubusercontent.com", df_gh$user, df_gh$repo, 
      "refs/heads/main/NAMESPACE", sep = "/")
namespace_master <- paste("https://raw.githubusercontent.com", df_gh$user, df_gh$repo, 
      "refs/heads/master/NAMESPACE", sep = "/")
```

```{r}
len <- nrow(df_gh)

namespace_main_contents <- vector(mode='list', len)
namespace_master_contents <- vector(mode='list', len)

for (i in 1:len){
  
  try(
  namespace_main_contents[[i]] <-  read_lines(namespace_main[i])
   )
  
  try(
  namespace_master_contents[[i]] <-  read_lines(namespace_master[i])
   )
  
  namespace_contents[[1]] <- ifelse(!is.null(namespace_master_contents[[i]]),
                                    namespace_master_contents[[i]],
                                ifelse(!is.null(namespace_main_contents[[i]]),
                                    namespace_main_contents[[i]], NA))
  
  Sys.sleep(.2)
}

length(namespace_main_contents)
length(namespace_master_contents)
length(namespace_contents)

df_gh %>% 
  mutate(namespace_main_contents = namespace_main_contents) %>% 
  unnest(namespace_contents = namespace_main_contents) %>% 
  select(user, repo, namespace_contents) %>% 
  remove_missing() ->
df_gh_main


df_gh %>% 
  mutate(namespace_master_contents = namespace_master_contents) %>% 
  unnest(namespace_contents = namespace_master_contents) %>% 
  select(user, repo, namespace_contents) %>% 
  remove_missing() ->
df_gh_master

df_gh_main %>% 
  bind_rows(df_gh_master) ->
df_gh


```



```{r}
df_gh %>% 
  filter(str_detect(namespace_contents,"^export")) %>%
  mutate(fun_exported = str_remove_all(namespace_contents,"export\\(|\\)")) %>% 
  select(-namespace_contents) ->
extensions_df_exports; dim(extensions_df_exports)

head(extensions_df_exports)

write_csv(extensions_df_exports, "exported_funs_exts_ggplot2_tidyverse_org.csv")
```


