---
title: "ma206distributions and ma206equations"
subtitle: "independent studies"
author: "ER, IF, MJ"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    seal: FALSE
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


class: inverse, center, middle
background-image: url(https://images.unsplash.com/photo-1509081438789-63838f150db7?ixlib=rb-4.0.3&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=387&q=80)
background-size: cover

# ma206distributions and ma206equations






---
class: inverse, middle, center

# Hello!

---
class: inverse, middle, center

#   Thank you for your willingess to have a look at some of the work we are doing in `ma206distributions` and `ma206equations`!

---
class: inverse, middle, center

#  Mathematical fluency involves frequent use of probability distributions and of course equations.

---
class: inverse, middle, center
#  However, visualizing these distributions with software and typing equations can be tedious.

---
class: inverse, middle, center
#    In a well-defined statistics and probability curriculum, the set of distributions and equations to be mastered is known and limited.
---
class: inverse, middle, center
#   The goal of this project is to provide ready-to-use functions that will make it easier to engage computationally with these distributions.
---
class: inverse, middle, center
#   Your feedback is on this work is greatly appreciated.

---
class: inverse, middle, center
#   Without further ado...

```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = F, warning = FALSE, comment = "", cache = F)

options(scipen = 8, digits = 3)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}



```








```{r newclass, include = FALSE}
library(R6)


cross_trials <- function(trial = prize_wheel, num_trials = 2){

  df <- trial
  names(df) <- paste0("t", 1,"_", names(df))

  if(num_trials > 1){
  for (i in 2:num_trials){

    temp <- trial
    names(temp) <- paste0("t", i,"_", names(trial))

    df <- tidyr::crossing(df, temp)

  }
  }

  df

}


bernoulli_trial <- function(prob = .25){

  tibble::tibble(outcome = 0:1, prob = c(1-prob, prob))

}

library(magrittr)
bernoulli_trial() %>% 
  cross_trials(num_trials = 3)


Trials <- R6Class("Trials",
                       public = list(
                         
                         # objects
                         trial = NULL,
                         index = NULL,
                         out = NULL,
                         

                         # functions
                         init = function(trial = NULL){
                           
                           self$trial <- trial
                           self$index <- 1
                           
                           self$out <- cross_trials(self$trial, num_trials = self$index)
                           
                           invisible(self)          #returns

                           
                         },
                         
                         update = function(increment = 1){ # a method
                           
                           self$index <- self$index + increment
                           
                           # displaying
                           self$out <- cross_trials(self$trial, num_trials = self$index)
                          
                           invisible(self)          #returns
                           
                                     },
                         
                         print = function() {  # print method; default is to print everything

                           print(self$out)

                         }
                       )
)
                      
                
```










```{r trialtest, eval = T, include = FALSE}

my_trials <- Trials$new() 

my_trials

my_trials$init(trial = bernoulli_trial())
my_trials$out
my_trials$update()
my_trials$out

```








```{r wrapandpipe, eval = T, include = FALSE}
# wrap and pipe
trial_init <- function(trial = NULL
                    ){
  
  my_trials <- Trials$new()
  
  
  my_trials$init(trial = trial)
  
  my_trials
  
}

bernoulli_trial() %>% trial_init()

trial_advance <- function(trials, increment = 1){
  
  my_trials <- trials
  
  my_trials$update(increment = increment)
  
  my_trials
  
}

add_trials <- function(trials, increment = 1){
  
  
  if(!is.R6(trials)){my_trials <- trial_init(trial = trials)
  
  my_trials <- trial_advance(trials = my_trials, 
                                       increment = increment)
  
  }
  if(is.R6(trials)){my_trials <- trial_advance(trials = trials, 
                                       increment = increment)}
  
  my_trials
  
}



bernoulli_trial() %>%
  trial_init() %>% 
  trial_advance() %>% 
  trial_advance(2)


sum_across <- function(data, var_key = "outcome"){
  
  dplyr::mutate(.data = data, 
                global_outcome =
                  rowSums(dplyr::across(dplyr::contains(var_key)))) 
  
}

seq_across <- function(data, var_key = "outcome"){
  
  col_list <- names(data)[names(data) %>% stringr::str_detect(var_key)]
  
  paste_collapse <- function(x){paste(x, collapse = ", ")}
  
  data$global_outcome  <- apply(data[,col_list], MARGIN = 1, FUN = paste_collapse)
  
  data
  
}


prod_across <- function(data, var_key = "prob"){
  
  col_list <- names(data)[names(data) %>% stringr::str_detect(var_key)]
  
  data$global_probs  <- apply(data[,col_list], MARGIN = 1, FUN = prod)
  
  data
  
}
```

---
class: inverse, center, middle

# tidy_dbinom() 

## The function tidy_dbinom() returns a dataframe that pairs outcomes and probabilities, using just two inputs. 


---

`r flipbookr::chunk_reveal("tidy_dbinom", title = "### tidy_dbinom() usage")`

```{r tidy_dbinom, include = F}
library(tidyverse)
library(ma206distributions)
library(ma206equations)

tidy_dbinom(single_trial_prob = 1/6, 
            num_trials = 8) ->
eight_die_rolls

tidy_dbinom(single_trial_prob = .5, 
            num_trials = 10) ->
ten_coin_flips

tidy_dbinom(single_trial_prob = 1/38, 
            num_trials = 3) ->
three_roulette_spins

```

---
class: inverse, center, middle

# geom_lollipop() and scale_x_counting() 

## Visualization functions geom_lollipop() and scale_x_counting() help concisely visualize discrete probability data frames like that produced with tidy_dbinom()  


---

`r flipbookr::chunk_reveal("tidy_dbinom_plot", title = "### geom_lollipop() and scale_x_counting() usage")`

```{r tidy_dbinom_plot, include = F}
library(ma206distributions)
library(tidyverse)
tidy_dbinom(single_trial_prob = 1/6, num_trials = 6) %>% # from ma206distributions
  ggplot() + 
  aes(x = num_successes, 
      y = probability) + 
  geom_lollipop() +
  scale_x_counting() ->
die_roll_viz

tidy_dbinom(single_trial_prob = .5, num_trials = 10) %>% # from ma206distributions
  ggplot() + 
  aes(x = num_successes, 
      y = probability) + 
  geom_lollipop() +
  scale_x_counting() ->
coin_flip_viz

```

---
class: inverse, center, middle

# stamp_dbinom()

## For convenience stamp_dbinom() wraps up the dataframe together with the plotting functionality. 


---

`r flipbookr::chunk_reveal("tidy_dbinom_stamp", title = "###  stamp_dbinom() usage")`

```{r tidy_dbinom_stamp, include = F}
ggplot() + 
  stamp_dbinom(single_trial_prob = 1/6, 
               num_trials = 10) +    
  scale_x_counting() +                                        
  labs(title = "Probability of rolling 0,1,2,...10 sixes when rolling a fair die 10 times") ->
die_roll_viz_w_stamp


ggplot() + 
  stamp_dbinom(single_trial_prob = .5, 
               num_trials = 5) +    
  labs(title = "Probability of flipping 0,1,2,3,4 or 5 heads\n when flipping a coin five times") ->
coin_flip_viz_w_stamp

```


---
class: inverse, center, middle

# stamp_eq_binomial()

## The function stamp_eq_binomial() tells you how the probability for each number of total successes is computed.

---

`r flipbookr::chunk_reveal("equations")`

```{r equations, include = F}
library(ma206equations)
library(ma206distributions)
ggplot() + 
  stamp_dbinom(single_trial_prob = .5, 
               num_trials = 5) + 
  # binomial computation
  stamp_eq_binomial(x = 4, y = .27, size = 7) + 
  # reminder of computing choose
  stamp_eq_choose(x = 4.5, y = .21)
```


---
class: inverse, center, middle

# bernoulli_trial() and adding trials()

## You can also build up binomial distributions from first principles uising bernoulli_trial() and adding trials().  


---


`r flipbookr::chunk_reveal("pipeline2", title = "## bernoulli_trial() and add_trials() usage", widths = c(1.2,3))`


```{r pipeline2, include = FALSE}
bernoulli_trial(prob = .25) %>% 
  add_trials() %>% 
  add_trials() %>% 
  .$out %>% 
  sum_across() %>% 
  prod_across() %>% 
  group_by(global_outcome) %>% 
  summarize(probs = 
              sum(global_probs))

# check consistency with 
# our automated computation
tidy_dbinom(
  single_trial_pro = .25,
  num_trials = 3)
```

