---
title: "ma206distributions and ma206equations"
subtitle: "independent studies"
author: "ER, IF, MJ"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = F, warning = FALSE, comment = "", cache = F)

options(scipen = 8, digits = 3)

options(tibble.print_min = 55)
options(knitr.duplicate.label = "allow")
options(width=300) # prevents data wrapping
```


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```



---




```{r newclass, include = FALSE}
library(R6)


cross_trials <- function(trial = prize_wheel, num_trials = 2){

  df <- trial
  names(df) <- paste0("t", 1,"_", names(df))

  if(num_trials > 1){
  for (i in 2:num_trials){

    temp <- trial
    names(temp) <- paste0("t", i,"_", names(trial))

    df <- tidyr::crossing(df, temp)

  }
  }

  df

}


bernoulli_trial <- function(prob = .25){

  tibble::tibble(outcome = 0:1, prob = c(1-prob, prob))

}

library(magrittr)
bernoulli_trial() %>% 
  cross_trials(num_trials = 3)


Trials <- R6Class("Trials",
                       public = list(
                         
                         # objects
                         trial = NULL,
                         index = NULL,
                         out = NULL,
                         

                         # functions
                         init = function(trial = NULL){
                           
                           self$trial <- trial
                           self$index <- 1
                           
                           self$out <- cross_trials(self$trial, num_trials = self$index)
                           
                           invisible(self)          #returns

                           
                         },
                         
                         update = function(increment = 1){ # a method
                           
                           self$index <- self$index + increment
                           
                           # displaying
                           self$out <- cross_trials(self$trial, num_trials = self$index)
                          
                           invisible(self)          #returns
                           
                                     },
                         
                         print = function() {  # print method; default is to print everything

                           print(self$out)

                         }
                       )
)
                      
                
```










```{r trialtest, eval = T, include = FALSE}

my_trials <- Trials$new() 

my_trials

my_trials$init(trial = bernoulli_trial())
my_trials$out
my_trials$update()
my_trials$out

```








```{r wrapandpipe, eval = T, include = FALSE}
# wrap and pipe
trial_init <- function(trial = NULL
                    ){
  
  my_trials <- Trials$new()
  
  
  my_trials$init(trial = trial)
  
  my_trials
  
}

bernoulli_trial() %>% trial_init()

trial_advance <- function(trials, increment = 1){
  
  my_trials <- trials
  
  my_trials$update(increment = increment)
  
  my_trials
  
}

add_trials <- function(trials, increment = 1){
  
  
  if(!is.R6(trials)){my_trials <- trial_init(trial = trials)
  
  my_trials <- trial_advance(trials = my_trials, 
                                       increment = increment)
  
  }
  if(is.R6(trials)){my_trials <- trial_advance(trials = trials, 
                                       increment = increment)}
  
  my_trials
  
}



bernoulli_trial() %>%
  trial_init() %>% 
  trial_advance() %>% 
  trial_advance(2)


sum_across <- function(data, var_key = "outcome"){
  
  dplyr::mutate(.data = data, 
                global_outcome =
                  rowSums(dplyr::across(dplyr::contains(var_key)))) 
  
}

seq_across <- function(data, var_key = "outcome"){
  
  col_list <- names(data)[names(data) %>% stringr::str_detect(var_key)]
  
  paste_collapse <- function(x){paste(x, collapse = ", ")}
  
  data$global_outcome  <- apply(data[,col_list], MARGIN = 1, FUN = paste_collapse)
  
  data
  
}


prod_across <- function(data, var_key = "prob"){
  
  col_list <- names(data)[names(data) %>% stringr::str_detect(var_key)]
  
  data$global_probs  <- apply(data[,col_list], MARGIN = 1, FUN = prod)
  
  data
  
}
```

---

### Mathematical fluency involves frequent use of probability distributions. However, visualizing these distributions can be tedious.   In a well-defined statistics and probability curriculum, the set of distributions to be mastered is known and limited.  In such settings having shortcuts to produce or ingest these distributions in analytic software for display and analysis could be beneficial for students and instructors alike.  The goal of this project is to provide ready-to-use objects (data frames) and functions that will make it easier to engage computationally with these distributions.  We make these objects and functions available in the R package `ma206distributions`.  In this independent study, we've focus especially on the binomial and geometric distributions as targets to be used and mastered to explore the potential for ma206distributions. 

---

`r flipbookr::chunk_reveal("tidy_dbinom", title = "# tidy_dbinom()")`

```{r tidy_dbinom, include = F}
library(ma206distributions)

tidy_dbinom(single_trial_prob = 1/6, 
            num_trials = 8)
```

---

`r flipbookr::chunk_reveal("tidy_dbinom_plot", title = "# tidy_dbinom() plotted")`

```{r tidy_dbinom_plot, include = F}
library(ma206distributions)
library(tidyverse)
tidy_dbinom(single_trial_prob = 1/6, num_trials = 6) %>% # from ma206distributions
  ggplot() + 
  aes(x = num_successes, 
      y = probability) + 
  geom_lollipop(annotate = T, 
                round_digits = 2) +  # from ma206distributions
  scale_x_counting()                 # from ma206distributions

```


---

`r flipbookr::chunk_reveal("tidy_dbinom_stamp", title = "# tidy_dbinom() stamp()")`

```{r tidy_dbinom_stamp, include = F}
ggplot() + 
  stamp_dbinom(single_trial_prob = 1/6, 
               num_trials = 10) +    # from ma206distributions
  scale_x_counting() +                                        # from ma206distributions
  labs(title = "Probability of rolling 0,1,2,...10 sixes when rolling a fair die 10 times")
```


---

`r flipbookr::chunk_reveal("pipeline")`


```{r pipeline, include = FALSE}
library(dplyr)
bernoulli_trial(prob = .5) %>% 
  add_trials() %>% 
  add_trials() %>% 
  add_trials() %>% 
  .$out %>% 
  sum_across() %>% 
  prod_across() %>% 
  group_by(global_outcome) %>% 
  summarize(probs = sum(global_probs))
```


---


`r flipbookr::chunk_reveal("pipeline2")`


```{r pipeline2, include = FALSE}
bernoulli_trial(prob = .25) %>% 
  add_trials(3) %>% 
  .$out %>% 
  seq_across() %>% 
  prod_across() %>% 
  group_by(global_outcome) %>% 
  summarize(probs = sum(global_probs))
```

