---
title: "wrap ggproto? (!?)"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

## Intro Thoughts

What if you just want to define a basic computational engine (geom_* or stat_* function) on the fly in a script.  It seems like it requires a good amount of code, but there are things that repeat.  Below, we see if we define a StatTemp within a function, and use that function to remove some of the repetition for vanilla-y extensions.

TLDR:  This seems to work, and surprisingly well (??).  I thought I'd only be able to use StatTemp once, but you seem to be able to define multiple geoms_* functions with the same define_temp_geom wrapper...

## Status Quo:  1. compute, 2. ggproto, 3. define layer

```{r cars}
library(tidyverse)
compute_panel_equilateral <- function(data, scales, n = 15){
  
  data |> 
    mutate(group = row_number()) |> 
    crossing(tibble(z = 0:n)) |>
    mutate(around = 2*pi*z/max(z)) |> 
    mutate(x = x0 + cos(around)*r,
           y = y0 + sin(around)*r) 
  
}

StatCircle <- ggproto(
  `_class` = "StatCircle",
  `_inherit` = ggplot2::Stat,
  compute_panel = compute_panel_equilateral,
  required_aes = c("x0", "y0", "r"))

geom_circle <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCircle,  # proto object from Step 2
    geom = ggplot2::GeomPolygon,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

data.frame(x0 = 0:1, y0 = 0:1, r = 1:2/3) |> 
  ggplot() + 
  aes(x0 = x0, y0 = y0, r = r) + 
  geom_circle() + 
  aes(fill = r)
```

## Experimental:  `define_temp_geom()` combines 2 and 3 in using a temp stat


```{r}
define_temp_geom_compute_panel <- function(
  required_aes,
  compute_panel, 
  geom = ggplot2::GeomPoint, 
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, 
  ...) {

StatTemp <- ggproto(
  `_class` = "StatTemp",
  `_inherit` = ggplot2::Stat,
  compute_panel = compute_panel,
  required_aes = required_aes)

  ggplot2::layer(
    stat = StatTemp,  # proto object from Step 2
    geom = geom,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

### Try it out

#### abbreviated definition `geom_circle()` using `define_temp_geom_compute_panel`

```{r}
compute_panel_circle <- function(data, scales, n = 15){
  
  data |> 
    mutate(group = row_number()) |> 
    crossing(tibble(z = 0:n)) |>
    mutate(around = 2*pi*z/max(z)) |> 
    mutate(x = x0 + cos(around)*r,
           y = y0 + sin(around)*r) 
  
}

geom_circle <- function(...){
  
  define_temp_geom_compute_panel(
    required_aes = c("x0", "y0", "r"),
    compute_panel = compute_panel_circle,
    geom = ggplot2::GeomPath,
    ...)
  
}
```


#### use `geom_circle()`

```{r}
data.frame(x0 = 0:1, y0 = 0:1, r = 1:2/3) |>
  ggplot() +
  aes(x0 = x0, y0 = y0, r = r) +
  geom_circle() +
  aes(fill = r)
```

### Can you define a second w/ the same StatTemp...

#### define geom_heart

```{r}
compute_panel_heart <- function(data, scales){

  data %>%
  mutate(group = row_number()) %>%
  tidyr::crossing(around = 0:15/15) %>%
    dplyr::mutate(
      y = y0 + r * (
        .85 * cos(2*pi*around)
        - .35 * cos(2 * 2*pi*around)
        - .25 * cos(3 * 2*pi*around)
        - .05 * cos(4 * 2*pi*around)
      ),
      x = x0 + r * (sin(2*pi*around)^3))

}

geom_heart <- function(...){

    define_temp_geom_compute_panel(
      required_aes = c("x0", "y0", "r"),
      compute_panel = compute_panel_heart,
      geom = ggplot2::GeomPolygon,
      ...)

  }
```

#### try using both geom_heart and geom_circle together...

```{r}
data.frame(x0 = 0:1, y0 = 0:1, r = 1:2/3) |>
  ggplot() +
  aes(x0 = x0, y0 = y0, r = r) +
  geom_heart(alpha = .3) +
  geom_circle(color = "red", 
              data = data.frame(x0 = 4,y0 = 2, r = 1)) + 
  annotate(geom = "point", x = .5, y = .5, size = 8, color = "green")

```


## Closing remarks, Other Relevant Work, Caveats
