---
title: "two-sample t test"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, out.width = "100%", dpi = 500)
options(tidyverse.quiet = TRUE)
```





## Step 1. Visualize the difference in means

```{r}
library(tidyverse)
library(ggt.test)

```


```{r, fig.width=9, echo = F}
"data_haircuts |> 
  ggplot() + 
  aes(x = price) + 
  geom_stacks() + 
  geom_mean() + 
  facet_align(sex) + # two samples
  geom_mean_diff() + 
  geom_mean_diff_label()" |>
  ggram::ggram(code = _)
```


## 'Interlude' - What if there was no relationship between vars - shuffling?  Is difference consistent with no relationship?


In 10 shuffles, how many times is the difference under than observed diff 32.21...

```{r, out.width="33%", fig.show='hold'}
data_haircuts |> 
  data_shuffle_var(var = price) |>  # shuffle - no association between vars
  ggplot() + 
  aes(x = shuffled) + #<< shuffled values 
  geom_stacks() +
  geom_mean() + 
  geom_mean_diff() +
  facet_align(sex) + 
  geom_mean_diff() +
  geom_mean_diff_label()


last_plot() + (data_haircuts |> data_shuffle_var(var = price))

last_plot() + (data_haircuts |> data_shuffle_var(var = price))

last_plot() + (data_haircuts |> data_shuffle_var(var = price))

last_plot() + (data_haircuts |> data_shuffle_var(var = price))

last_plot() + (data_haircuts |> data_shuffle_var(var = price))

```


## Step 2. Visualize distribution for null (and standard error)


```{r}
data_haircuts |> 
  count(sex)

data_haircuts |> 
  group_by(sex) |>
  summarise(sd = sd(price))
```


```{r, fig.width=9, echo = F}
'data_haircuts |> 
  ggplot() + 
  aes(outcome = price, 
      sample = sex) + 
  geom_sd_span() +
  stamp_no_diff() +
  stamp_no_diff_label() +
  geom_mean_diff_normalized() +
  geom_mean_diff_normalized_label() +
  stamp_equation() + 
  geom_tdist_null_two_sample()' |>
  ggram::ggram(code = _)
```


# What proportion of the density is as or more extreme than the observed difference 32.2. (What's the area in the tail of the distribution for the null?)

```{r}
t.test(data_haircuts$price ~ data_haircuts$sex, var.equal = T)

t.test(data_haircuts$price ~ data_haircuts$sex)

wilcox.test(data_haircuts$price ~ data_haircuts$sex)
```




```{r, fig.width=9, echo = F}
"data_haircuts |> 
  ggplot() + 
  aes(x = price |> rank()) + 
  geom_rug() +
  geom_stacks() + 
  geom_mean() + 
  facet_align(sex) + # two samples
  geom_mean_diff() + 
  geom_mean_diff_label()" |>
  ggram::ggram(code = _)
```
