---
title: "Untitled"
author: "Dean Data Cell"
date: "5/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 0: Use base ggplo2 to get the job done

```{r cars}
library(tidyverse)
library(gapminder)
gapminder %>%  
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  select(country, pop) ->  
prep  

packcircles::circleProgressiveLayout(prep$pop,  
                                         sizetype = 'area') ->  
pack  

cbind(prep, pack) %>%
  mutate(id = row_number())


pack %>%  
  packcircles::circleLayoutVertices(npoints = 50) ->  
circle_outlines  


circle_outlines %>%  
  ggplot() +  
  aes(x = x, y = y) +  
  geom_polygon(colour = "black", alpha = 0.6) +  
  aes(group = id) +  
  aes(fill = factor(id)) +  
  geom_text(data = cbind(prep, pack),  
            aes(x, y, size = pop, label = country,  
                group = NULL, fill = NULL)) +  
  theme(legend.position = "none") +  
  coord_equal()
```







---

## Step 1: computation

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
# you won't use the scales argument, but ggplot will later
compute_group_circle_pack <- function(data, scales){
 
  data %>%
    mutate(id = row_number()) ->
  data1
 
  data1 %>%  
    pull(area) %>%
    packcircles::circleProgressiveLayout(
                                         sizetype = 'area') %>%
    packcircles::circleLayoutVertices(npoints = 300) %>%
    left_join(data1) #%>%
    # rename(group = id)
   
}


# step 1b test the computation function
gapminder::gapminder %>%
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  # input must have required aesthetic inputs as columns
  rename(area = pop) %>%
  compute_group_circle_pack() %>%
  head()


# step 1b test the computation function
gapminder::gapminder %>%
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  # input must have required aesthetic inputs as columns
  rename(area = pop) %>%
  compute_group_circle_pack() %>% 
  ggplot() + 
  aes(x = x, y = y, fill = country) + 
  geom_polygon()

my_setup_data <- function(data, params){
                                    if(data$group[1] == -1){
                                      nrows <- nrow(data)
                                      data$group <- seq_len(nrows)
                                    }
                                    data
                                  }

```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs?
- where does the function from above go?

```{r}
StatCirclepack <- ggplot2::ggproto(`_class` = "StatCirclepack",
                                  `_inherit` = ggplot2::Stat,
                                  # setup_data = my_setup_data,
                                  required_aes = c("area"),
                                  compute_group = compute_group_circle_pack#,
                                  # default_aes = aes(fill = after_stat(area))
                                  )
```


---

## Step 3: define geom_* function

```{r}
geom_polygon_circlepack <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCirclepack, # proto object from Step 2
    geom = ggplot2::GeomPolygon, # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
gapminder::gapminder %>%
  filter(year == 2002) %>%
  ggplot() +
  aes(area = pop) +
  geom_polygon_circlepack(alpha = .5)

last_plot() +
  facet_wrap(facets = vars(continent)) 

last_plot() +
  aes(fill = country) + 
  guides(fill = "none")
```




```{r my_cars, include = FALSE}
library(ggsankey)
gapminder::gapminder %>% 
  group_by(country) %>%  
  uncount(weights = round(pop/1000000)) %>% 
  arrange(country, year) %>% 
  mutate(lead_year = lead(year)) %>% 
  mutate(lead_continent = lead(continent)) %>% 
  # ungroup() %>% 
  ggplot() + 
  aes(x = year) + 
  aes(next_x = lead_year) + 
  aes(node = continent) + 
  aes(next_node = lead_continent) + 
  geom_alluvial() + 
  aes(fill = continent)

last_plot() %>% layer_data()
```


---


# Try again

## Step 1: computation

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
# you won't use the scales argument, but ggplot will later
compute_group_circle_pack <- function(data, scales){
 
  data

   
}


my_setup_data <- function(data, params){
                                     data %>%
    mutate(id = row_number()) ->
  data1
 
  data1 %>%  
    pull(area) %>%
    packcircles::circleProgressiveLayout(
                                         sizetype = 'area') %>%
    packcircles::circleLayoutVertices(npoints = 300) %>%
    left_join(data1) #%>%
    # rename(group = id)
                                  }

```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs?
- where does the function from above go?

```{r}
StatCirclepack <- ggplot2::ggproto(`_class` = "StatCirclepack",
                                  `_inherit` = ggplot2::Stat,
                                  # setup_data = my_setup_data,
                                  required_aes = c("area"),
                                  compute_group = compute_group_circle_pack#,
                                  # default_aes = aes(fill = after_stat(area))
                                  )
```


---

## Step 4: Enjoy! Use your function

```{r}
gapminder::gapminder %>%
  filter(year == 2002) %>%
  ggplot() +
  aes(area = pop) +
  geom_polygon_circlepack(alpha = .5)

last_plot() +
  facet_wrap(facets = vars(continent)) 

last_plot() +
  aes(fill = country) + 
  guides(fill = "none")
```

---

## Step 3: define geom_* function

```{r}
geom_polygon_circlepack <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCirclepack, # proto object from Step 2
    geom = ggplot2::GeomPolygon, # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
gapminder::gapminder %>%
  filter(year == 2002) %>%
  ggplot() +
  aes(area = pop) +
  geom_polygon_circlepack(alpha = .5)

last_plot() +
  facet_wrap(facets = vars(continent)) 

last_plot() +
  aes(fill = country) + 
  guides(fill = "none")
```


