---
title: "geom_diamond()"
author: "Evangeline Reynolds"
date: "8/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
```


# step 0: get it done with base ggplot2

```{r}
create_diamond <- function(x0, y0, width = 1, height = 1){
  data.frame(
    x = c(x0 + width, x0, x0 - width, x0, x0 + width),
    y = c(y0, y0 + height, y0, y0 - height , y0)
  )
}

cars %>% 
  .[1:5,] %>% 
  mutate(diamond = purrr::map2(speed, dist, create_diamond)) %>% 
  mutate(group = row_number()) %>% 
  unnest() %>% 
  ggplot() + 
  geom_point(data = cars[1:5,], aes(x = speed, y = dist)) + 
  geom_path(aes(x = x, y = y, group = group))

```


# step 1a:  write set up data function



```{r cars}
setup_data_diamonds <- function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  }
```

# step 1b: test set up data function

```{r}
cars %>% 
  .[1:2,] %>% 
  mutate(group = row_number()) %>% 
  setup_data_diamonds()
```



# step 2a: write compute panel function


# step2a.01 write row processing function

# in this case we write a function to process each row for the data

```{r}

create_diamond <- function(x0, y0, width = 1, height = 1){
  data.frame(
    x = c(x0 + width, x0, x0 - width, x0, x0 + width),
    y = c(y0, y0 + height, y0, y0 - height , y0)
  )
}
```


# step2a.02 test row processing function

```{r}
create_diamond(1, 2)
```



# step 2a.03 the compute panel function 

This function takes a full data frame, goes through each of the rows, and uses the row processing function on each, and returns a data frame

```{r}
compute_panel_diamonds <- function(data, scales) {
    
    cols_to_keep <- setdiff(names(data), c("x0", "y0"))
    
    diamonds <- lapply(seq_len(nrow(data)), function(i) {
      
      diamonds_path <- create_diamond(data$x0[i], data$y0[i])
      cbind(diamonds_path, unclass(data[i, cols_to_keep]))
      
    })
    
    do.call(rbind, diamonds)
}
```

# step 2b: Test the compute_panel

```{r}
cars %>% 
  .[1:2,] %>% 
  rename(x0 = speed) %>% 
  rename(y0 = dist) %>% 
  mutate(group = row_number()) %>% 
  compute_panel_diamonds()
```


# Step 3: write ggproto

```{r}
StatDiamond <- ggproto("StatDiamond", Stat, 
  setup_data = setup_data_diamonds,
  compute_panel = compute_panel_diamonds,
  required_aes = c("x0", "y0")
)
```



# Step 4: write c

```{r}
geom_diamond <- function(mapping = NULL, 
                        data = NULL, 
                        stat = "diamond", 
                        position = "identity", 
                        ..., 
                        width = 1, 
                        height = 1,
                        arrow = NULL, 
                        lineend = "butt", 
                        linejoin = "round", 
                        na.rm = FALSE, 
                        show.legend = NA, 
                        inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPolygon, 
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(
      width = width, 
      height = height,
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}
```

# Step 5: Enjoy!

```{r}

set.seed(1244)
cars %>% 
  .[1:5,] %>%
  ggplot() +
  aes(x = speed, y = dist) +
  geom_point() +
  aes(x0 = speed, y0 = dist) +
  geom_diamond(alpha = .2) + 
  aes(fill = speed > 6) + 
  geom_diamond(height = 2, alpha = 0, color = "black")
```



# Step 6: Post Mordem using layer_data


```{r}
layer_data(last_plot(), 2)
```

