---
title: "equation steps"
author: "Evangeline Reynolds"
date: "2/10/2022"
output: html_document
---


## Step 0: use base ggplot2 to get the job done


```{r}
library(tidyverse)

cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_text(aes(label = paste0("x[", id_number, "]")), 
             hjust = 1.2, parse = T)


cars %>% 
  mutate(id_number = 1:n()) %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_text(data = . %>% slice(50), 
            aes(label = paste0('over(1, n) + over(1, (n - 1)) * bgroup("(", (over(x[i]-bar(x),s[x])),")") ^2')), 
             hjust = 1.2, parse = T, check_overlap = T)
```

## Step 1: computation

- define computation that ggplot2 should do for you, before plotting
  - here it's computing a variable with labels for each observation
- test that functionality!

```{r}
# you won't use the scales argument, but ggplot will later
compute_group_row_number <- function(data, scales){
  
  data %>% 
    # add an additional column called label
    # the geom we inherit from requires the label aesthetic
    mutate(label = 1:n())
  
}

# step 1b test the computation function 
cars %>% 
  # input must have required aesthetic inputs as columns
  rename(x = speed, y = dist) %>% 
  compute_group_row_number() %>% 
  head()
```

---

## Step 2: define ggproto

- what's the naming convention for the proto object?
- which aesthetics are required as inputs
- where does the function from above go?

```{r}
StatRownumber <- ggplot2::ggproto(`_class` = "StatRownumber",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("x", "y"),
                                  compute_group = compute_group_row_number)
```


---

## Step 3: define geom_* function

```{r}
geom_text_equation <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatRownumber, # proto object from Step 2
    geom = ggplot2::GeomText, # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```



---

## Step 4: Enjoy! Use your function

```{r}
cars %>% 
  ggplot() + 
  aes(x = speed, y = dist) + 
  geom_point() + 
  geom_text_equation(hjust = 1.2) # function in action
```
