---
title: "adding usable (?) stat_* columns via default_aes"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Would like to be able to add some bonus aes by adding defaults.  

## We see, adding 'known' (anticipated?) aes works nicely (i.e. label)

```{r error = T}
library(tidyverse)
diamonds %>% 
  ggplot() + 
  aes(y = cut) + 
  geom_bar() + 
  stat_count(geom = "text", 
             aes(label = after_stat(count)))

StatCount$default_aes
StatCount$default_aes <- aes(x = after_stat(count),
                             y = after_stat(count),
                             weight = 1,
                             label = after_stat(count),
                             percent = after_stat(100*count/sum(count)) %>% 
                               round(1) %>% 
                               paste0("%"),
                             another_aes = after_stat("bonus aes"))

geom_barlab <- function(...){stat_count(geom = "text", ...)}

diamonds %>% 
  ggplot() + 
  aes(y = cut) + 
  geom_bar() + 
  geom_barlab()
```

# looks like we should be able to use it, inspecting w/ layer_data()

```{r}
layer_data()  # wow!  percent is here.  (but we can't seem to access it)
```

# let's use ggtrace 'layer_all_stages' :-) to see when columns should be available?

```{r}
layer_all_stages <- function(){

a <- ggtrace::layer_before_stat() %>% head(3)
b <- ggtrace::layer_after_stat() %>% head(3)
c <- ggtrace::layer_before_geom() %>% head(3)
d <- ggtrace::layer_after_scale() %>% head(3)

list(before_stat = a, after_stat = b, 
     before_geom = c, after_scale = d)

}



layer_all_stages()
```

# but... throws error when we actually try 'Caused by error in `compute_geom_1()` `geom_text()` requires the following missing aesthetics: label'

```{r, error=T}
# but
diamonds %>% 
  ggplot() + 
  aes(y = cut) + 
  geom_bar() + 
  stat_count(geom = "text", aes(label = after_scale(percent))) 


```



## Closing remarks, Other Relevant Work, Caveats
