---
title: "StatTable: flexible, step-by-step, declarative table creation using ggplot2"
subtitle: "An early 'compute_layer' exercize"
author: "Gina Reynolds"
output:
  xaringan::moon_reader:
    # seal: FALSE
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r, include = F}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(fig.width = 7, message = FALSE, warning = FALSE, comment = "", cache = T)
library(flipbookr)
library(tidyverse)
load("current_image.Rdata")
```

---

# Step 0: Use base ggplot2

---


`r flipbookr::chunk_reveal("base")`


```{r base, include = F}
Titanic %>% 
  data.frame() %>%  
  uncount(Freq) ->
tidy_titanic

tidy_titanic %>% 
  group_by(Sex, Survived) %>%  # cross tab def
  count() %>% 
  group_by(Sex) %>% # scope for proportion
  mutate(prop = n/sum(n)) %>% 
  ggplot() + 
  facet_grid(rows = vars(Sex), cols = vars(Survived)) + 
  geom_text(aes(x = 0, y = 0, label = round(prop, 3)))
```


---


# Step 1.a: Define compute!


---

We will have our cells within ggplot2 panels (facets).  Another route is x and y to define cells, but facet_grid allows you to include an arbitrary number of panels.

---


`r flipbookr::chunk_reveal("compute", break_type = 2)`


```{r compute, eval = T, include = F}
library(tidyverse)

compute_layer_prop_by <- function(data, ...){
  
  if(is.null(data$value)){data$value <- 1}
  if(is.null(data$prop_by)){data$prop_by <- 1}
  if(is.null(data$wt)){data$wt <- 1}
  
  data$value <- data$value * data$wt
  
  data %>% 
    group_by(prop_by) %>% 
    summarize(group_sum = sum(.data$value)) ->
  group_sum_df
  
  data %>% 
    left_join(group_sum_df, by = "prop_by") %>% 
    group_by(PANEL, prop_by) %>% 
    summarize(sum = sum(.data$value),
              group_sum = mean(.data$group_sum)) %>% 
    mutate(prop = sum/group_sum) %>% 
    mutate(prop_label = round(prop, 3)) %>% 
    mutate(percent_label = paste0(prop_label * 100, "%")) %>% 
    mutate(count_percent_label = paste0(sum, "\n(", percent_label, ")" )) %>% 
    mutate(percent_count_label = paste0(percent_label, "\n(", sum, ")" )) %>% 
    mutate(x = 0, y = 0) 

}

```

---

# Step 1.b - Test Compute


---


`r flipbookr::chunk_reveal("compute_test")`


```{r compute_test, include = F}
tidy_titanic %>% 
  mutate(prop_by = Sex) %>% 
  mutate(
    PANEL = 
      paste(as.numeric(Sex),
            as.numeric(Survived),
            sep = "-")) %>%
  ungroup() %>% 
  compute_layer_prop_by()
```

---


# Step 2.a: Define Stat


---

`r flipbookr::chunk_reveal("StatTable", break_type = 2, widths = c(4,1))`


```{r StatTable, include=FALSE}
StatTable <- ggproto(`_class` = "StatTable",
                    `_inherit` = Stat,
                    compute_layer = compute_layer_prop_by,
                    default_aes = aes(label = after_stat(sum)))
```

---

# Step 2.b: Test Stat with layer() (placeholder)

---

`r flipbookr::chunk_reveal("stat_test")`


```{r stat_test, include = F}
tidy_titanic %>% 
  ggplot() + 
  layer(geom = "text", 
        stat = StatTable,
        position = "identity") ->
all

all + 
  facet_grid(rows = vars(Survived),
             cols = vars(Sex)) ->
two_by_two

two_by_two + 
  aes(label = after_stat(percent_label)) ->
two_by_two_percent

two_by_two_percent +
  aes(prop_by = Sex) ->
percent_within_Sex

library(patchwork)
((all + two_by_two)/
    (two_by_two_percent + 
       percent_within_Sex))

```

---

# Step 3.a. Define user-facing function

---

```{r}
stat_table <- function(mapping = NULL, 
                       data = NULL,
                       geom = ggplot2::GeomText,
                       position = "identity", 
                       na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatTable,        # proto object from step 2
    geom = geom,   # inherit other behavior
    data = data, 
    mapping = mapping,
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}



```


---

# Step 3.c some convenience functions


```{r}
geom_table_text <- stat_table
geom_table_tile <- function(){stat_table(geom = "tile")}


and <- paste  # not super careful category combos
```

---

`r chunk_reveal(chunk_name = "the_chunk", break_type = "auto", widths = c(3,2))`


```{r the_chunk, include = F}
tidy_titanic %>% 
  ggplot() + 
  facet_grid(~Sex) + # specify table layout
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        text = element_text(size = 22)) +
  geom_table_tile() + 
  geom_table_text(color = "whitesmoke", 
                  size = 8) + 
  facet_grid(Sex ~ Survived) + 
  aes(label = after_stat(prop_label)) +
  aes(label = after_stat(percent_label)) + 
  aes(prop_by = Sex) + 
  aes(fill = after_stat(prop_by)) + 
  aes(label = after_stat(count_percent_label)) +
  aes(label = after_stat(percent_count_label)) +
  facet_grid(rows = vars(Sex),
             cols = vars(Age, Survived)) + 
  aes(prop_by = NULL) +
  facet_grid(rows = vars(Age)) 
```

---


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```

