---
title: "Spring 2026 Slides"
author: "Evangeline 'Gina' Reynolds"
output:
  xaringan::moon_reader:
    seal: FALSE
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, dpi = 300, dev = "ragg_png")
library(flipbookr)
library(tidyverse)
options(tidyverse.quiet = TRUE)
```


class: inverse, bottom
background-image: url(https://images.unsplash.com/photo-1588591795084-1770cb3be374?q=80&w=2070&auto=format&fit=crop&ixlib=rb-4.1.0&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D)
background-size: cover

## .large[R Medicine 2026: ggregions for building intuitive coding interfaces for anatomical visualizations]
## .small[]
#### .tiny[Dr. Evangeline Reynolds | `r Sys.Date()` | DU, Image credit: Ross Sneddon, Upsplash]


???

Title slide
```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, fig.height = 4.5, message = FALSE, warning = FALSE, comment = "", cache = T)
library(flipbookr)
library(ggplot2)
library(ggregions)
```


---
class: inverse

# Motivations for ggregions

> ### I am told there are people who do not care for maps, and I find it hard to believe. - Robert Louis Stevenson

--

> ### 'My teacher, Mr. Jayson, says a map is a picture of someplace from above.  It's like flying over that spot in an airplane.' - 'Lisa' in Loreen Leedy's in 'Mappying Penny's World'

---

# But ...

> ### I hate when I work with spatial data that so much of my mindshare goes to thinking about 'how do I wrangle this map and what on earth is going on' versus the actual meaning behind what I'm trying to map. - Emily Riederer


---
class: inverse


--

## There is great support for mapping in R and in the data visualization library {ggplot2} which supports many R packages.  

--

## But, producing a simple map, like a choropleth, can feel much harder than producing other plots, like a scatterplot for example in ggplot2. 



```{r, include = F}
library(dplyr)

# Step 1. prep some of their geo reference data
usmapdata::us_map() |>
  select(state_name = full, 
         state_abbr = abbr,
         state_fips = fips,
         geometry = geom) ->
us_states_ref
```


```{r geom_state, echo = F}
# Step 2. use `write_geom_region`

#' @export
geom_state <- write_geom_region(us_states_ref)

#' @export
stamp_state <- write_stamp_region(us_states_ref)

#' @export
geom_state_text <- write_geom_region_text(us_states_ref)

#' @export
stamp_state_text <- write_stamp_region_text(us_states_ref)
```



---

`r chunk_reveal("classic_v_semantic")`

```{r classic_v_semantic, include=F}
tribble(~var1, ~var2, ~var3,
            1,     4,   T,
            2,     3,   F) |> 
  ggplot() + 
  aes(x = var1,
      y = var2) + 
  geom_point(size = 4) +
  aes(color = var3)
```

---

`r chunk_reveal("semantic")`

```{r semantic, include=F}
tribble(~place, ~var2,
       "Texas",     T,
     "Indiana",     F) |> 
ggplot() + 
  aes(state_name = place) + 
  geom_state() + 
  aes(fill = var2)
```



---
class: inverse

## Semantic 

--

## Positional 

--

## Encoding 



---
class: inverse

## Semantic (*meaning* of words or symbols)



## Positional 



## Encoding 

--


### 'normal' ggplot2 positional encoding

--

### `aes(x = c(1, 2), y = c(4, 2))`

--

### S.P.E.

--

### `aes(state_name = c("Texas", "Indiana"))`


<!-- #### ggregions posits that by investing in methods to easily create new geo-specific ggplot2 layer extensions (geom_*s), producing maps with ggplot2 can become as intuitive as creating other plots. About specifying a map, you might say, 'use this data, draw regions, position and color fill according to var1 and var2.' -->

<!-- The proposed regions mapping might look something like this: -->






---
class: inverse, center, middle

# Does {ggregions} deliver?

```{r, eval = T, echo = F}
us_income <- tidyr::us_rent_income |> 
  filter(variable == "income") |>
  rename(income = estimate)

# head(us_income)

library(ggplot2)
(theme_gray() + 
  theme(palette.colour.continuous = "magma")) |> 
  set_theme()
```






---


`r chunk_reveal("usmapdata0")`

```{r usmapdata0, include = F}
library(ggplot2)

us_income |> 
  ggplot() + 
  aes(state_name = NAME) + 
  geom_state() + 
  aes(fill = income)
```


---


> ## [ggregions-generated] APIs feels like they actually puts the "question" at the center. - Emily Riederer

---
class: inverse, center, middle


# How does it work?

---

## Effortlessly write new region-specific geom_* functions.

<!-- The {ggregions} allows API creators to quickly create easy-to-use ggplot2 layer extensions (e.g. geom_*()s) that adhere to the ggplot2 'grammar'. -->


---

```{r, eval = F, echo = F}
usethis::create_package(".", check_name = F)
```

---


`r chunk_reveal("us_states_ref", title = "## Example with geographic regions: US states\n\n### Step 1. prepare data")`


```{r us_states_ref, include = F}
library(dplyr)

# Step 1. prep some of their geo reference data
usmapdata::us_map() |>
  select(state_name = full, 
         state_abbr = abbr,
         state_fips = fips,
         geometry = geom) ->
us_states_ref
```






---

### Step 2. write user-facing functions with ggregion 'write' utilities

<!-- Then, authors will be able use `write_geom_region()` to create a regions-specific ggplot2 layer, for example `geom_state()` below, which can be made available in their geo packages. -->

```{r}
library(ggregions)

geom_state <- 
  write_geom_region(us_states_ref, required_aes = "state_name|state_abbr|state_fips")

stamp_state <- 
  write_stamp_region(us_states_ref, required_aes = "state_name|state_abbr|state_fips")

geom_state_text <- 
  write_geom_region_text(us_states_ref, required_aes = "state_name|state_abbr|state_fips")

stamp_state_text <- 
  write_stamp_region_text(us_states_ref, required_aes = "state_name|state_abbr|state_fips")

```


---

### Step 2b. (optional) Inspect user-facing function

```{r}
geom_state
```


---

#### Steb 2bb (for the truely curious) Inspect StatRegions compute

```{r}

ggregions:::StatRegion$compute_panel

```

---

### What does `ggregions::write_geom_region()` do?


```{r}
ggregions::write_geom_region
```

--

```{r}
ggregions:::geom_region
```



<!-- ## Deliver intuitive, newcomer-welcoming spatial viz experience -->

<!-- By including the newly specified `geom_` function in their geo data package (perhaps instead or in addition to a convenience wrapper that are typical of geo packages) the package *users* will be able to use the` data + aes + geom` formulation for defining their plots.  Instead of positional aesthetics like x, y or geometry, a region identifier like `state_name` or `fips` can be used.  Internally, the regions `geom_*` will translate between place name or other identifier to the necessary borders (`sfc`s). -->

<!-- Below is some tabular data with geographic region identifiers, but no boundaries: -->

<!-- Even though there is no boundary information, analysts can map this data with the familiar  data + aesthetic mapping + layer syntax that makes ggplot2 easy to read, write, and reason about. -->




---




`r chunk_reveal("usmapdata")`

```{r usmapdata, include = F}
library(ggplot2)
library(usmapdata) 

us_income |> 
  ggplot() + 
  aes(state_name = NAME) + 
  geom_state() + 
  aes(fill = income) +
  scale_fill_viridis_c(option = "magma") + 
  geom_state_text(color = "white") + 
  aes(label = after_stat(state_abbr))
```

<!-- Packages that might take avantage of ggregion's functionality include -->
<!-- [afrimapr](https://afrimapr.github.io/afrimapr.website/), [geobr](https://ipeagit.github.io/geobr/), [cancensus](https://mountainmath.github.io/cancensus/index.html), [chilemapas](https://pacha.dev/chilemapas/), [RCzechia](https://github.com/jlacko/RCzechia), [geofi](https://ropengov.github.io/geofi/) -->
<!-- , [geokz](https://github.com/arodionoff/geokz), [mapsPERU](https://github.com/musajajorge/mapsPERU) and [geoidep](https://geografo.pe/geoidep/index.html), [mapSpain](https://github.com/rOpenSpain/mapSpain/), [geographr](https://github.com/britishredcrosssociety/geographr), [geouy](https://github.com/RichDeto/geouy), [tigris](https://github.com/walkerke/tigris), -->
<!-- and [rgeoboundaries](https://github.com/wmgeolab/rgeoboundaries). -->
<!--  Since these geo packages have many users, the impacts of `{ggregions}` could be significant. -->






---
class: inverse, center, middle

# Anatomical regions



---

### ggregions with teethr::dental_arcade_mapping

https://github.com/bbartholdy/teethr

---

### data preparation

```{r}
teethr::dental_arcade_mapping |> 
  as_tibble() |> 
  left_join(teethr::tooth_notation |> rename(tooth = text)) |>
  select(tooth_id = tooth, fdi = FDI, standard = standards, geometry) ->
teeth_ref_data
```

--


### ggplot2 'layer' preparation

```{r}  
library(ggregions)
geom_tooth <- write_geom_region(teeth_ref_data, required_aes = "tooth_id|fdi|standard")
geom_tooth_text <- write_geom_region_text(teeth_ref_data, required_aes = "tooth_id|fdi|standard")
stamp_tooth <- write_stamp_region(teeth_ref_data, required_aes = "tooth_id|fdi|standard")
stamp_tooth_text <- write_stamp_region_text(teeth_ref_data, required_aes = "tooth_id|fdi|standard")
```


```{r, echo = F}
library(teethr)

caries_ratios <- mb11_caries %>% 
  dental_longer(-id) %>%
  dental_join() %>% 
  count_caries(caries = score, no_lesion = "none") %>% # convert location to lesion count
  group_by(tooth) %>% 
  dental_ratio(count = caries_count) %>%
  dental_recode(tooth, "FDI", "text") 

```


---

`r chunk_reveal("arcade", title = "### yields user-friendly API" )`


```{r arcade, include = F}
caries_ratios |> 
  ggplot() + 
  aes(tooth_id = tooth, 
      fill = ratio) + 
  geom_tooth(alpha = .5) + 
  stamp_tooth_text(size = 2) + 
  scale_fill_viridis_c()
```

---


```{r, echo = F, fig.width= 9, fig.height=5, out.width="65%"}

"
library(teethr)

caries_ratios |> 
  ggplot() + 
  aes(tooth_id = tooth, #<<
      fill = ratio) + 
  geom_tooth(alpha = .5) + #<<
  stamp_tooth_text(size = 1.5) + #<<
  scale_fill_viridis_c()" |>
  ggram::ggram(code = _,
               subtitle =  "🚧🧫🚧🔬🚧🧪 Checking out caries ratios in this amazing {teethr} tooth arcade... Uh-oh LLM1!",
               caption =   "A {teethr} X {ggregions} experiment: ggregions helps build region-specific geom_*()s")

```

---

## ggregions X gganatogram::hgFemale_list


```{r echo = F}
to_sf_routine <- function(data){
  
  data |>
  mutate(y = -y) |>
  sf::st_as_sf(coords = c("x", "y"), agr = "constant") |>
  group_by(id, group) |>
  summarize(do_union = F) |> 
  ungroup() |> 
  group_by(id, group) |>
  summarise() |>
  mutate(geometry = geometry |> sf::st_cast("POLYGON")) |> 
  mutate(geometry = geometry |> sf::st_cast("MULTIPOLYGON")) |> 
  ungroup() 
  
}
```

### Step 1: Data preparation


```{r, error = T, echo = T}
female_sf <- gganatogram::hgFemale_list[c(1:156, 180:195)] |> # return to this!!
  bind_rows() |>
  remove_missing() |>
  to_sf_routine()

ref_female_sf <- female_sf |> 
  select(organ = id, geometry)
```

--

### Step 2: User-facing function preparation

```{r, }  
library(ggregions)
stamp_organ <- write_stamp_region(ref_female_sf, required_aes = "organ")
geom_organ <- write_geom_region(ref_female_sf, required_aes = "organ")
stamp_organ_text <- write_stamp_region_text(ref_female_sf, required_aes = "organ")
geom_organ_text <- write_geom_region_text(ref_female_sf, required_aes = "organ")
```

---

`r chunk_reveal("gganatogram", title = "Enjoy!")`

```{r gganatogram, include = F}
tribble(~my_organ, ~color,   
        "stomach", "cadetblue",
        "brain",   "pink3",
        "colon",   "darkseagreen4",
        "lung",    "plum",
        "heart",    "coral") |> 
ggplot() + 
  stamp_organ(alpha = .2) +
  aes(organ = my_organ) +
  geom_organ() +
  aes(fill = I(color)) + 
  geom_organ_text() + 
  ggplyr:::data_slice(5)
```


---


class: inverse, bottom
background-image: url(https://encrypted-tbn3.gstatic.com/shopping?q=tbn:ANd9GcTSNaV2hzHW71ovNcoPji1OZMZBz_EJQxwCRClzMeX-QmKTiP5OpNcMvuIVE7fmgObCuOsrcRLUMUxKG2P2kZSIkrSf8Puhpg)
background-size: contain

# Thank you!
github:

## EvaMaeRey/ggregions


---




## Unresolved questions for this project:


#### 1. Perhaps an *alternative to sf* would make sense for anatomical regions - we don't really anticipate *'crs'* to be specified.  Would allowing another preparation/Stat make sense - polygon-based?  Or just more attention to converting polygons to sf (relates to #3).

--

#### 2. The gganatogram objects bring up the fact that regions *overplot* - the heart is covered by the lungs.  So should there be a way to control the ordering (And Even if there is no regional overlap, overplotting can happen if we have the region more than once in the dataset)

--

#### 3. ggregions' focus is on gg-type, friendly API, *but how much focus should be on preparing reference data.*  The teethr and colon case might help here. 

--

#### 4. ggregions allows you to 'try-on' the ggregions framework (positional aesthetics that are symantic, 'texas' can be the answer to 'where'?).  But, *what if you don't want to take the ggregions dependency?*  Should there be more instruction regarding this?  Licensing considerations? 

--

#### 5. More relevant to the geographic case, it might be useful to allow different resolution of the regions reference data.  How important is addressing this a this point? 

--

#### 6. Can this project strengthen the R ecosystem? Motivation to get into packaging?  How do we make that clearer, and the pathway easier? 


<!-- adjust font size in this css code chunk, currently 80 -->

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 70%}

@media print {
  .has-continuation {
    display: block;
  }
}


```

<!-- code.r.hljs.remark-code{ -->
<!--   position: relative; -->
<!--   overflow-x: hidden; -->
<!-- } -->


<!-- code.r.hljs.remark-code:hover{ -->
<!--   overflow-x:visible; -->
<!--   width: 500px; -->
<!--   border-style: solid; -->
<!-- } -->

