---
title: "another experiment"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```




## Intro Thoughts


## Status Quo

```{r cars}
library(tidyverse)


box_df <- tribble(
~ task, ~minutes,
"start timeboxing", 1,
"red rocks yoga confirmation", 5,
"find docs quote code", 5,
"grab tidytuedsday data url", 5,
"plan meet up w mom", 1, 
"mother's day ideas", 5,
"teachers appreciation ideas", 10, 
"tidypivot motivation", 20,
"ggverbatim motivation", 20,
"R Ladies Denver minimal viable package", 20,
"ggexpress as discussion", 15,
"break and dinner prep", 15,
"Break - kitchen cleanup", 1,
"ggcallout step 0 work", 20,
"pp docs", 5,
"ggtimebox", 20
)

```

## Experiment

```{r, fig.height=9}
# box_df |>
# ggtimebox:::compute_panel_timebox()

box_df %>% 
ggplot() + 
  aes(task = task, minutes = minutes) + 
  ggtimebox:::stat_timebox(alpha = .2, fill = "cadetblue", color = "white") + 
  ggtimebox:::stat_timebox(geom = "text", hjust = 0, vjust = 1) + 
  ggtimebox:::stat_timebox(geom = "text", aes(label = after_stat(clock_hour_minute)), hjust = 1, vjust = 1) + 
  ggtimebox:::coord_timebox() + 
  ggtimebox:::stamp_currenttime() + 
  expand_limits(x = -.5) + 
  ggstamp::theme_void_fill(fill = "snow")
```

# red rocks

https://www.redrocksonline.com/yoga/

# find docs quote code

I think I can find that in 'featurette'

https://evamaerey.github.io/featurette/2023-11-02-help-quote-mtcars/help-quote-mtcars.html#9

https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

```{r}
# tidytuesday data
library(magrittr)
tmp <- tempfile()
help(mtcars, package = datasets) %>%
  as.character() %>%
  utils:::.getHelpFile() %>%
  tools:::Rd2latex(Rd = ., out = tmp)
pandoc::pandoc_convert(file = tmp,
                       from = "latex",
                       to = 'markdown') ->
  md


```


# TidyTuesday url

```{r}

```

# ggverbatim

function to modify values (round, good values, in sds?)


# mother's day

books. Brunch. card.

# Teachers.

M. Negro.  Orquidas. Vanilla. Basquet Ball, Olive Garden, Macys, Mango, Ordenar la casa, salir con familia

I. Pink, Chocolate, Roses, Cheetos, Cake, Chiles, chick-fil-et, Starbucks, Strawberry, Banana, Target/Walmart.  For class: Puzzles, books





#  Minimal viable package

We're going to move to a two-track model! Develop without the github assumption, or w/ it. 




## Closing remarks, Other Relevant Work, Caveats
