---
title: "pipeable, declarative table building"
author: "Evangeline Reynolds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

# synopsis

Proposes group, compute and pivot to table in a succinct function, eg: 

`pivot_count(data = titanic, rows = survived, cols = c(sex, age))`

Uses R6 to update table specification and creates user facing functions to incrementally build tables, eg:

```{r, eval = F}

ggtable(data = tidytitanic::tidy_titanic) |> # overall count
  set_rows(sex) |> # within sex
  set_cols(age) |> # within age and sex
  set_rows(NULL) |> # overwrites previous declarations
  set_rows(c(sex,survived)) |> 
  set_cols(c(age, class))

```




# all in one count and pivot function.

Here we write a nice declarative (describe it and get it) table making function... This one just counts.  More functionality is worked out in the pivotr function here: https://github.com/EvaMaeRey/tidypivot. 

```{r}
pivot_count <- function(data, rows = NULL, cols = NULL){
  
  cols_quo <- rlang::enquo(cols)
  rows_quo <- rlang::enquo(rows)

  if(rlang::quo_is_null(cols_quo) & rlang::quo_is_null(rows_quo)){
    data %>% 
      count() ->
    outdata
    
  }else{
  data %>% 
    group_by(across(c({{rows}}, {{cols}}))) %>% 
    count() ->
  outdata
  }
  
  if(!rlang::quo_is_null(cols_quo)){
   outdata %>%
      pivot_wider(names_from = {{cols}}, values_from = n) ->
    outdata
  }
  
  outdata
}
```

## let's use it on a version of the Titanic data ('titanic')  {.tabset}


```{r}
library(tidyverse)
Titanic %>% 
  data.frame() %>% 
  uncount(Freq) ->
titanic

names(titanic) <- tolower(names(titanic))

head(titanic)
```

### Example 1

```{r}
pivot_count(data = titanic, rows = survived, cols = sex)
```

### Example 2

```{r}
pivot_count(data = titanic, cols = sex)
```

### Example 3

```{r}
pivot_count(data = titanic) 
```

### Example 4

```{r}
pivot_count(data = titanic, rows = survived, cols = c(sex, age))
```

# Making declarations incremental with OOP

But we might want to do step-wise (piped) declarations, this will require OOP.  I choose R6 encapsulated/classical.  I think people find it easier just starting off as is my case!


```{r}
library(R6)
Pivot <- R6Class("Pivot",
  public = list(
    data = NULL,
    rows = NULL,
    cols = NULL,
    value = NULL,
    wt = NULL,
    fun = NULL,
    initialize = function(data = NULL, rows = NULL, cols = NULL, 
                          value = NULL, wt = NULL,
                          fun = NULL){
      self$data <- data
      self$rows <- rows
      self$cols <- cols
      self$value <- value
      self$wt <- wt
      self$fun <- fun
      
      theData <<- data   # hacky?

    }, 
    set_data = function(val){
      self$data <- val
      
      theData <<- self$data  # hacky?

    },
    set_rows = function(val){
      self$rows <- val
    },
    set_cols = function(val){
      self$cols <- val
    },
    set_value = function(val){
      self$value <- val
    },
    set_wt = function(val){
      self$wt <- val
    },
    set_fun = function(val){
      self$fun <- val
    },
    return_table_code = function(){

      # out <- 'pivot_count(data = theData, rows, cols)' 
      
      out <- 'tidypivot::pivot_helper(data = theData, rows, cols, value, wt, fun)'
      
      str_replace_or_null <- function(x, pattern, replacement){
        
        if(is.null(replacement)){x %>% str_replace(pattern, 'NULL')}
        else{x %>% str_replace(pattern, replacement)}
        
      }
      
      # out <- str_replace(out, "data", self$data)
      out <- str_replace_or_null(out, "rows", self$rows)
      out <- str_replace_or_null(out, "cols", self$cols)
      out <- str_replace_or_null(out, "fun", self$fun)
      out <- str_replace_or_null(out, "value", self$value)
      out <- str_replace_or_null(out, "wt", self$wt)
      out <- str_replace_or_null(out, "fun", self$fun)

      eval(parse(text = out))    # hacky?
      
    }
    
    ),
  
  
  )

```


## Let's try it out {.tabset}

### Step 1: Declare new table specification

```{r}
table_specification <- Pivot$new()
table_specification
```


### Step 2: Set data

```{r}
table_specification$set_data(tidytitanic::tidy_titanic)
table_specification
table_specification$return_table_code()
```


### Step 3: Set Rows

```{r}
table_specification$set_rows("survived")
table_specification
table_specification$return_table_code()
```


### Step 4: Set Cols

```{r}
table_specification$set_cols("sex")
table_specification
table_specification$return_table_code()
```


### Step 5: Over write which var defines rows

```{r}
table_specification$set_rows("age")
table_specification
table_specification$return_table_code()
```




# Now try to create user-facing interface for piping

```{r}

ggtable <- function(data, rows = NULL, cols = NULL){

  # ts <- Pivot$new()
  ts$initialize()
  ts$set_data(data)
  ts$set_rows(deparse(substitute(rows)))   # hacky?
  ts$set_cols(deparse(substitute(cols)))

  print(ts$return_table_code())
  
  invisible(ts)

}


set_rows <- function(piping_spacer, # hacky?
                     vars){
  
  ts$set_rows(val = deparse(substitute(vars)))
  ts$return_table_code()
  
  print(ts$return_table_code())
  
  invisible(ts)

}


set_cols <- function(piping_spacer, vars){
  
  ts$set_cols(val = deparse(substitute(vars)))
  ts$return_table_code()

  print(ts$return_table_code())
  
  invisible(ts)

}


set_value <- function(piping_spacer, vars){
  
  ts$set_value(val = deparse(substitute(vars)))
  ts$return_table_code()

  print(ts$return_table_code())
  
  invisible(ts)

}

set_wt <- function(piping_spacer, vars){
  
  ts$set_wt(val = deparse(substitute(vars)))
  ts$return_table_code()

  print(ts$return_table_code())
  
  invisible(ts)

}

set_fun <- function(piping_spacer, fun){
  
  ts$set_fun(val = deparse(substitute(fun)))
  ts$return_table_code()

  print(ts$return_table_code())
  
  invisible(ts)

}


last_table <- function(){
  
  print(ts$return_table_code())
  
  invisible(ts)
  
}

```

## Try it out {.tabset}

### Step 1: Set data featuring ugly prestep

```{r}
# it seemed like I needed this before any of the following would get going
ts <- Pivot$new()    # hacky? ugly?

ggtable(data = tidytitanic::tidy_titanic)
```

### Step 2: Set rows

```{r} 
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) 
```

### Step 3: Set cols

```{r}
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) |>
  set_cols(survived)
```

### Step 4: Reset cols

```{r}
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) |>
  set_cols(survived) |>
  set_cols(age)
```


### Step 5: NULL rows

```{r}
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) |>
  set_cols(survived) |>
  set_cols(age) |>
  set_rows(NULL)
```

### Step 6: Reset to multiple rows 

```{r}
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) |>
  set_cols(survived) |>
  set_cols(age) |>
  set_rows(NULL) |>
  set_rows(c(sex,survived))
```


### Step 7: Reset to multiple cols 

```{r tablebuildup}
ggtable(data = tidytitanic::tidy_titanic) |>
  set_rows(sex) |>
  set_cols(survived) |>
  set_cols(age) |>
  set_rows(NULL) |>
  set_rows(c(sex, survived)) |>
  set_cols(c(age, class))
```


### Step 8: not piping

```{r}
ggtable(data = tidytitanic::tidy_titanic, age, class)

```


### Step 9: Another dataset: mtcars

```{r}
ggtable(data = mtcars, rows = cyl)
```


### Step 9: using last_table() to print a family of tables


```{r, error = T}
ggtable(data = titanic)

last_table() |>
  set_cols(sex)

last_table() |>
  set_rows(survived)

last_table() |>
  set_cols(NULL)
```

### Step 10: Definitely not doing print the right way...

print is not done in the right way yet. A collect function will be nice to output the printed table, versus the specification.

```{r}
my_table <- ggtable(data = mtcars, rows = cyl) # shouldn't print

str(my_table) # object is specification yay!

my_table |>       # pipes in okay way.
  set_cols(am)
```


### Step 11: Nice to have in the future... on the fly variable definition - like ggplot2


```{r, eval = F}
ggtable(data = mtcars, rows = cyl)

last_table() |>
  set_cols(ifelse(am, "manual", "auto"))


```

## Functions, weighting


```{r}
ts <- Pivot$new()    # hacky? ugly?

ggtable(mtcars) |>
  set_value(mpg)


```




# References

- https://r6.r-lib.org/articles/Introduction.html#fields-containing-reference-objects
- https://adv-r.hadley.nz/r6.html

