---
title: "{codehover} x {ggcube}"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F, 
                      dpi = 300,
                      cache = T)
library(codehover)
library(tidyverse)
library(ggcube)


```

`r chunk_code_hover(chunk_name = "my_cars")`


```{r my_cars, eval=FALSE, echo=FALSE}
ggplot(mountain) +
  aes(x = x) + 
  aes(y = y) + 
  coord_3d() +
  aes(z = 0) +
  geom_surface_3d() + 
  aes(fill = z) +
  scale_fill_viridis_c() + 
  aes(z = z) +
  guides(fill = guide_colorbar_3d()) +
  theme_light()
```




```{css, echo = F}
.column {
  float: left;
  width: 50%;
  padding: 10px;
  background-color:white;
}
```


```{css, echo = F}
table {
  cursor: pointer; 
  cursor: hand;
  border: none;
  margin-left: auto;
  margin-right: auto;
  width:"80%";
  border:"0"; 
  cellspacing:"0"; 
  cellpadding:"0";
  font-family: "Lucida Console";
}

.hover {
  background-color: #F1D95A92;
}

.image-crop {
  height: 600px;;
}

tab1 { padding-left: 4em; }
tab2 { padding-left: 8em; }
tab3 { padding-left: 12em; }
tab4 { padding-left: 16em; }
tab5 { padding-left: 20em; }
tab6 { padding-left: 24em; }
tab7 { padding-left: 28em; }
tab8 { padding-left: 32em; }
tab9 { padding-left: 36em; }
tab10 { padding-left: 40em; }
tab11 { padding-left: 44em; }
tab12 { padding-left: 48em; }
tab13 { padding-left: 52em; }
tab14 { padding-left: 56em; }
tab15 { padding-left: 60em; }
tab16 { padding-left: 64em; }



```




<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.5.1/jquery.min.js"></script>
<script>

$('.pipehover_incremental tr').hover(function() {
  $(this).removeClass()
  $(this).prevAll().removeClass()
  $(this).nextAll().removeClass()
  $(this).addClass('hover');
  $(this).prevAll().addClass('hover');
  $(this).closest('div').next().find('img').attr("src", $(this).attr("link"));
});


$('.pipehover_select_one_row tr').hover(function() {
  $(this).removeClass()
  $(this).prevAll().removeClass()
  $(this).nextAll().removeClass()
  $(this).addClass('hover');
  $(this).closest('div').next().find('img').attr("src", $(this).attr("link"));
});

</script>
